### 2017-10-27 paper/tables/descriptiveStatisticsPartisanIN.tex

% latex table generated in R 3.4.1 by xtable 1.8-2 package
% Fri Oct 20 13:41:38 2017
\begin{table}[ht]
\centering
\begin{tabular}{rrr}
  \hline
 & Democratic & Republican \\ 
  \hline
Cities & 15 & 17 \\ 
  Documents & 10257 & 5859 \\ 
  Tokens & 6101752 & 2310072 \\ 
  Token assignments & 6006202 & 2259362 \\ 
  Topics & 103 & 97 \\ 
   \hline
\end{tabular}
\caption{Descriptive statistics for Indiana. ``Tokens'' describes the number
             of words in each party's documents, ``token assignments'' the tokens assigned
             to each party in the topic model depending on the ratio of Democratic to Republican 
             tokens in it (both weighted by the total number of tokens per party).} 
\label{tabDescriptiveIN}
\end{table}