### 2017-11-15 paper/tables/stmTopWordsINDem.tex
improvements to stm, hclust, and some other stuff

% latex table generated in R 3.4.2 by xtable 1.8-2 package
% Mon Nov 13 15:36:58 2017
\begin{table}[ht]
\centering
\begin{tabular}{llllllll}
  \hline
-0.026 & -0.019 & -0.018 & -0.018 & -0.018 & -0.013 & -0.011 & -0.011 \\ 
  \hline
fort & propos & said & prosecutor & digest & fund & servic & plan \\ 
  citi & author & ask & charg & introduc & grant & inform & develop \\ 
  ordin & district & state & feloni & author & budget & board & communiti \\ 
  approv & street & will & counti & counti & counti & applic & citi \\ 
  purchas & public & chair & case & appoint & state & project & neighborhood \\ 
  depart & control & propos & crime & board & feder & approv & program \\ 
  properti & amend & year & crimin & approv & depart & technolog & provid \\ 
  will & intersect & move & offic & district & appropri & system & public \\ 
  resolut & counti & need & victim & fund & increas & manag & resourc \\ 
  contract & committe & council & sentenc & street & agenc & agreement & area \\ 
   \hline
\end{tabular}
\caption{Top Democratic topics and words (Indiana), according to STM. 
The words are the top words for the most Democratic/Republican topic, determined
by the size (and significance) of the coefficient (see table header) of the party covariate.} 
\label{tabSTMINDem}
\end{table}