### 2017-11-15 paper/tables/stmTopWordsLADem.tex

% latex table generated in R 3.4.2 by xtable 1.8-2 package
% Mon Nov 13 18:17:05 2017
\begin{table}[ht]
\centering
\begin{tabular}{llllllll}
  \hline
-0.144 & -0.077 & -0.039 & -0.02 & -0.018 & -0.015 & -0.012 & -0.011 \\ 
  \hline
stream & obj & otherwis & polic & wherea & type & communiti & art \\ 
  length & ascent & provid & crime & ordin & content & neighborhood & call \\ 
  predictor & flag & respect & investig & citi & width & citi & cost \\ 
  acrobat & descent & thereto & offic & resolut & pdf & program & sponsor \\ 
  librari & encod & author & arrest & provid & height & public & free \\ 
  pdf & interpol & citi & inform & code & filter & develop & amp \\ 
  ash & matrix & district & suspect & street & parent & opportun & music \\ 
  hum & kid & ordin & victim & shall & count & particip & artist \\ 
  gym & baa & amend & block & chapter & resourc & creat & museum \\ 
  parent & filter & locat & street & provis & size & provid & home \\ 
   \hline
\end{tabular}
\caption{Top Democratic topics and words (Louisiana), according to STM. 
             The words are the top words for the most Democratic/Republican topic, determined
             by the size (and significance) of the coefficient (see table header) of the party covariate.} 
\label{tabSTMLADem}
\end{table}