### 2017-11-15 paper/tables/stmTopWordsLARep.tex

% latex table generated in R 3.4.2 by xtable 1.8-2 package
% Mon Nov 13 18:17:04 2017
\begin{table}[ht]
\centering
\begin{tabular}{llllllll}
  \hline
0.085 & 0.062 & 0.062 & 0.057 & 0.021 & 0.019 & 0.018 & 0.017 \\ 
  \hline
event & ordin & citi & flood & west & properti & park & councilman \\ 
  park & bid & mayor & emerg & north & plan & street & said \\ 
  inform & summari & said & citi & east & zone & citi & citi \\ 
  food & amount & year & storm & south & lot & improv & mayor \\ 
  music & citi & state & resid & traffic & applic & recreat & want \\ 
  show & depart & announc & weather & boulevard & shall & work & council \\ 
  visit & public & work & inform & parish & requir & public & just \\ 
  market & contract & offici & hurrican & drive & approv & repair & think \\ 
  weekend & approv & resid & disast & road & use & program & vote \\ 
  begin & accord & million & power & locat & site & includ & thank \\ 
   \hline
\end{tabular}
\caption{Top Republican topics and words (Louisiana), according to STM. 
             The words are the top words for the most Democratic/Republican topic, determined
             by the size (and significance) of the coefficient (see table header) of the party covariate.} 
\label{tabSTMLARep}
\end{table}