### 2017-12-19 paper/presentation/Markus_SPSA_2018_Presentation.tex
updates to the presentation

\documentclass[12pt]{beamer}
\setbeamertemplate{navigation symbols}{}
%\usepackage[latin1]{inputenc}
\usepackage{pgfplots}
\usepackage{adjustbox}
\usepackage{graphicx}
%\usepackage{draftwatermark}
\usepackage{tikz}
\usepackage{colortbl}
\usepackage{subfigure}
%\usetheme{Amsterdam}
\usecolortheme{default}
\title[]{Government websites as data: A methodological pipeline for
collection, processing, and text analysis}
\author{Markus Neumann \\ Fridolin Linder \\ Bruce Desmarais}
\institute{The Pennsylvania State University}
\date{January 6, 2018}

\makeatletter
\defbeamertemplate*{footline}{myminiframes theme}
{%
	\begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
	\end{beamercolorbox}
	\hbox{%
		\begin{beamercolorbox}[wd=.1\paperwidth,ht=2.5ex,dp=1.125ex,%
			leftskip=.3cm,rightskip=.3cm,center]{title in head/foot}%
			{\usebeamerfont{author in head/foot}\usebeamercolor[fg]{author in head/foot}}{\insertframenumber}%/\inserttotalframenumber}%
		\end{beamercolorbox}%
		\begin{beamercolorbox}[wd=.75\paperwidth,ht=2.5ex,dp=1.125ex,%
			leftskip=.3cm,rightskip=.3cm plus1fil,center]{title in head/foot}%
			\leavevmode{\usebeamerfont{title in head/foot}\insertshorttitle}%
		\end{beamercolorbox}%
		\begin{beamercolorbox}[wd=.15\paperwidth,ht=2.5ex,dp=1.125ex,%
			leftskip=.3cm,rightskip=.3cm plus1fil,center]{title in head/foot}%
			{\usebeamerfont{author in head/foot}\usebeamercolor[fg]{author in head/foot}}%\insertshortauthor}
		\end{beamercolorbox}%
	}%
	\begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
	\end{beamercolorbox}
}
\makeatother

\begin{document}

\section{Introduction}
\begin{frame}
\titlepage
\end{frame}

%\begin{frame}{Spot the difference}
%\begin{figure}
%	\hfill
%	\subfigure{\includegraphics[width=5cm]{Images/Clinton_Rally_cr.jpg}}
%	\hfill
%	\subfigure{\includegraphics[width=5cm]{Images/Trump_Rally_cr.jpg}}
%\end{figure}
%\end{frame}

\begin{frame}{Government Websites}
\linespread{1.5}
{\large
	\begin{itemize}
		\item Content of government websites is an important source of information \& transparency
		\item After coming into power, the Trump administration  has made some controversial changes to the websites of federal agencies
		\item Website content is political
		\item Partisanship of city government is expected to have an effect
	\end{itemize}

}
\end{frame}

\begin{frame}{Example Website}
	\includegraphics[width = \linewidth]{arcadia_water.png}
\end{frame}

\begin{frame}{Local Government Websites}
\linespread{1.5}
{\large
	\begin{itemize}
		\item Most local (i.e. mayoral) elections are non-partisan
		\item Few states have exclusively partisan local elections
		\item Data for these elections can be difficult to find
		\item We selected Indiana and Louisiana
	\end{itemize}

	}
\end{frame}

\begin{frame}{Data Overview}
\begin{figure}
	\hfill
	\subfigure{\includegraphics[width=5cm, height = .7\textheight]{../figures/indiana_map.pdf}}
	\hfill
	\subfigure{\includegraphics[width=5cm, height = .7\textheight]{../figures/louisiana_map.pdf}}
\end{figure}
\end{frame}

\begin{frame}{The Pipeline}
	\textbf{Data Collection} \hspace{2mm} $\rightarrow$ \hspace{2mm} \textbf{Preprocessing} \hspace{2mm} $\rightarrow$ \hspace{2mm} \textbf{Analysis}
\begin{columns}
	\column{0.33\textwidth}
	%Data Collection
	\vspace{5mm}
	\begin{itemize}
		\item Identify URLs
		\item Verify URLs (browser automation)
		\item Download websites
		\item Determine file type
		\item Convert to txt
	\end{itemize}		
	\column{0.33\textwidth}
	%Preprocessing
	\vspace{5mm}
	\begin{itemize}
		\item Remove punctuation, dates, etc.
		\item To lowercase
		\item Boilerplate removal
		\item Spellchecking
		\item Lemmatization \\ (city \& cities = city)
	\end{itemize}
	\column{0.33\textwidth}
	%Analysis
	\vspace{5mm}
	\begin{itemize}
		%\item Hierarchical clustering
		\item Fightin' Words
		\item Structural topic model
		%\begin{itemize}
		%	\item LDA
		%	\item STM
		%\end{itemize}
	\end{itemize}
\end{columns}
\end{frame}

\begin{frame}{The Pipeline}
	\textbf{Data Collection} \hspace{2mm} $\rightarrow$ \hspace{2mm} \textbf{Preprocessing} \hspace{2mm} $\rightarrow$ \hspace{2mm} \textbf{Analysis}
	\begin{columns}
		\column{0.33\textwidth}
		%Data Collection
		\vspace{5mm}
		\begin{itemize}
			\item Identify URLs
			\item Verify URLs (browser automation)
			\item Download websites
			\item Determine file type
			\item Convert to txt
		\end{itemize}		
		\column{0.33\textwidth}
		%Preprocessing
		\vspace{5mm}
		\begin{itemize}
			\item Remove punctuation, dates, etc.
			\item To lowercase
			\item \textbf{Boilerplate removal}
			\item Spellchecking
			\item Lemmatization \\ (city \& cities = city)
		\end{itemize}
		\column{0.33\textwidth}
		%Analysis
		\vspace{5mm}
		\begin{itemize}
			%\item Hierarchical clustering
			\item Fightin' Words
			\item Structural topic model
			%\begin{itemize}
			%	\item LDA
			%	\item STM
			%\end{itemize}
		\end{itemize}
	\end{columns}
\end{frame}

\section{Preprocessing}

%\begin{frame}{Document filetype \& conversion}
%\begin{itemize}
%	\item Websites contain a number of different filetypes
%	\item txt, pdf, html, doc and docx are useable
%	\item convert everything to txt
%	\item the file ENDING does not always 
%\end{itemize}
%\end{frame}

\begin{frame}{Boilerplate Removal}
\includegraphics[width = \linewidth]{arcadia_water.png}
\end{frame}

\begin{frame}{Boilerplate Removal}
\includegraphics[width = \linewidth]{duplicateLines.png}
\end{frame}

\begin{frame}{Boilerplate Removal}
	\includegraphics[width = \linewidth]{duplicateLineRemoval.png}
\end{frame}

\begin{frame}{Boilerplate Removal}
\begin{itemize}
	\item Within each city, there is a lot of shared text
	\item If not removed, the text clusters into cities
\end{itemize}
\begin{itemize}
	\item Solution: Compare each line in each document to every other line in every document of that city
	\item Count duplicates
	\item Remove a line if it is duplicated within a city above some threshold
	\item Fast \& efficient implementation
\end{itemize}
\end{frame}

\begin{frame}{The Pipeline}
	\textbf{Data Collection} \hspace{2mm} $\rightarrow$ \hspace{2mm} \textbf{Preprocessing} \hspace{2mm} $\rightarrow$ \hspace{2mm} \textbf{Analysis}
	\begin{columns}
		\column{0.33\textwidth}
		%Data Collection
		\vspace{5mm}
		\begin{itemize}
			\item Identify URLs
			\item Verify URLs (browser automation)
			\item Download websites
			\item Determine file type
			\item Convert to txt
		\end{itemize}		
		\column{0.33\textwidth}
		%Preprocessing
		\vspace{5mm}
		\begin{itemize}
			\item Remove punctuation, dates, etc.
			\item To lowercase
			\item Boilerplate removal
			\item Spellchecking
			\item Lemmatization \\ (city \& cities = city)
		\end{itemize}
		\column{0.33\textwidth}
		%Analysis
		\vspace{5mm}
		\begin{itemize}
			%\item Hierarchical clustering
			\item Fightin' Words
			\item Structural topic model
			%\begin{itemize}
			%	\item LDA
			%	\item STM
			%\end{itemize}
		\end{itemize}
	\end{columns}
\end{frame}

%\begin{frame}{Other preprocessing}
%\begin{itemize}
%	\item Remove:
%	\begin{itemize}
%		\item Dates
%		\item Numbers
%		\item Punctuation
%		\item Words not recognized by an English dictionary
%		\item Stopwords
%		\item Documents with too few unique words
%	\end{itemize}
%\item Set to lowercase
%\item Lemmatization
%\end{itemize}
%\end{frame}

\section{Analysis}
%\begin{frame}{Hierarchical Clustering}
%\centering
%\includegraphics[height=\textheight]{../figures/heatmap_hclust_40.png}
%\end{frame}

%\begin{frame}{Fightin' Words}
%\begin{figure}
%	\hfill
%	\subfigure{\includegraphics[width=5cm]{../figures/fightinWordsIN.png}}
%	\hfill
%	\subfigure{\includegraphics[width=5cm]{../figures/fightinWordsLA.png}}
%\end{figure}
%\end{frame}

\begin{frame}{Fightin' Words}
	\input{fightinwordsIN.tex} %../tables/stmTopWordsINDem.tex
\end{frame}

%\begin{frame}{Latent Dirichlet Allocation}
%\input{partisanTopWords.tex}
%\end{frame}

\begin{frame}{Structural Topic Model}
\input{stmTopWordsINDem.tex} %../tables/stmTopWordsINDem.tex
\end{frame}

\begin{frame}{Structural Topic Model}
\input{stmTopWordsINRep.tex} %../tables/stmTopWordsINRep.tex
\end{frame}

\section{Conclusion}
\begin{frame}{Conclusion}
\linespread{1.5}
\large
\begin{itemize}
	%\item Hierarchical clustering only creates 'topics' if content is VERY similar
	%\item Fightin words is extremely effective at forcing the text into two categories, but also somewhat limited
	%\item LDA is sensitive to preprocessing, but the most flexible
	%\item STM works too, but also very sensitive to duplicate content
	\item Manual vs. automatic analysis of government websites
	\item Our pipeline facilities scalable comparative analysis

	\vspace{5mm}
	\item Partisanship affects website content:
	\item Democrats focus on raising and spending money
	\item Republicans focus on infrastructure and utilities
\end{itemize}
\end{frame}

\end{document}