### 2017-12-19 paper/presentation/stmTopWordsINDem.tex

%\usepackage[table]{xcolor}
% latex table generated in R 3.4.2 by xtable 1.8-2 package
% Mon Nov 13 15:36:58 2017
\begin{table}[ht]
\centering
\begin{tabular}{llllll}
  \hline
-0.026 & -0.019 & -0.018 & -0.018 & -0.018 & -0.013 \\ 
  \hline
fort & propos & said & prosecutor & digest & \cellcolor{blue!25} fund \\ 
  citi & author & ask & charg & introduc & \cellcolor{blue!25} grant \\ 
  ordin & district & state & feloni & author & \cellcolor{blue!25} budget \\ 
  approv & street & will & counti & counti & counti \\ 
  \cellcolor{blue!25} purchas & public & chair & case & appoint & state \\ 
  depart & control & propos & crime & board & feder \\ 
  properti & amend & year & crimin & approv & depart \\ 
  will & intersect & move & offic & district & appropri \\ 
  resolut & counti & need & victim & \cellcolor{blue!25} fund & increas  \\ 
  contract & committe & council & sentenc & street & agenc \\ 
   \hline
\end{tabular}
%\caption{Top Democratic topics and words (Indiana), according to STM. 
%The words are the top words for the most Democratic/Republican topic, determined
%by the size (and significance) of the coefficient (see table header) of the party covariate.} 
\label{tabSTMINDem}
\end{table}