### 2017-12-19 paper/presentation/stmTopWordsINRep.tex

% latex table generated in R 3.4.2 by xtable 1.8-2 package
% Mon Nov 13 15:36:58 2017
\begin{table}[ht]
\centering
\begin{tabular}{llllll}
  \hline
0.024 & 0.019 & 0.019 & 0.016 & 0.016 & 0.016 \\ 
  \hline
motion & plan & inc & request & council & \cellcolor{red!25} traffic \\ 
  second & \cellcolor{red!25} zone & electr & board & citi & amp \\ 
  made & applic & build & member & ordin & vehicl \\ 
  approv & \cellcolor{red!25} properti & construct & servic & common & stop \\ 
  mayor & approv & home & \cellcolor{red!25} street & councilor & sign \\ 
  present & sign & street & approv & amend & \cellcolor{red!25} road \\ 
  state & \cellcolor{red!25} site & meridian & purchas & resolut & block \\ 
  will & locat & servic & citi & adopt & signal \\ 
  citi & commiss & west & move & wherea & \cellcolor{red!25} street \\ 
  council & file & main & good & approv & driver \\ 
   \hline
\end{tabular}
%\caption{Top Republican topics and words (Indiana), according to STM. 
%The words are the top words for the most Democratic/Republican topic, determined
%by the size (and significance) of the coefficient (see table header) of the party covariate.} 
\label{tabSTMINRep}
\end{table}