### 2017-12-25 paper/tables/stmTopWordsLADem.tex
LA is finally working completely, results are much better

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Mon Dec 25 01:20:56 2017
\begin{table}[ht]
\centering
\begin{tabular}{llllllll}
  \hline
-0.044 & -0.04 & -0.021 & -0.019 & -0.016 & -0.012 & -0.011 & -0.01 \\ 
  \hline
owner & provide & victim & application & employee & city & revenue & community \\ 
  applicant & otherwise & crime & motion & policy & ordinance & fund & neighborhood \\ 
  new & respect & violence & second & city & whereas & tax & district \\ 
  application & thereto & arrest & street & report & bond & budget & police \\ 
  per & city & domestic & favor & department & resolution & million & public \\ 
  exist & authorize & officer & retention & audit & code & total & meeting \\ 
  build & ordinance & suspect & john & review & provide & year & engagement \\ 
  proposal & amend & report & arc & management & shall & general & crime \\ 
  receive & district & child & window & procedure & provision & service & officer \\ 
  construction & locate & police & heather & work & fund & forecast & department \\ 
   \hline
\end{tabular}
\caption{Top Democratic topics and words (Louisiana), according to STM. 
The words are the top words for the most Democratic/Republican topic, determined
by the size (and significance) of the coefficient (see table header) of the party covariate.} 
\label{tabSTMLADem}
\end{table}