### 2017-12-25 paper/tables/stmTopWordsLARep.tex

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Mon Dec 25 01:20:56 2017
\begin{table}[ht]
\centering
\begin{tabular}{llllllll}
  \hline
0.08 & 0.069 & 0.065 & 0.038 & 0.03 & 0.015 & 0.012 & 0.009 \\ 
  \hline
emergency & call & ordinance & say & say & city & home & school \\ 
  city & art & bid & city & councilman & mayor & church & program \\ 
  resident & event & city & new & mayor & council & house & youth \\ 
  weather & sponsor & summary & make & city & member & building & student \\ 
  storm & cost & authorize & work & council & board & build & recreation \\ 
  area & free & resolution & can & ordinance & commission & state & community \\ 
  official & music & amount & year & budget & director & old & child \\ 
  hurricane & show & department & take & meeting & office & first & high \\ 
  work & information & public & get & ask & district & store & city \\ 
  mayor & feature & contract & one & follow & public & one & center \\ 
   \hline
\end{tabular}
\caption{Top Republican topics and words (Louisiana), according to STM. 
The words are the top words for the most Democratic/Republican topic, determined
by the size (and significance) of the coefficient (see table header) of the party covariate.} 
\label{tabSTMLARep}
\end{table}