### 2017-12-26 paper/tables/stmTopWordsLADem.tex
results for LA (with population in the stm)

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Tue Dec 26 15:50:18 2017
\begin{table}[ht]
\centering
\begin{tabular}{llllllll}
  \hline
-0.136 & -0.102 & -0.043 & -0.02 & -0.02 & -0.012 & -0.012 & -0.012 \\ 
  \hline
art & otherwise & whereas & water & street & shall & police & event \\ 
  call & provide & city & main & inc & city & crime & city \\ 
  cost & respect & ordinance & sewer & drive & agreement & officer & park \\ 
  home & city & bond & project & construction & party & suspect & rental \\ 
  sponsor & thereto & provide & infrastructure & permit & provide & arrest & use \\ 
  church & authorize & resolution & street & service & property & report & hour \\ 
  amp & ordinance & code & system & avenue & owner & victim & hotel \\ 
  free & district & chapter & improvement & oak & provision & information & public \\ 
  museum & amend & shall & remark & park & section & murder & provide \\ 
  artist & locate & otherwise & phase & lane & agree & block & term \\ 
   \hline
\end{tabular}
\caption{Top Democratic topics and words (Louisiana), according to STM. 
The words are the top words for the most Democratic/Republican topic, determined
by the size (and significance) of the coefficient (see table header) of the party covariate.} 
\label{tabSTMLADem}
\end{table}