### 2018-02-13 paper/tables/groundtruth_bootstrapped.tex
A LOT of changes, additions and fixes to the whole city and mayor scraping, processing, analyzing, etc.

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Tue Feb 13 10:36:40 2018
\begin{table}[ht]
\centering
\begin{tabular}{rllll}
  \hline
 & dem.groundtruth & rep.groundtruth & dem.cities & rep.cities \\ 
  \hline
dem.groundtruth & 1, 1 & 0.807, 0.896 & 0.714, 0.729 & 0.68, 0.698 \\ 
  rep.groundtruth & 0.807, 0.896 & 1, 1 & 0.647, 0.697 & 0.641, 0.693 \\ 
  dem.cities & 0.714, 0.729 & 0.647, 0.697 & 1, 1 & 0.937, 0.944 \\ 
  rep.cities & 0.68, 0.698 & 0.641, 0.693 & 0.937, 0.944 & 1, 1 \\ 
   \hline
\end{tabular}
\caption{Ground truth test, comparing campaign websites of mayors of the 100 largest cities in the US and cities in Indiana and Louisiana. The values are bootstrapped confidence bounds for cosine similarities between concatenated document collections.} 
\label{groundtruth_bootstrapped}
\end{table}