### 2018-04-23 paper/tables/stmTopWords.tex
new results table for stm run on new data

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Mon Apr 23 14:12:56 2018
\begin{table}[ht]
\centering
\begingroup\footnotesize
\begin{tabular}{lllllll}
  \hline
1 & 2 & 3 & 4 & 5 & 6 & 7 \\ 
  \hline
\cellcolor{red!10}yon & \cellcolor{red!10}borough & \cellcolor{red!10}port & \cellcolor{red!10}waterfront & \cellcolor{red!10}queens & \cellcolor{red!10}boroughs & \cellcolor{red!10}island \\ 
  \cellcolor{blue!10}noise & \cellcolor{blue!10}impacts & \cellcolor{blue!10}mitigation & \cellcolor{blue!10}vibration & \cellcolor{blue!10}ambient & \cellcolor{blue!10}adverse & \cellcolor{blue!10}thresholds \\ 
  \cellcolor{red!10}tax & \cellcolor{red!10}exemption & \cellcolor{red!10}taxes & \cellcolor{red!10}estate & \cellcolor{red!10}assessed & \cellcolor{red!10}real & \cellcolor{red!10}taxpayer \\ 
  \cellcolor{red!10}para & \cellcolor{red!10}personas & \cellcolor{red!10}antes & \cellcolor{red!10}persona & \cellcolor{red!10}horas & \cellcolor{red!10}junta & \cellcolor{red!10}con \\ 
  \cellcolor{blue!10}election & \cellcolor{blue!10}ethics & \cellcolor{blue!10}appointed & \cellcolor{blue!10}elections & \cellcolor{blue!10}ballot & \cellcolor{blue!10}charter & \cellcolor{blue!10}elected \\ 
  \cellcolor{red!10}wetland & \cellcolor{red!10}habitats & \cellcolor{red!10}riparian & \cellcolor{red!10}habitat & \cellcolor{red!10}wetlands & \cellcolor{red!10}tidal & \cellcolor{red!10}freshwater \\ 
  \cellcolor{blue!10}parking & \cellcolor{blue!10}bus & \cellcolor{blue!10}transit & \cellcolor{blue!10}mall & \cellcolor{blue!10}buses & \cellcolor{blue!10}campus & \cellcolor{blue!10}arena \\ 
  \cellcolor{red!30}click & \cellcolor{red!30}download & \cellcolor{red!30}online & \cellcolor{red!30}please & \cellcolor{red!30}email & \cellcolor{red!30}website & \cellcolor{red!30}visit \\ 
  \cellcolor{blue!10}draft & \cellcolor{blue!10}comments & \cellcolor{blue!10}comment & \cellcolor{blue!10}meetings & \cellcolor{blue!10}update & \cellcolor{blue!10}presentation & \cellcolor{blue!10}briefing \\ 
  \cellcolor{blue!10}ave & \cellcolor{blue!10}blvd & \cellcolor{blue!10}glen & \cellcolor{blue!10}pkwy & \cellcolor{blue!10}hwy & \cellcolor{blue!10}cove & \cellcolor{blue!10}fwy \\ 
  \cellcolor{blue!20}neighborhoods & \cellcolor{blue!20}strategy & \cellcolor{blue!20}vision & \cellcolor{blue!20}strategies & \cellcolor{blue!20}businesses & \cellcolor{blue!20}opportunities & \cellcolor{blue!20}vibrant \\ 
  \cellcolor{white}bid & \cellcolor{white}contract & \cellcolor{white}invoices & \cellcolor{white}procurement & \cellcolor{white}purchasing & \cellcolor{white}bids & \cellcolor{white}vendor \\ 
  \cellcolor{white}marijuana & \cellcolor{white}cannabis & \cellcolor{white}licensee & \cellcolor{white}taxicab & \cellcolor{white}license & \cellcolor{white}mischief & \cellcolor{white}citation \\ 
  \cellcolor{blue!10}complaint & \cellcolor{blue!10}discrimination & \cellcolor{blue!10}defendants & \cellcolor{blue!10}bankruptcy & \cellcolor{blue!10}trial & \cellcolor{blue!10}harassment & \cellcolor{blue!10}defendant \\ 
  \cellcolor{red!10}rouge & \cellcolor{red!10}baton & \cellcolor{red!10}foods & \cellcolor{red!10}parish & \cellcolor{red!10}vegetables & \cellcolor{red!10}vending & \cellcolor{red!10}cooked \\ 
  \cellcolor{red!10}child & \cellcolor{red!10}violence & \cellcolor{red!10}abuse & \cellcolor{red!10}mental & \cellcolor{red!10}clients & \cellcolor{red!10}inmates & \cellcolor{red!10}homelessness \\ 
  \cellcolor{red!10}applicants & \cellcolor{red!10}landlord & \cellcolor{red!10}tenant & \cellcolor{red!10}rent & \cellcolor{red!10}exam & \cellcolor{red!10}tenants & \cellcolor{red!10}applications \\ 
  \cellcolor{red!10}think & \cellcolor{red!10}say & \cellcolor{red!10}really & \cellcolor{red!10}thing & \cellcolor{red!10}okay & \cellcolor{red!10}got & \cellcolor{red!10}maybe \\ 
  \cellcolor{blue!20}setback & \cellcolor{blue!20}fence & \cellcolor{blue!20}yard & \cellcolor{blue!20}zoned & \cellcolor{blue!20}front & \cellcolor{blue!20}height & \cellcolor{blue!20}accessory \\ 
  \cellcolor{red!10}shall & \cellcolor{red!10}subsection & \cellcolor{red!10}article & \cellcolor{red!10}provisions & \cellcolor{red!10}pursuant & \cellcolor{red!10}thereof & \cellcolor{red!10}forth \\ 
  \cellcolor{red!20}explained & \cellcolor{red!20}asked & \cellcolor{red!20}said & \cellcolor{red!20}legislator & \cellcolor{red!20}commented & \cellcolor{red!20}advised & \cellcolor{red!20}leg \\ 
  \cellcolor{blue!10}respondents & \cellcolor{blue!10}census & \cellcolor{blue!10}population & \cellcolor{blue!10}compared & \cellcolor{blue!10}average & \cellcolor{blue!10}trends & \cellcolor{blue!10}comparison \\ 
  \cellcolor{red!20}infection & \cellcolor{red!20}symptoms & \cellcolor{red!20}breastfeeding & \cellcolor{red!20}syphilis & \cellcolor{red!20}doses & \cellcolor{red!20}asthma & \cellcolor{red!20}tuberculosis \\ 
  \cellcolor{red!20}yes & \cellcolor{red!20}name & \cellcolor{red!20}signature & \cellcolor{red!20}mailing & \cellcolor{red!20}zip & \cellcolor{red!20}attach & \cellcolor{red!20}form \\ 
  \cellcolor{red!30}games & \cellcolor{red!30}tournament & \cellcolor{red!30}swim & \cellcolor{red!30}players & \cellcolor{red!30}player & \cellcolor{red!30}camp & \cellcolor{red!30}game \\ 
  \cellcolor{blue!10}goals & \cellcolor{blue!10}implementation & \cellcolor{blue!10}policies & \cellcolor{blue!10}policy & \cellcolor{blue!10}specific & \cellcolor{blue!10}implement & \cellcolor{blue!10}comprehensive \\ 
  \cellcolor{red!20}ems & \cellcolor{red!20}fires & \cellcolor{red!20}preparedness & \cellcolor{red!20}disaster & \cellcolor{red!20}evacuation & \cellcolor{red!20}fire & \cellcolor{red!20}firefighters \\ 
  \cellcolor{red!10}subcontractor & \cellcolor{red!10}subcontractors & \cellcolor{red!10}agrees & \cellcolor{red!10}proposer & \cellcolor{red!10}contractor & \cellcolor{red!10}grantee & \cellcolor{red!10}breach \\ 
  \cellcolor{blue!10}realm & \cellcolor{blue!10}massing & \cellcolor{blue!10}facades & \cellcolor{blue!10}entrances & \cellcolor{blue!10}plazas & \cellcolor{blue!10}elements & \cellcolor{blue!10}proponents \\ 
  \cellcolor{blue!10}employee & \cellcolor{blue!10}allegation & \cellcolor{blue!10}overtime & \cellcolor{blue!10}named & \cellcolor{blue!10}leave & \cellcolor{blue!10}grievance & \cellcolor{blue!10}wage \\ 
  \cellcolor{blue!10}parks & \cellcolor{blue!10}playground & \cellcolor{blue!10}beach & \cellcolor{blue!10}park & \cellcolor{blue!10}picnic & \cellcolor{blue!10}marina & \cellcolor{blue!10}trails \\ 
  \cellcolor{blue!10}lanes & \cellcolor{blue!10}bicycle & \cellcolor{blue!10}bike & \cellcolor{blue!10}intersections & \cellcolor{blue!10}bicyclists & \cellcolor{blue!10}roadway & \cellcolor{blue!10}pedestrians \\ 
  \cellcolor{blue!20}absent & \cellcolor{blue!20}aye & \cellcolor{blue!20}khan & \cellcolor{blue!20}voting & \cellcolor{blue!20}berry & \cellcolor{blue!20}nays & \cellcolor{blue!20}tagged \\ 
  \cellcolor{blue!20}budget & \cellcolor{blue!20}revenue & \cellcolor{blue!20}million & \cellcolor{blue!20}revenues & \cellcolor{blue!20}budgeted & \cellcolor{blue!20}fund & \cellcolor{blue!20}expenditures \\ 
  \cellcolor{blue!10}whereas & \cellcolor{blue!10}resolution & \cellcolor{blue!10}amending & \cellcolor{blue!10}resolved & \cellcolor{blue!10}hereby & \cellcolor{blue!10}authorizes & \cellcolor{blue!10}digest \\ 
  \cellcolor{blue!10}assets & \cellcolor{blue!10}statements & \cellcolor{blue!10}governmental & \cellcolor{blue!10}accounting & \cellcolor{blue!10}liabilities & \cellcolor{blue!10}net & \cellcolor{blue!10}financial \\ 
  \cellcolor{red!30}honored & \cellcolor{red!30}joined & \cellcolor{red!30}proud & \cellcolor{red!30}fort & \cellcolor{red!30}announces & \cellcolor{red!30}won & \cellcolor{red!30}worth \\ 
  \cellcolor{white}analyst & \cellcolor{white}technician & \cellcolor{white}specialist & \cellcolor{white}performs & \cellcolor{white}prepares & \cellcolor{white}coordinates & \cellcolor{white}assists \\ 
  \cellcolor{white}server & \cellcolor{white}wireless & \cellcolor{white}software & \cellcolor{white}aircraft & \cellcolor{white}servers & \cellcolor{white}airport & \cellcolor{white}technology \\ 
  \cellcolor{blue!10}improvements & \cellcolor{blue!10}project & \cellcolor{blue!10}projects & \cellcolor{blue!10}phase & \cellcolor{blue!10}replacement & \cellcolor{blue!10}reconstruction & \cellcolor{blue!10}upgrades \\ 
  \cellcolor{red!10}recycling & \cellcolor{red!10}recycle & \cellcolor{red!10}garbage & \cellcolor{red!10}bags & \cellcolor{red!10}waste & \cellcolor{red!10}recyclable & \cellcolor{red!10}recyclables \\ 
  \cellcolor{blue!10}housing & \cellcolor{blue!10}affordable & \cellcolor{blue!10}households & \cellcolor{blue!10}affordability & \cellcolor{blue!10}income & \cellcolor{blue!10}moderate & \cellcolor{blue!10}homeless \\ 
  \cellcolor{red!20}effluent & \cellcolor{red!20}wastewater & \cellcolor{red!20}discharges & \cellcolor{red!20}contaminants & \cellcolor{red!20}sludge & \cellcolor{red!20}infiltration & \cellcolor{red!20}solids \\ 
  \cellcolor{red!10}fee & \cellcolor{red!10}permit & \cellcolor{red!10}inspection & \cellcolor{red!10}permits & \cellcolor{red!10}inspections & \cellcolor{red!10}fees & \cellcolor{red!10}occupancy \\ 
  \cellcolor{blue!10}uses & \cellcolor{blue!10}mixed & \cellcolor{blue!10}density & \cellcolor{blue!10}land & \cellcolor{blue!10}industrial & \cellcolor{blue!10}residential & \cellcolor{blue!10}commercial \\ 
  \cellcolor{red!10}energy & \cellcolor{red!10}renewable & \cellcolor{red!10}coal & \cellcolor{red!10}electricity & \cellcolor{red!10}climate & \cellcolor{red!10}solar & \cellcolor{red!10}greenhouse \\ 
  \cellcolor{blue!10}bonds & \cellcolor{blue!10}refunding & \cellcolor{blue!10}securities & \cellcolor{blue!10}bond & \cellcolor{blue!10}debt & \cellcolor{blue!10}issuer & \cellcolor{blue!10}maturity \\ 
  \cellcolor{red!20}artists & \cellcolor{red!20}artist & \cellcolor{red!20}performances & \cellcolor{red!20}music & \cellcolor{red!20}orchestra & \cellcolor{red!20}symphony & \cellcolor{red!20}arts \\ 
  \cellcolor{blue!80}arrested & \cellcolor{blue!80}suspect & \cellcolor{blue!80}homicide & \cellcolor{blue!80}suspects & \cellcolor{blue!80}shooting & \cellcolor{blue!80}sergeant & \cellcolor{blue!80}arrests \\ 
  \cellcolor{white}retiree & \cellcolor{white}actuarial & \cellcolor{white}retirement & \cellcolor{white}deductible & \cellcolor{white}retirees & \cellcolor{white}copay & \cellcolor{white}dental \\ 
  \cellcolor{red!10}ferrets & \cellcolor{red!10}dogs & \cellcolor{red!10}cats & \cellcolor{red!10}rabies & \cellcolor{red!10}pets & \cellcolor{red!10}spay & \cellcolor{red!10}stray \\ 
  \cellcolor{blue!10}audit & \cellcolor{blue!10}auditor & \cellcolor{blue!10}audits & \cellcolor{blue!10}procedures & \cellcolor{blue!10}controls & \cellcolor{blue!10}implemented & \cellcolor{blue!10}accountability \\ 
  \cellcolor{blue!10}avenue & \cellcolor{blue!10}west & \cellcolor{blue!10}east & \cellcolor{blue!10}south & \cellcolor{blue!10}north & \cellcolor{blue!10}thence & \cellcolor{blue!10}street \\ 
  \cellcolor{white}historic & \cellcolor{white}revival & \cellcolor{white}landmarks & \cellcolor{white}landmark & \cellcolor{white}craftsman & \cellcolor{white}bungalow & \cellcolor{white}style \\ 
  \cellcolor{red!10}remodel & \cellcolor{red!10}monoxide & \cellcolor{red!10}alarms & \cellcolor{red!10}roofing & \cellcolor{red!10}heater & \cellcolor{red!10}description & \cellcolor{red!10}bathroom \\ 
  \cellcolor{red!10}motion & \cellcolor{red!10}alderman & \cellcolor{red!10}seconded & \cellcolor{red!10}carried & \cellcolor{red!10}whiting & \cellcolor{red!10}unanimously & \cellcolor{red!10}ayes \\ 
  \cellcolor{white}pruning & \cellcolor{white}tree & \cellcolor{white}forestry & \cellcolor{white}trees & \cellcolor{white}mulch & \cellcolor{white}planting & \cellcolor{white}planted \\ 
  \cellcolor{red!10}fittings & \cellcolor{red!10}joints & \cellcolor{red!10}thickness & \cellcolor{red!10}pipe & \cellcolor{red!10}trench & \cellcolor{red!10}valve & \cellcolor{red!10}psi \\ 
  \cellcolor{white}students & \cellcolor{white}learning & \cellcolor{white}schools & \cellcolor{white}student & \cellcolor{white}academic & \cellcolor{white}career & \cellcolor{white}education \\ 
  \cellcolor{red!20}plat & \cellcolor{red!20}sign & \cellcolor{red!20}pud & \cellcolor{red!20}petitioner & \cellcolor{red!20}petition & \cellcolor{red!20}variance & \cellcolor{red!20}subdivision \\ 
   \hline
\end{tabular}
\endgroup
\end{table}