### 2018-05-23 paper/tables/boilerplateClassifierRFMetrics.tex
added random forest with inverse probability weighting

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Wed May 23 15:57:06 2018
\begin{table}[ht]
\centering
\begin{tabular}{rr}
  \hline
 & Value \\ 
  \hline
Percent Correctly Predicted & 0.88 \\ 
  Precision & 0.88 \\ 
  Recall & 0.91 \\ 
  F1-Score & 0.89 \\ 
   \hline
\end{tabular}
\caption{Performance metrics for random forest boilerplate classifier, with IPW weighting.} 
\end{table}