### 2018-05-23 paper/tables/boilerplateClassifierRFMetrics.tex
comparison between weighting schemes for RF. Differences are fairly small.

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Wed May 23 16:31:52 2018
\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
 & IPW & Prob & No Weight \\ 
  \hline
Percent Correctly Predicted & 0.88 & 0.89 & 0.89 \\ 
  Precision & 0.88 & 0.89 & 0.89 \\ 
  Recall & 0.91 & 0.92 & 0.91 \\ 
  F1-Score & 0.90 & 0.90 & 0.90 \\ 
   \hline
\end{tabular}
\caption{Performance metrics for random forest boilerplate classifier, with either Inverse Probability Weights (IPW), probabilty (of being in the training sample) (prob), or no weighting.} 
\end{table}