### 2018-06-14 paper/tables/stmTopWords2.tex
changes to the results table; last sentence of the abstract

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Thu Jun 14 14:34:01 2018
\begin{table}[ht]
\centering
\begingroup\scriptsize
\begin{tabular}{rlllllllr}
  \hline
Topic & Top Word 1 & Top Word 2 & Top Word 3 & Top Word 4 & Top Word 5 & Top Word 6 & Top Word 7 & Tokens assigned \\ 
  \hline
 38 & \cellcolor{red!40}artist & \cellcolor{red!40}poetry & \cellcolor{red!40}music & \cellcolor{red!40}fun & \cellcolor{red!40}dance & \cellcolor{red!40}exhibition & \cellcolor{red!40}player & 3770 \\ 
    5 & \cellcolor{red!30}please & \cellcolor{red!30}email & \cellcolor{red!30}mail & \cellcolor{red!30}copy & \cellcolor{red!30}contact & \cellcolor{red!30}click & \cellcolor{red!30}form & 260 \\ 
   43 & \cellcolor{red!20}epidemiology & \cellcolor{red!20}infection & \cellcolor{red!20}vaccine & \cellcolor{red!20}antibody & \cellcolor{red!20}asthma & \cellcolor{red!20}hygiene & \cellcolor{red!20}symptom & 2469 \\ 
   20 & \cellcolor{red!20}snow & \cellcolor{red!20}hurricane & \cellcolor{red!20}tornado & \cellcolor{red!20}plow & \cellcolor{red!20}evacuate & \cellcolor{red!20}pothole & \cellcolor{red!20}spray & 1290 \\ 
   52 & \cellcolor{red!20}reappoints & \cellcolor{red!20}legislator & \cellcolor{red!20}cat & \cellcolor{red!20}leg & \cellcolor{red!20}sander & \cellcolor{red!20}dog & \cellcolor{red!20}spay & 1152 \\ 
   51 & \cellcolor{red!20}drinking & \cellcolor{red!20}wastewater & \cellcolor{red!20}water & \cellcolor{red!20}pump & \cellcolor{red!20}sludge & \cellcolor{red!20}sewage & \cellcolor{red!20}disinfection & 487 \\ 
   32 & \cellcolor{red!20}think & \cellcolor{red!20}really & \cellcolor{red!20}okay & \cellcolor{red!20}thing & \cellcolor{red!20}something & \cellcolor{red!20}seem & \cellcolor{red!20}nice & 1940 \\ 
   36 & \cellcolor{red!10}shall & \cellcolor{red!10}herein & \cellcolor{red!10}forth & \cellcolor{red!10}deem & \cellcolor{red!10}thereof & \cellcolor{red!10}hereunder & \cellcolor{red!10}construe & 433 \\ 
   27 & \cellcolor{red!10}library & \cellcolor{red!10}branch & \cellcolor{red!10}learn & \cellcolor{red!10}book & \cellcolor{red!10}online & \cellcolor{red!10}view & \cellcolor{red!10}upcoming & 302 \\ 
   58 & \cellcolor{red!10}buffalo & \cellcolor{red!10}announce & \cellcolor{red!10}warren & \cellcolor{red!10}lovely & \cellcolor{red!10}honor & \cellcolor{red!10}ceremony & \cellcolor{red!10}news & 1298 \\ 
   33 & \cellcolor{red!10}fire & \cellcolor{red!10}fort & \cellcolor{red!10}worth & \cellcolor{red!10}beach & \cellcolor{red!10}alarm & \cellcolor{red!10}firefighter & \cellcolor{red!10}rescue & 459 \\ 
   34 & \cellcolor{red!10}fee & \cellcolor{red!10}charge & \cellcolor{red!10}per & \cellcolor{red!10}billing & \cellcolor{red!10}bill & \cellcolor{red!10}refund & \cellcolor{red!10}monthly & 241 \\ 
   35 & \cellcolor{red!10}youth & \cellcolor{red!10}student & \cellcolor{red!10}parent & \cellcolor{red!10}school & \cellcolor{red!10}teacher & \cellcolor{red!10}academic & \cellcolor{red!10}classroom & 710 \\ 
   56 & \cellcolor{red!10}garland & \cellcolor{red!10}auburn & \cellcolor{red!10}councilor & \cellcolor{red!10}plain & \cellcolor{red!10}hall & \cellcolor{red!10}ward & \cellcolor{red!10}city & 229 \\ 
   21 & \cellcolor{red!10}bid & \cellcolor{red!10}proposer & \cellcolor{red!10}subcontractor & \cellcolor{red!10}bidder & \cellcolor{red!10}contractor & \cellcolor{red!10}subcontract & \cellcolor{red!10}solicitation & 485 \\ 
   59 & \cellcolor{red!10}motion & \cellcolor{red!10}adjourn & \cellcolor{red!10}unanimously & \cellcolor{red!10}second & \cellcolor{red!10}ayes & \cellcolor{red!10}carry & \cellcolor{red!10}prince & 487 \\ 
   49 & \cellcolor{red!10}garbage & \cellcolor{red!10}recycling & \cellcolor{red!10}bin & \cellcolor{red!10}recyclable & \cellcolor{red!10}recyclables & \cellcolor{red!10}cart & \cellcolor{red!10}bag & 1635 \\ 
   31 & \cellcolor{red!10}deductible & \cellcolor{red!10}dental & \cellcolor{red!10}medicare & \cellcolor{red!10}coinsurance & \cellcolor{red!10}copay & \cellcolor{red!10}aircraft & \cellcolor{red!10}dentist & 706 \\ 
   54 & \cellcolor{red!10}duct & \cellcolor{red!10}conduit & \cellcolor{red!10}bolt & \cellcolor{red!10}splice & \cellcolor{red!10}valve & \cellcolor{red!10}piping & \cellcolor{red!10}fitting & 1477 \\ 
   14 & \cellcolor{red!10}immigrant & \cellcolor{red!10}discrimination & \cellcolor{red!10}gender & \cellcolor{red!10}immigration & \cellcolor{red!10}racial & \cellcolor{red!10}refugee & \cellcolor{red!10}transgender & 1095 \\ 
    1 & \cellcolor{red!10}storm & \cellcolor{red!10}runoff & \cellcolor{red!10}drainage & \cellcolor{red!10}infiltration & \cellcolor{red!10}drain & \cellcolor{red!10}discharge & \cellcolor{red!10}sediment & 490 \\ 
    2 & \cellcolor{red!10}yon & \cellcolor{red!10}ave & \cellcolor{red!10}blvd & \cellcolor{red!10}greenwood & \cellcolor{red!10}suite & \cellcolor{red!10}comm & \cellcolor{red!10}bldg & 1317 \\ 
    4 & \cellcolor{red!10}para & \cellcolor{red!10}persona & \cellcolor{red!10}ante & \cellcolor{red!10}horas & \cellcolor{red!10}junta & \cellcolor{red!10}largo & \cellcolor{red!10}sin & 1377 \\ 
   55 & \cellcolor{red!10}alderman & \cellcolor{red!10}whereas & \cellcolor{red!10}hereby & \cellcolor{red!10}ordain & \cellcolor{red!10}resolution & \cellcolor{red!10}resolve & \cellcolor{red!10}amend & 457 \\ 
   26 & \cellcolor{red!10}sampling & \cellcolor{red!10}petroleum & \cellcolor{red!10}sample & \cellcolor{red!10}concentration & \cellcolor{red!10}hydrocarbon & \cellcolor{red!10}pesticide & \cellcolor{red!10}coal & 1278 \\ 
   48 & \cellcolor{red!10}premise & \cellcolor{red!10}marijuana & \cellcolor{red!10}permit & \cellcolor{red!10}licensee & \cellcolor{red!10}license & \cellcolor{red!10}cannabis & \cellcolor{red!10}violation & 489 \\ 
   12 & \cellcolor{red!10}server & \cellcolor{red!10}wireless & \cellcolor{red!10}software & \cellcolor{red!10}digital & \cellcolor{red!10}telecommunication & \cellcolor{red!10}technology & \cellcolor{red!10}desktop & 917 \\ 
    7 & \cellcolor{red!10}energy & \cellcolor{red!10}renewable & \cellcolor{red!10}solar & \cellcolor{red!10}climate & \cellcolor{red!10}electricity & \cellcolor{red!10}greenhouse & \cellcolor{red!10}efficiency & 740 \\ 
   16 & \cellcolor{white}recreation & \cellcolor{white}golf & \cellcolor{white}playground & \cellcolor{white}park & \cellcolor{white}picnic & \cellcolor{white}zoo & \cellcolor{white}tennis & 702 \\ 
   60 & \cellcolor{white}exhaust & \cellcolor{white}air & \cellcolor{white}boiler & \cellcolor{white}diesel & \cellcolor{white}ozone & \cellcolor{white}fuel & \cellcolor{white}header & 316 \\ 
    3 & \cellcolor{white}rouge & \cellcolor{white}baton & \cellcolor{white}issuer & \cellcolor{white}maturity & \cellcolor{white}parish & \cellcolor{white}jun & \cellcolor{white}portfolio & 502 \\ 
   23 & \cellcolor{white}economic & \cellcolor{white}attract & \cellcolor{white}downtown & \cellcolor{white}economy & \cellcolor{white}industry & \cellcolor{white}revitalization & \cellcolor{white}tourism & 862 \\ 
   25 & \cellcolor{white}incumbent & \cellcolor{white}exam & \cellcolor{white}supervise & \cellcolor{white}supervision & \cellcolor{white}knowledge & \cellcolor{white}examination & \cellcolor{white}ability & 683 \\ 
    8 & \cellcolor{white}actuarial & \cellcolor{white}retirement & \cellcolor{white}pension & \cellcolor{white}contribution & \cellcolor{white}retiree & \cellcolor{white}valuation & \cellcolor{white}unfunded & 289 \\ 
    9 & \cellcolor{white}facade & \cellcolor{white}awning & \cellcolor{white}roof & \cellcolor{white}porch & \cellcolor{white}balcony & \cellcolor{white}exterior & \cellcolor{white}wall & 1103 \\ 
   15 & \cellcolor{white}shoreline & \cellcolor{white}marsh & \cellcolor{white}coastal & \cellcolor{white}habitat & \cellcolor{white}wetland & \cellcolor{white}salmon & \cellcolor{white}floodplain & 1454 \\ 
   42 & \cellcolor{white}tax & \cellcolor{white}exemption & \cellcolor{white}taxable & \cellcolor{white}real & \cellcolor{white}abatement & \cellcolor{white}property & \cellcolor{white}appraisal & 343 \\ 
   30 & \cellcolor{blue!10}population & \cellcolor{blue!10}census & \cellcolor{blue!10}respondent & \cellcolor{blue!10}figure & \cellcolor{blue!10}trend & \cellcolor{blue!10}comparison & \cellcolor{blue!10}survey & 540 \\ 
   53 & \cellcolor{blue!10}historic & \cellcolor{blue!10}landmark & \cellcolor{blue!10}revival & \cellcolor{blue!10}century & \cellcolor{blue!10}historian & \cellcolor{blue!10}archaeological & \cellcolor{blue!10}artifact & 2518 \\ 
   11 & \cellcolor{blue!10}parking & \cellcolor{blue!10}vehicle & \cellcolor{blue!10}passenger & \cellcolor{blue!10}tow & \cellcolor{blue!10}garage & \cellcolor{blue!10}taxicab & \cellcolor{blue!10}taxi & 435 \\ 
   18 & \cellcolor{blue!10}prune & \cellcolor{blue!10}tree & \cellcolor{blue!10}forestry & \cellcolor{blue!10}shrub & \cellcolor{blue!10}deer & \cellcolor{blue!10}planting & \cellcolor{blue!10}deciduous & 2279 \\ 
   45 & \cellcolor{blue!10}variance & \cellcolor{blue!10}plat & \cellcolor{blue!10}setback & \cellcolor{blue!10}zoning & \cellcolor{blue!10}fence & \cellcolor{blue!10}yard & \cellcolor{blue!10}lot & 300 \\ 
   19 & \cellcolor{blue!10}noise & \cellcolor{blue!10}mitigation & \cellcolor{blue!10}impact & \cellcolor{blue!10}fugitive & \cellcolor{blue!10}adverse & \cellcolor{blue!10}significant & \cellcolor{blue!10}vibration & 360 \\ 
   13 & \cellcolor{blue!10}agency & \cellcolor{blue!10}yes & \cellcolor{blue!10}federal & \cellcolor{blue!10}entity & \cellcolor{blue!10}recipient & \cellcolor{blue!10}deficiency & \cellcolor{blue!10}compliance & 239 \\ 
    6 & \cellcolor{blue!10}improvement & \cellcolor{blue!10}project & \cellcolor{blue!10}upgrade & \cellcolor{blue!10}capital & \cellcolor{blue!10}appropriated & \cellcolor{blue!10}replacement & \cellcolor{blue!10}phase & 189 \\ 
   46 & \cellcolor{blue!10}employee & \cellcolor{blue!10}overtime & \cellcolor{blue!10}sick & \cellcolor{blue!10}bargaining & \cellcolor{blue!10}wage & \cellcolor{blue!10}salary & \cellcolor{blue!10}grievance & 398 \\ 
   28 & \cellcolor{blue!10}allegation & \cellcolor{blue!10}complainant & \cellcolor{blue!10}defendant & \cellcolor{blue!10}misconduct & \cellcolor{blue!10}allege & \cellcolor{blue!10}bankruptcy & \cellcolor{blue!10}plaintiff & 1747 \\ 
   40 & \cellcolor{blue!10}tab & \cellcolor{blue!10}mode & \cellcolor{blue!10}accessibility & \cellcolor{blue!10}false & \cellcolor{blue!10}focus & \cellcolor{blue!10}else & \cellcolor{blue!10}null & 257 \\ 
   10 & \cellcolor{blue!10}density & \cellcolor{blue!10}us & \cellcolor{blue!10}mixed & \cellcolor{blue!10}village & \cellcolor{blue!10}urban & \cellcolor{blue!10}orient & \cellcolor{blue!10}mix & 336 \\ 
   39 & \cellcolor{blue!10}comment & \cellcolor{blue!10}draft & \cellcolor{blue!10}review & \cellcolor{blue!10}preliminary & \cellcolor{blue!10}planning & \cellcolor{blue!10}propose & \cellcolor{blue!10}revision & 274 \\ 
   37 & \cellcolor{blue!10}audit & \cellcolor{blue!10}auditor & \cellcolor{blue!10}internal & \cellcolor{blue!10}procedure & \cellcolor{blue!10}implement & \cellcolor{blue!10}oversight & \cellcolor{blue!10}timely & 402 \\ 
   44 & \cellcolor{blue!10}housing & \cellcolor{blue!10}affordable & \cellcolor{blue!10}homeless & \cellcolor{blue!10}homelessness & \cellcolor{blue!10}landlord & \cellcolor{blue!10}affordability & \cellcolor{blue!10}rent & 340 \\ 
   17 & \cellcolor{blue!10}debt & \cellcolor{blue!10}governmental & \cellcolor{blue!10}bond & \cellcolor{blue!10}obligation & \cellcolor{blue!10}financial & \cellcolor{blue!10}accounting & \cellcolor{blue!10}statement & 259 \\ 
   41 & \cellcolor{blue!10}bicycle & \cellcolor{blue!10}bike & \cellcolor{blue!10}lane & \cellcolor{blue!10}intersection & \cellcolor{blue!10}pedestrian & \cellcolor{blue!10}crosswalk & \cellcolor{blue!10}crossing & 527 \\ 
   24 & \cellcolor{blue!10}strategy & \cellcolor{blue!10}goal & \cellcolor{blue!10}outreach & \cellcolor{blue!10}priority & \cellcolor{blue!10}strategic & \cellcolor{blue!10}stakeholder & \cellcolor{blue!10}engagement & 313 \\ 
   50 & \cellcolor{blue!20}aye & \cellcolor{blue!20}absent & \cellcolor{blue!20}nay & \cellcolor{blue!20}councilman & \cellcolor{blue!20}khan & \cellcolor{blue!20}voting & \cellcolor{blue!20}berry & 674 \\ 
   22 & \cellcolor{blue!20}budget & \cellcolor{blue!20}revenue & \cellcolor{blue!20}expenditure & \cellcolor{blue!20}million & \cellcolor{blue!20}appropriation & \cellcolor{blue!20}forecast & \cellcolor{blue!20}allocation & 236 \\ 
   47 & \cellcolor{blue!20}digest & \cellcolor{blue!20}authorize & \cellcolor{blue!20}inc & \cellcolor{blue!20}consolidated & \cellcolor{blue!20}contingency & \cellcolor{blue!20}agreement & \cellcolor{blue!20}reinvestment & 215 \\ 
   29 & \cellcolor{blue!20}chair & \cellcolor{blue!20}election & \cellcolor{blue!20}agenda & \cellcolor{blue!20}committee & \cellcolor{blue!20}speaker & \cellcolor{blue!20}ballot & \cellcolor{blue!20}commission & 353 \\ 
   57 & \cellcolor{blue!80}robbery & \cellcolor{blue!80}homicide & \cellcolor{blue!80}sergeant & \cellcolor{blue!80}arrest & \cellcolor{blue!80}suspect & \cellcolor{blue!80}crime & \cellcolor{blue!80}burglary & 1255 \\ 
   \hline
\end{tabular}
\endgroup
\caption{Top words from a structural topic model with 60 topics and FREX scoring. Colors depict partisanship based on coefficient size. White cells are non-significant topics. Based on data preprocessed with the classifier.} 
\end{table}