### 2018-06-26 paper/tables/stmTopWords120.tex
stm with 120 topics, similar findings as the one with 60

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Jun 26 15:42:11 2018
\begin{table}[ht]
\centering
\begingroup\scriptsize
\begin{tabular}{rllllllll}
  \hline
 \# & Top Word 1 & Top Word 2 & Top Word 3 & Top Word 4 & Top Word 5 & Top Word 6 & \multicolumn{2}{c}{Tokens assigned} \\ 
  \hline
 93 & \cellcolor{red!30}kindness & \cellcolor{red!30}winner & \cellcolor{red!30}hero & \cellcolor{red!30}famous & \cellcolor{red!30}tribute & \cellcolor{red!30}wager & \mybar{3042} \\ 
   36 & \cellcolor{red!20}copy & \cellcolor{red!20}record & \cellcolor{red!20}request & \cellcolor{red!20}mail & \cellcolor{red!20}submit & \cellcolor{red!20}fax & \mybar{111} \\ 
   98 & \cellcolor{red!20}community & \cellcolor{red!20}resident & \cellcolor{red!20}mission & \cellcolor{red!20}quality & \cellcolor{red!20}excellent & \cellcolor{red!20}life & \mybar{78} \\ 
   52 & \cellcolor{red!20}county & \cellcolor{red!20}leg & \cellcolor{red!20}legislator & \cellcolor{red!20}legislature & \cellcolor{red!20}town & \cellcolor{red!20}municipality & \mybar{132} \\ 
   18 & \cellcolor{red!20}often & \cellcolor{red!20}always & \cellcolor{red!20}sometimes & \cellcolor{red!20}never & \cellcolor{red!20}easy & \cellcolor{red!20}even & \mybar{505} \\ 
   20 & \cellcolor{red!20}click & \cellcolor{red!20}blog & \cellcolor{red!20}email & \cellcolor{red!20}copyright & \cellcolor{red!20}dream & \cellcolor{red!20}sorry & \mybar{336} \\ 
   38 & \cellcolor{red!10}camp & \cellcolor{red!10}yoga & \cellcolor{red!10}library & \cellcolor{red!10}camper & \cellcolor{red!10}fun & \cellcolor{red!10}librarian & \mybar{1009} \\ 
   43 & \cellcolor{red!10}antibody & \cellcolor{red!10}infection & \cellcolor{red!10}hepatitis & \cellcolor{red!10}tuberculosis & \cellcolor{red!10}infect & \cellcolor{red!10}viral & \mybar{1551} \\ 
   66 & \cellcolor{red!10}drinking & \cellcolor{red!10}water & \cellcolor{red!10}reservoir & \cellcolor{red!10}contaminant & \cellcolor{red!10}irrigation & \cellcolor{red!10}tap & \mybar{228} \\ 
   68 & \cellcolor{red!10}spray & \cellcolor{red!10}mosquito & \cellcolor{red!10}pesticide & \cellcolor{red!10}pest & \cellcolor{red!10}repellent & \cellcolor{red!10}soap & \mybar{898} \\ 
   33 & \cellcolor{red!10}fire & \cellcolor{red!10}alarm & \cellcolor{red!10}firefighter & \cellcolor{red!10}rescue & \cellcolor{red!10}apparatus & \cellcolor{red!10}emergency & \mybar{271} \\ 
   56 & \cellcolor{red!10}holiday & \cellcolor{red!10}weekend & \cellcolor{red!10}parade & \cellcolor{red!10}event & \cellcolor{red!10}auburn & \cellcolor{red!10}host & \mybar{283} \\ 
   44 & \cellcolor{red!10}microchip & \cellcolor{red!10}cat & \cellcolor{red!10}euthanasia & \cellcolor{red!10}spay & \cellcolor{red!10}rabies & \cellcolor{red!10}neuter & \mybar{1229} \\ 
   70 & \cellcolor{red!10}election & \cellcolor{red!10}ethic & \cellcolor{red!10}ballot & \cellcolor{red!10}political & \cellcolor{red!10}candidate & \cellcolor{red!10}lobbyist & \mybar{382} \\ 
   60 & \cellcolor{red!10}shall & \cellcolor{red!10}unless & \cellcolor{red!10}except & \cellcolor{red!10}mean & \cellcolor{red!10}deem & \cellcolor{red!10}forth & \mybar{103} \\ 
   59 & \cellcolor{red!10}motion & \cellcolor{red!10}unanimously & \cellcolor{red!10}adjourn & \cellcolor{red!10}prince & \cellcolor{red!10}carry & \cellcolor{red!10}ken & \mybar{220} \\ 
   81 & \cellcolor{red!10}effluent & \cellcolor{red!10}sludge & \cellcolor{red!10}wastewater & \cellcolor{red!10}mercury & \cellcolor{red!10}lbs & \cellcolor{red!10}gal & \mybar{537} \\ 
   89 & \cellcolor{red!10}ask & \cellcolor{red!10}explain & \cellcolor{red!10}say & \cellcolor{red!10}reply & \cellcolor{red!10}suggest & \cellcolor{red!10}ruff & \mybar{427} \\ 
    5 & \cellcolor{red!10}home & \cellcolor{red!10}family & \cellcolor{red!10}homeowner & \cellcolor{red!10}single & \cellcolor{red!10}residence & \cellcolor{red!10}cottage & \mybar{94} \\ 
  105 & \cellcolor{red!10}proposer & \cellcolor{red!10}breach & \cellcolor{red!10}franchisee & \cellcolor{red!10}indemnify & \cellcolor{red!10}agree & \cellcolor{red!10}hereunder & \mybar{273} \\ 
  119 & \cellcolor{red!10}alderman & \cellcolor{red!10}councilor & \cellcolor{red!10}councilwoman & \cellcolor{red!10}alderwoman & \cellcolor{red!10}common & \cellcolor{red!10}roll & \mybar{260} \\ 
   14 & \cellcolor{red!10}borough & \cellcolor{red!10}exam & \cellcolor{red!10}trademark & \cellcolor{red!10}veteran & \cellcolor{red!10}immigrant & \cellcolor{red!10}new & \mybar{359} \\ 
  116 & \cellcolor{red!10}asthma & \cellcolor{red!10}overdose & \cellcolor{red!10}diabetes & \cellcolor{red!10}obesity & \cellcolor{red!10}hospitalization & \cellcolor{red!10}prevalence & \mybar{609} \\ 
   37 & \cellcolor{red!10}dental & \cellcolor{red!10}medicare & \cellcolor{red!10}deductible & \cellcolor{red!10}coinsurance & \cellcolor{red!10}prescription & \cellcolor{red!10}copay & \mybar{444} \\ 
   67 & \cellcolor{red!10}plat & \cellcolor{red!10}thence & \cellcolor{red!10}easement & \cellcolor{red!10}pud & \cellcolor{red!10}petitioner & \cellcolor{red!10}annexation & \mybar{271} \\ 
   35 & \cellcolor{red!10}parent & \cellcolor{red!10}youth & \cellcolor{red!10}child & \cellcolor{red!10}mentor & \cellcolor{red!10}literacy & \cellcolor{red!10}foster & \mybar{326} \\ 
   75 & \cellcolor{red!10}website & \cellcolor{red!10}plain & \cellcolor{red!10}please & \cellcolor{red!10}online & \cellcolor{red!10}customize & \cellcolor{red!10}contact & \mybar{98} \\ 
   21 & \cellcolor{red!10}bid & \cellcolor{red!10}bidder & \cellcolor{red!10}contractor & \cellcolor{red!10}subcontractor & \cellcolor{red!10}contract & \cellcolor{red!10}procurement & \mybar{238} \\ 
   95 & \cellcolor{red!10}duct & \cellcolor{red!10}valve & \cellcolor{red!10}splice & \cellcolor{red!10}piping & \cellcolor{red!10}conduit & \cellcolor{red!10}conductor & \mybar{850} \\ 
   83 & \cellcolor{red!10}storm & \cellcolor{red!10}runoff & \cellcolor{red!10}drainage & \cellcolor{red!10}sewer & \cellcolor{red!10}sanitary & \cellcolor{red!10}infiltration & \mybar{224} \\ 
   64 & \cellcolor{red!10}discrimination & \cellcolor{red!10}gender & \cellcolor{red!10}disability & \cellcolor{red!10}race & \cellcolor{red!10}religion & \cellcolor{red!10}racial & \mybar{437} \\ 
   32 & \cellcolor{red!10}think & \cellcolor{red!10}really & \cellcolor{red!10}thing & \cellcolor{red!10}something & \cellcolor{red!10}maybe & \cellcolor{red!10}just & \mybar{899} \\ 
   49 & \cellcolor{red!10}recycling & \cellcolor{red!10}recycle & \cellcolor{red!10}garbage & \cellcolor{red!10}trash & \cellcolor{red!10}waste & \cellcolor{red!10}bin & \mybar{405} \\ 
    3 & \cellcolor{red!10}maturity & \cellcolor{red!10}portfolio & \cellcolor{red!10}rating & \cellcolor{red!10}jun & \cellcolor{red!10}yield & \cellcolor{red!10}investment & \mybar{276} \\ 
    8 & \cellcolor{red!10}invoice & \cellcolor{red!10}payment & \cellcolor{red!10}card & \cellcolor{red!10}cash & \cellcolor{red!10}account & \cellcolor{red!10}amt & \mybar{222} \\ 
   12 & \cellcolor{red!10}password & \cellcolor{red!10}header & \cellcolor{red!10}archive & \cellcolor{red!10}browser & \cellcolor{red!10}folder & \cellcolor{red!10}text & \mybar{552} \\ 
   90 & \cellcolor{red!10}student & \cellcolor{red!10}school & \cellcolor{red!10}elementary & \cellcolor{red!10}college & \cellcolor{red!10}academic & \cellcolor{red!10}graduate & \mybar{303} \\ 
   48 & \cellcolor{red!10}application & \cellcolor{red!10}applicant & \cellcolor{red!10}must & \cellcolor{red!10}certificate & \cellcolor{red!10}license & \cellcolor{red!10}proof & \mybar{150} \\ 
   92 & \cellcolor{red!10}food & \cellcolor{red!10}calorie & \cellcolor{red!10}meat & \cellcolor{red!10}vend & \cellcolor{red!10}utensil & \cellcolor{red!10}salad & \mybar{1291} \\ 
   86 & \cellcolor{red!10}whereas & \cellcolor{red!10}hereby & \cellcolor{red!10}resolve & \cellcolor{red!10}bond & \cellcolor{red!10}anticipation & \cellcolor{red!10}redemption & \mybar{194} \\ 
   26 & \cellcolor{red!10}petroleum & \cellcolor{red!10}spill & \cellcolor{red!10}contamination & \cellcolor{red!10}asbestos & \cellcolor{red!10}contaminate & \cellcolor{red!10}radioactive & \mybar{444} \\ 
    4 & \cellcolor{red!10}para & \cellcolor{red!10}persona & \cellcolor{red!10}ante & \cellcolor{red!10}horas & \cellcolor{red!10}junta & \cellcolor{red!10}sin & \mybar{644} \\ 
    7 & \cellcolor{red!10}energy & \cellcolor{red!10}renewable & \cellcolor{red!10}solar & \cellcolor{red!10}electricity & \cellcolor{red!10}climate & \cellcolor{red!10}efficiency & \mybar{416} \\ 
   22 & \cellcolor{red!10}wireless & \cellcolor{red!10}server & \cellcolor{red!10}software & \cellcolor{red!10}telecommunication & \cellcolor{red!10}cable & \cellcolor{red!10}technology & \mybar{376} \\ 
   76 & \cellcolor{red!10}year & \cellcolor{red!10}fiscal & \cellcolor{red!10}five & \cellcolor{red!10}annual & \cellcolor{red!10}last & \cellcolor{red!10}three & \mybar{50} \\ 
   34 & \cellcolor{red!10}fee & \cellcolor{red!10}charge & \cellcolor{red!10}per & \cellcolor{red!10}cost & \cellcolor{red!10}plus & \cellcolor{red!10}hourly & \mybar{109} \\ 
  109 & \cellcolor{red!10}city & \cellcolor{red!10}fort & \cellcolor{red!10}manager & \cellcolor{red!10}worth & \cellcolor{red!10}hall & \cellcolor{red!10}municipal & \mybar{10} \\ 
  103 & \cellcolor{red!10}com & \cellcolor{red!10}perm & \cellcolor{red!10}tor & \cellcolor{red!10}cigarette & \cellcolor{red!10}loo & \cellcolor{red!10}comm & \mybar{1386} \\ 
   79 & \cellcolor{red!10}tow & \cellcolor{red!10}plow & \cellcolor{red!10}vehicle & \cellcolor{red!10}trailer & \cellcolor{red!10}motor & \cellcolor{red!10}truck & \mybar{594} \\ 
   54 & \cellcolor{red!10}dwell & \cellcolor{red!10}building & \cellcolor{red!10}remodel & \cellcolor{red!10}unit & \cellcolor{red!10}occupancy & \cellcolor{red!10}alteration & \mybar{132} \\ 
   23 & \cellcolor{red!10}name & \cellcolor{red!10}address & \cellcolor{red!10}description & \cellcolor{red!10}number & \cellcolor{red!10}list & \cellcolor{red!10}zip & \mybar{92} \\ 
   69 & \cellcolor{red!10}vista & \cellcolor{red!10}ranch & \cellcolor{red!10}suite & \cellcolor{red!10}trinity & \cellcolor{red!10}coliseum & \cellcolor{red!10}mesa & \mybar{657} \\ 
   61 & \cellcolor{red!10}cannabis & \cellcolor{red!10}marijuana & \cellcolor{red!10}cultivation & \cellcolor{red!10}dispensary & \cellcolor{red!10}collective & \cellcolor{red!10}liquor & \mybar{470} \\ 
  114 & \cellcolor{red!10}beach & \cellcolor{red!10}orange & \cellcolor{red!10}platinum & \cellcolor{red!10}resort & \cellcolor{red!10}ocean & \cellcolor{red!10}angel & \mybar{517} \\ 
    1 & \cellcolor{red!10}landlord & \cellcolor{red!10}tenant & \cellcolor{red!10}lease & \cellcolor{red!10}lessee & \cellcolor{red!10}golf & \cellcolor{red!10}rent & \mybar{288} \\ 
  118 & \cellcolor{white}flood & \cellcolor{white}earthquake & \cellcolor{white}floodplain & \cellcolor{white}hurricane & \cellcolor{white}tornado & \cellcolor{white}disaster & \mybar{513} \\ 
   99 & \cellcolor{white}roof & \cellcolor{white}porch & \cellcolor{white}awning & \cellcolor{white}masonry & \cellcolor{white}brick & \cellcolor{white}vinyl & \mybar{729} \\ 
  111 & \cellcolor{white}buffalo & \cellcolor{white}player & \cellcolor{white}league & \cellcolor{white}ballpark & \cellcolor{white}baseball & \cellcolor{white}football & \mybar{542} \\ 
  112 & \cellcolor{white}excavation & \cellcolor{white}trench & \cellcolor{white}excavate & \cellcolor{white}gravel & \cellcolor{white}silt & \cellcolor{white}concrete & \mybar{696} \\ 
   53 & \cellcolor{white}downtown & \cellcolor{white}mall & \cellcolor{white}hotel & \cellcolor{white}midtown & \cellcolor{white}uptown & \cellcolor{white}shopping & \mybar{531} \\ 
   \hline
\end{tabular}
\endgroup
\caption{Top words from a structural topic model with 120 topics (first 60 topics displayed here) and FREX scoring. Colors depict partisanship based on coefficient size. White cells are non-significant topics.} 
\end{table}