### 2018-06-26 paper/tables/stmTopWords120_2.tex

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Jun 26 15:42:11 2018
\begin{table}[ht]
\centering
\begingroup\scriptsize
\begin{tabular}{rllllllll}
  \hline
 \# & Top Word 1 & Top Word 2 & Top Word 3 & Top Word 4 & Top Word 5 & Top Word 6 & \multicolumn{2}{c}{Tokens assigned} \\ 
  \hline
 53 & \cellcolor{white}downtown & \cellcolor{white}mall & \cellcolor{white}hotel & \cellcolor{white}midtown & \cellcolor{white}uptown & \cellcolor{white}shopping & \mybar{531} \\ 
  117 & \cellcolor{white}police & \cellcolor{white}patrol & \cellcolor{white}chief & \cellcolor{white}lieutenant & \cellcolor{white}captain & \cellcolor{white}swear & \mybar{283} \\ 
    2 & \cellcolor{red!10}page & \cellcolor{red!10}yon & \cellcolor{red!10}rev & \cellcolor{red!10}sou & \cellcolor{red!10}spec & \cellcolor{red!10}gen & \mybar{165} \\ 
  100 & \cellcolor{white}senate & \cellcolor{white}house & \cellcolor{white}butler & \cellcolor{white}hook & \cellcolor{white}rep & \cellcolor{white}haven & \mybar{590} \\ 
   55 & \cellcolor{white}chapter & \cellcolor{white}code & \cellcolor{white}section & \cellcolor{white}subsection & \cellcolor{white}article & \cellcolor{white}amend & \mybar{124} \\ 
   19 & \cellcolor{white}fugitive & \cellcolor{white}noise & \cellcolor{white}exhaust & \cellcolor{white}receptor & \cellcolor{white}coal & \cellcolor{white}ozone & \mybar{437} \\ 
   31 & \cellcolor{white}aviation & \cellcolor{white}taxicab & \cellcolor{white}airport & \cellcolor{white}runway & \cellcolor{white}airline & \cellcolor{white}hangar & \mybar{498} \\ 
   85 & \cellcolor{white}homeless & \cellcolor{white}homelessness & \cellcolor{white}supportive & \cellcolor{white}client & \cellcolor{white}transitional & \cellcolor{white}encampment & \mybar{232} \\ 
   42 & \cellcolor{white}tax & \cellcolor{white}exemption & \cellcolor{white}taxable & \cellcolor{white}deduction & \cellcolor{white}taxpayer & \cellcolor{white}appraisal & \mybar{160} \\ 
   80 & \cellcolor{white}artist & \cellcolor{white}artwork & \cellcolor{white}art & \cellcolor{white}exhibition & \cellcolor{white}gallery & \cellcolor{white}artistic & \mybar{1060} \\ 
   65 & \cellcolor{white}density & \cellcolor{white}land & \cellcolor{white}us & \cellcolor{white}urban & \cellcolor{white}village & \cellcolor{white}growth & \mybar{102} \\ 
  101 & \cellcolor{white}marsh & \cellcolor{white}riparian & \cellcolor{white}habitat & \cellcolor{white}wetland & \cellcolor{white}grassland & \cellcolor{white}freshwater & \mybar{1110} \\ 
  120 & \cellcolor{blue!10}bend & \cellcolor{blue!10}rogers & \cellcolor{blue!10}walnut & \cellcolor{blue!10}grape & \cellcolor{blue!10}parenthood & \cellcolor{blue!10}shalom & \mybar{315} \\ 
  108 & \cellcolor{blue!10}owner & \cellcolor{blue!10}inspector & \cellcolor{blue!10}property & \cellcolor{blue!10}inspection & \cellcolor{blue!10}unsafe & \cellcolor{blue!10}nuisance & \mybar{156} \\ 
  110 & \cellcolor{white}incumbent & \cellcolor{white}ability & \cellcolor{white}supervise & \cellcolor{white}knowledge & \cellcolor{white}supervision & \cellcolor{white}essential & \mybar{378} \\ 
  102 & \cellcolor{blue!10}parking & \cellcolor{blue!10}space & \cellcolor{blue!10}height & \cellcolor{blue!10}garage & \cellcolor{blue!10}foot & \cellcolor{blue!10}lot & \mybar{83} \\ 
   30 & \cellcolor{blue!10}figure & \cellcolor{blue!10}census & \cellcolor{blue!10}population & \cellcolor{blue!10}respondent & \cellcolor{blue!10}comparison & \cellcolor{blue!10}table & \mybar{240} \\ 
   71 & \cellcolor{blue!10}economic & \cellcolor{blue!10}workforce & \cellcolor{blue!10}economy & \cellcolor{blue!10}industry & \cellcolor{blue!10}sector & \cellcolor{blue!10}job & \mybar{312} \\ 
   91 & \cellcolor{blue!10}prune & \cellcolor{blue!10}forestry & \cellcolor{blue!10}tree & \cellcolor{blue!10}planting & \cellcolor{blue!10}shrub & \cellcolor{blue!10}root & \mybar{1092} \\ 
   28 & \cellcolor{blue!10}conviction & \cellcolor{blue!10}guilty & \cellcolor{blue!10}offense & \cellcolor{blue!10}convict & \cellcolor{blue!10}misdemeanor & \cellcolor{blue!10}felony & \mybar{762} \\ 
  115 & \cellcolor{blue!10}mitigation & \cellcolor{blue!10}impact & \cellcolor{blue!10}adverse & \cellcolor{blue!10}significant & \cellcolor{blue!10}mitigate & \cellcolor{blue!10}measure & \mybar{135} \\ 
   27 & \cellcolor{blue!10}workshop & \cellcolor{blue!10}learn & \cellcolor{blue!10}tour & \cellcolor{blue!10}upcoming & \cellcolor{blue!10}get & \cellcolor{blue!10}view & \mybar{119} \\ 
   16 & \cellcolor{blue!10}park & \cellcolor{blue!10}recreation & \cellcolor{blue!10}playground & \cellcolor{blue!10}picnic & \cellcolor{blue!10}trail & \cellcolor{blue!10}zoo & \mybar{253} \\ 
   15 & \cellcolor{blue!10}landmark & \cellcolor{blue!10}historic & \cellcolor{blue!10}revival & \cellcolor{blue!10}preservation & \cellcolor{blue!10}archaeological & \cellcolor{blue!10}historical & \mybar{936} \\ 
   10 & \cellcolor{blue!10}ave & \cellcolor{blue!10}rainier & \cellcolor{blue!10}beacon & \cellcolor{blue!10}aurora & \cellcolor{blue!10}greenwood & \cellcolor{blue!10}capitol & \mybar{353} \\ 
   51 & \cellcolor{blue!10}waterfront & \cellcolor{blue!10}boat & \cellcolor{blue!10}shoreline & \cellcolor{blue!10}maritime & \cellcolor{blue!10}dock & \cellcolor{blue!10}port & \mybar{788} \\ 
   82 & \cellcolor{blue!10}avenue & \cellcolor{blue!10}east & \cellcolor{blue!10}west & \cellcolor{blue!10}north & \cellcolor{blue!10}street & \cellcolor{blue!10}south & \mybar{78} \\ 
   73 & \cellcolor{blue!10}actuarial & \cellcolor{blue!10}pension & \cellcolor{blue!10}retirement & \cellcolor{blue!10}retiree & \cellcolor{blue!10}unfunded & \cellcolor{blue!10}contribution & \mybar{181} \\ 
    6 & \cellcolor{blue!10}variance & \cellcolor{blue!10}setback & \cellcolor{blue!10}fence & \cellcolor{blue!10}exception & \cellcolor{blue!10}yard & \cellcolor{blue!10}nonconforming & \mybar{122} \\ 
   46 & \cellcolor{blue!10}allegation & \cellcolor{blue!10}complainant & \cellcolor{blue!10}misconduct & \cellcolor{blue!10}complaint & \cellcolor{blue!10}bias & \cellcolor{blue!10}allege & \mybar{631} \\ 
   25 & \cellcolor{blue!10}bankruptcy & \cellcolor{blue!10}plaintiff & \cellcolor{blue!10}examiner & \cellcolor{blue!10}creditor & \cellcolor{blue!10}trial & \cellcolor{blue!10}appeal & \mybar{843} \\ 
   78 & \cellcolor{blue!10}violent & \cellcolor{blue!10}gang & \cellcolor{blue!10}violence & \cellcolor{blue!10}inmate & \cellcolor{blue!10}crime & \cellcolor{blue!10}offender & \mybar{710} \\ 
   97 & \cellcolor{blue!10}employee & \cellcolor{blue!10}sick & \cellcolor{blue!10}wage & \cellcolor{blue!10}grievance & \cellcolor{blue!10}bargaining & \cellcolor{blue!10}overtime & \mybar{243} \\ 
   77 & \cellcolor{blue!10}board & \cellcolor{blue!10}appoint & \cellcolor{blue!10}chairperson & \cellcolor{blue!10}secretary & \cellcolor{blue!10}member & \cellcolor{blue!10}vice & \mybar{137} \\ 
   24 & \cellcolor{blue!10}grant & \cellcolor{blue!10}funding & \cellcolor{blue!10}program & \cellcolor{blue!10}fund & \cellcolor{blue!10}federal & \cellcolor{blue!10}match & \mybar{49} \\ 
  104 & \cellcolor{blue!10}project & \cellcolor{blue!10}improvement & \cellcolor{blue!10}upgrade & \cellcolor{blue!10}replacement & \cellcolor{blue!10}phase & \cellcolor{blue!10}appropriated & \mybar{84} \\ 
   94 & \cellcolor{blue!10}audit & \cellcolor{blue!10}auditing & \cellcolor{blue!10}deficiency & \cellcolor{blue!10}auditor & \cellcolor{blue!10}internal & \cellcolor{blue!10}weakness & \mybar{195} \\ 
   13 & \cellcolor{blue!10}yes & \cellcolor{blue!10}agency & \cellcolor{blue!10}successor & \cellcolor{blue!10}redevelopment & \cellcolor{blue!10}oversight & \cellcolor{blue!10}disposition & \mybar{128} \\ 
    9 & \cellcolor{blue!10}realm & \cellcolor{blue!10}design & \cellcolor{blue!10}proponent & \cellcolor{blue!10}courtyard & \cellcolor{blue!10}facade & \cellcolor{blue!10}concept & \mybar{468} \\ 
   96 & \cellcolor{blue!10}propose & \cellcolor{blue!10}draft & \cellcolor{blue!10}comment & \cellcolor{blue!10}alternative & \cellcolor{blue!10}plan & \cellcolor{blue!10}planning & \mybar{68} \\ 
  106 & \cellcolor{blue!10}sidewalk & \cellcolor{blue!10}crosswalk & \cellcolor{blue!10}signal & \cellcolor{blue!10}traffic & \cellcolor{blue!10}intersection & \cellcolor{blue!10}curb & \mybar{269} \\ 
   41 & \cellcolor{blue!10}bicycle & \cellcolor{blue!10}bike & \cellcolor{blue!10}transit & \cellcolor{blue!10}bus & \cellcolor{blue!10}route & \cellcolor{blue!10}mobility & \mybar{279} \\ 
   63 & \cellcolor{blue!10}memorandum & \cellcolor{blue!10}council & \cellcolor{blue!10}resolution & \cellcolor{blue!10}negotiation & \cellcolor{blue!10}manager & \cellcolor{blue!10}ward & \mybar{132} \\ 
   39 & \cellcolor{blue!10}commission & \cellcolor{blue!10}committee & \cellcolor{blue!10}commissioner & \cellcolor{blue!10}advisory & \cellcolor{blue!10}chair & \cellcolor{blue!10}discussion & \mybar{126} \\ 
   84 & \cellcolor{blue!10}implement & \cellcolor{blue!10}monitor & \cellcolor{blue!10}performance & \cellcolor{blue!10}inventory & \cellcolor{blue!10}process & \cellcolor{blue!10}track & \mybar{146} \\ 
  107 & \cellcolor{blue!10}budget & \cellcolor{blue!10}appropriation & \cellcolor{blue!10}fund & \cellcolor{blue!10}expenditure & \cellcolor{blue!10}adopt & \cellcolor{blue!10}levy & \mybar{99} \\ 
   62 & \cellcolor{blue!10}affordable & \cellcolor{blue!10}housing & \cellcolor{blue!10}affordability & \cellcolor{blue!10}household & \cellcolor{blue!10}income & \cellcolor{blue!10}renter & \mybar{224} \\ 
   17 & \cellcolor{blue!10}million & \cellcolor{blue!10}revenue & \cellcolor{blue!10}forecast & \cellcolor{blue!10}offset & \cellcolor{blue!10}deficit & \cellcolor{blue!10}projection & \mybar{187} \\ 
   74 & \cellcolor{blue!10}neighborhood & \cellcolor{blue!10}vision & \cellcolor{blue!10}attractive & \cellcolor{blue!10}node & \cellcolor{blue!10}amenity & \cellcolor{blue!10}corridor & \mybar{351} \\ 
   45 & \cellcolor{blue!10}zoning & \cellcolor{blue!10}district & \cellcolor{blue!10}zone & \cellcolor{blue!10}acre & \cellcolor{blue!10}dist & \cellcolor{blue!10}rezoning & \mybar{71} \\ 
  113 & \cellcolor{blue!10}debt & \cellcolor{blue!10}governmental & \cellcolor{blue!10}asset & \cellcolor{blue!10}net & \cellcolor{blue!10}statement & \cellcolor{blue!10}obligation & \mybar{133} \\ 
   72 & \cellcolor{blue!10}rouge & \cellcolor{blue!10}parish & \cellcolor{blue!10}baton & \cellcolor{blue!10}hogan & \cellcolor{blue!10}councilman & \cellcolor{blue!10}bowman & \mybar{528} \\ 
   88 & \cellcolor{blue!10}position & \cellcolor{blue!10}staffing & \cellcolor{blue!10}citywide & \cellcolor{blue!10}analyst & \cellcolor{blue!10}strategic & \cellcolor{blue!10}allocation & \mybar{111} \\ 
   40 & \cellcolor{blue!10}accessibility & \cellcolor{blue!10}mode & \cellcolor{blue!10}false & \cellcolor{blue!10}null & \cellcolor{blue!10}else & \cellcolor{blue!10}tab & \mybar{105} \\ 
   11 & \cellcolor{blue!10}strategy & \cellcolor{blue!10}goal & \cellcolor{blue!10}stakeholder & \cellcolor{blue!10}strategic & \cellcolor{blue!10}engagement & \cellcolor{blue!10}outreach & \mybar{168} \\ 
   58 & \cellcolor{blue!10}news & \cellcolor{blue!10}warren & \cellcolor{blue!10}announce & \cellcolor{blue!10}lovely & \cellcolor{blue!10}release & \cellcolor{blue!10}today & \mybar{501} \\ 
   50 & \cellcolor{blue!10}aye & \cellcolor{blue!10}absent & \cellcolor{blue!10}khan & \cellcolor{blue!10}nay & \cellcolor{blue!10}berry & \cellcolor{blue!10}voting & \mybar{318} \\ 
   87 & \cellcolor{blue!20}digest & \cellcolor{blue!20}proposal & \cellcolor{blue!20}reappoints & \cellcolor{blue!20}sander & \cellcolor{blue!20}gray & \cellcolor{blue!20}metropolitan & \mybar{232} \\ 
   29 & \cellcolor{blue!20}agenda & \cellcolor{blue!20}speaker & \cellcolor{blue!20}item & \cellcolor{blue!20}divided & \cellcolor{blue!20}speak & \cellcolor{blue!20}refrain & \mybar{146} \\ 
   47 & \cellcolor{blue!20}consolidated & \cellcolor{blue!20}reinvestment & \cellcolor{blue!20}contingency & \cellcolor{blue!20}contract & \cellcolor{blue!20}authorize & \cellcolor{blue!20}engineering & \mybar{131} \\ 
   57 & \cellcolor{blue!80}suspect & \cellcolor{blue!80}fatal & \cellcolor{blue!80}shoot & \cellcolor{blue!80}pronounce & \cellcolor{blue!80}stopper & \cellcolor{blue!80}gunshot & \mybar{500} \\ 
   \hline
\end{tabular}
\endgroup
\caption{Top words from a structural topic model with 120 topics (second 60 topics displayed here) and FREX scoring. Colors depict partisanship based on coefficient size. White cells are non-significant topics.} 
\end{table}