### 2018-07-13 paper/tables/tabFiletypeFrequencies.tex
updated the table describing filetypes

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Fri Jul 13 15:10:43 2018
\begin{table}[ht]
\centering
\begin{tabular}{lll}
  \hline
Filetype & Occurances After & Occurances After \\ 
  \hline
html & 211682 & 887362 \\ 
  pdf & 464842 & 638802 \\ 
  jpg & 0 & 36958 \\ 
  xml & 0 & 29638 \\ 
  Other & 162681 & 9475 \\ 
  ics & 435 & 8950 \\ 
  png & 0 & 8863 \\ 
  doc & 6972 & 8430 \\ 
  txt & 317 & 6025 \\ 
   & 793990 & 5234 \\ 
  docx & 3137 & 4319 \\ 
  TOTAL & 1644056 & 1644056 \\ 
   \hline
\end{tabular}
\caption{Number of files per type, before and after detecing them via their magic number. The table shows that a lot of files originally have the wrong type, and that converting them correctly has a large impact on how many of them end up being usable.} 
\end{table}