### 2018-07-29 paper/tables/boilerplateClassifierRFMetrics.tex
updated random forest classifier

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Sat Jul 28 20:34:44 2018
\begin{table}[ht]
\centering
\begin{tabular}{rr}
  \hline
 & Value \\ 
  \hline
Percent Correctly Predicted & 0.87 \\ 
  Precision & 0.87 \\ 
  Recall & 0.91 \\ 
  F1-Score & 0.89 \\ 
   \hline
\end{tabular}
\caption{Performance metrics for random forest boilerplate classifier, with inverse probability weights.} 
\label{randomForest}
\end{table}