### 2018-07-29 paper/tables/tabBoilerplateIllustration.tex
updated the boilerplate illustration figures and table to the new weights

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Sat Jul 28 20:51:59 2018
\begin{table}[ht]
\centering
\begingroup\small
\begin{tabular}{lrr}
  \hline
Line & Substantive & Boilerplate \\ 
  \hline
have questions & 0.00 & 1.00 \\ 
  stay connected & 0.00 & 1.00 \\ 
  accident or injury & 0.00 & 1.00 \\ 
  fire damaged buildings & 0.00 & 1.00 \\ 
  gas connections & 0.00 & 1.00 \\ 
  harboring of vagrants & 0.00 & 1.00 \\ 
  roof system problem & 0.00 & 1.00 \\ 
  violation plat note & 0.00 & 1.00 \\ 
  violation setback & 0.00 & 1.00 \\ 
  violation site plan & 0.00 & 1.00 \\ 
   \hline
\end{tabular}
\endgroup
\caption{Lines (or the first 50 characters of a line) in the corpus of Anchorage, AK, with the 10 highest probabilities of being classified as boilerplate.} 
\end{table}