### 2018-08-05 paper/tables/stmTopWords120.tex
updated the model with 120 topics

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Sun Aug  5 19:54:59 2018
\begin{table}[ht]
\centering
\begingroup\scriptsize
\begin{tabular}{rllllllll}
  \hline
 \# & Top Word 1 & Top Word 2 & Top Word 3 & Top Word 4 & Top Word 5 & Top Word 6 & \multicolumn{2}{c}{Tokens assigned} \\ 
  \hline
115 & \cellcolor{red!40}garland & \cellcolor{red!40}celebration & \cellcolor{red!40}blog & \cellcolor{red!40}dream & \cellcolor{red!40}sorry & \cellcolor{red!40}copyright & \mybar{994} \\ 
   52 & \cellcolor{red!20}dog & \cellcolor{red!20}legislator & \cellcolor{red!20}spay & \cellcolor{red!20}neuter & \cellcolor{red!20}animal & \cellcolor{red!20}microchip & \mybar{761} \\ 
   44 & \cellcolor{red!20}copy & \cellcolor{red!20}record & \cellcolor{red!20}mail & \cellcolor{red!20}request & \cellcolor{red!20}notice & \cellcolor{red!20}notify & \mybar{120} \\ 
   98 & \cellcolor{red!20}neighborhood & \cellcolor{red!20}community & \cellcolor{red!20}resident & \cellcolor{red!20}safe & \cellcolor{red!20}life & \cellcolor{red!20}quality & \mybar{95} \\ 
   88 & \cellcolor{red!20}war & \cellcolor{red!20}professor & \cellcolor{red!20}sister & \cellcolor{red!20}bachelor & \cellcolor{red!20}daughter & \cellcolor{red!20}soldier & \mybar{2516} \\ 
   43 & \cellcolor{red!20}camp & \cellcolor{red!20}yoga & \cellcolor{red!20}camper & \cellcolor{red!20}fun & \cellcolor{red!20}librarian & \cellcolor{red!20}library & \mybar{1080} \\ 
   42 & \cellcolor{red!10}infection & \cellcolor{red!10}tuberculosis & \cellcolor{red!10}breastfeed & \cellcolor{red!10}hepatitis & \cellcolor{red!10}vaccine & \cellcolor{red!10}condom & \mybar{1608} \\ 
   72 & \cellcolor{red!10}drinking & \cellcolor{red!10}water & \cellcolor{red!10}contaminant & \cellcolor{red!10}reservoir & \cellcolor{red!10}pipeline & \cellcolor{red!10}irrigation & \mybar{216} \\ 
   84 & \cellcolor{red!10}say & \cellcolor{red!10}ask & \cellcolor{red!10}explain & \cellcolor{red!10}reply & \cellcolor{red!10}horn & \cellcolor{red!10}advise & \mybar{454} \\ 
   18 & \cellcolor{red!10}player & \cellcolor{red!10}coach & \cellcolor{red!10}game & \cellcolor{red!10}umpire & \cellcolor{red!10}ball & \cellcolor{red!10}shirt & \mybar{1595} \\ 
   61 & \cellcolor{red!10}unanimously & \cellcolor{red!10}motion & \cellcolor{red!10}prince & \cellcolor{red!10}adjourn & \cellcolor{red!10}carry & \cellcolor{red!10}ken & \mybar{192} \\ 
   63 & \cellcolor{red!10}mosquito & \cellcolor{red!10}spray & \cellcolor{red!10}rodent & \cellcolor{red!10}pesticide & \cellcolor{red!10}repellent & \cellcolor{red!10}pest & \mybar{851} \\ 
   81 & \cellcolor{red!10}effluent & \cellcolor{red!10}sludge & \cellcolor{red!10}lbs & \cellcolor{red!10}mercury & \cellcolor{red!10}wastewater & \cellcolor{red!10}gal & \mybar{540} \\ 
   60 & \cellcolor{red!10}shall & \cellcolor{red!10}deem & \cellcolor{red!10}forth & \cellcolor{red!10}unless & \cellcolor{red!10}except & \cellcolor{red!10}thereof & \mybar{119} \\ 
   69 & \cellcolor{red!10}ethic & \cellcolor{red!10}candidate & \cellcolor{red!10}lobbyist & \cellcolor{red!10}filer & \cellcolor{red!10}political & \cellcolor{red!10}officeholder & \mybar{355} \\ 
   33 & \cellcolor{red!10}think & \cellcolor{red!10}really & \cellcolor{red!10}something & \cellcolor{red!10}thing & \cellcolor{red!10}just & \cellcolor{red!10}go & \mybar{826} \\ 
  119 & \cellcolor{red!10}firefighter & \cellcolor{red!10}fire & \cellcolor{red!10}chief & \cellcolor{red!10}police & \cellcolor{red!10}captain & \cellcolor{red!10}patrol & \mybar{248} \\ 
   37 & \cellcolor{red!10}physician & \cellcolor{red!10}nursing & \cellcolor{red!10}medical & \cellcolor{red!10}nurse & \cellcolor{red!10}outpatient & \cellcolor{red!10}medicaid & \mybar{352} \\ 
    5 & \cellcolor{red!10}home & \cellcolor{red!10}homeowner & \cellcolor{red!10}alarm & \cellcolor{red!10}detector & \cellcolor{red!10}monoxide & \cellcolor{red!10}header & \mybar{209} \\ 
   23 & \cellcolor{red!10}proposer & \cellcolor{red!10}bidder & \cellcolor{red!10}subcontractor & \cellcolor{red!10}bid & \cellcolor{red!10}contractor & \cellcolor{red!10}subcontract & \mybar{239} \\ 
  116 & \cellcolor{red!10}councilor & \cellcolor{red!10}alderman & \cellcolor{red!10}councilwoman & \cellcolor{red!10}alderwoman & \cellcolor{red!10}quill & \cellcolor{red!10}councilors & \mybar{268} \\ 
   15 & \cellcolor{red!10}trademark & \cellcolor{red!10}borough & \cellcolor{red!10}new & \cellcolor{red!10}immigration & \cellcolor{red!10}immigrant & \cellcolor{red!10}pour & \mybar{274} \\ 
   67 & \cellcolor{red!10}discrimination & \cellcolor{red!10}disability & \cellcolor{red!10}gender & \cellcolor{red!10}religion & \cellcolor{red!10}accommodation & \cellcolor{red!10}origin & \mybar{373} \\ 
  117 & \cellcolor{red!10}asthma & \cellcolor{red!10}overdose & \cellcolor{red!10}obesity & \cellcolor{red!10}hospitalization & \cellcolor{red!10}diabetes & \cellcolor{red!10}prevalence & \mybar{659} \\ 
   94 & \cellcolor{red!10}duct & \cellcolor{red!10}valve & \cellcolor{red!10}sprinkler & \cellcolor{red!10}combustible & \cellcolor{red!10}splice & \cellcolor{red!10}conductor & \mybar{778} \\ 
   58 & \cellcolor{red!10}event & \cellcolor{red!10}firework & \cellcolor{red!10}parade & \cellcolor{red!10}press & \cellcolor{red!10}holiday & \cellcolor{red!10}troy & \mybar{335} \\ 
   70 & \cellcolor{red!10}whereas & \cellcolor{red!10}hereby & \cellcolor{red!10}resolve & \cellcolor{red!10}duly & \cellcolor{red!10}authorize & \cellcolor{red!10}therefore & \mybar{202} \\ 
   30 & \cellcolor{red!10}disaster & \cellcolor{red!10}emergency & \cellcolor{red!10}preparedness & \cellcolor{red!10}evacuation & \cellcolor{red!10}dispatch & \cellcolor{red!10}homeland & \mybar{365} \\ 
   38 & \cellcolor{red!10}student & \cellcolor{red!10}parent & \cellcolor{red!10}school & \cellcolor{red!10}teacher & \cellcolor{red!10}academic & \cellcolor{red!10}youth & \mybar{354} \\ 
   93 & \cellcolor{red!10}city & \cellcolor{red!10}fort & \cellcolor{red!10}worth & \cellcolor{red!10}manager & \cellcolor{red!10}hall & \cellcolor{red!10}charter & \mybar{16} \\ 
   75 & \cellcolor{red!10}online & \cellcolor{red!10}click & \cellcolor{red!10}plain & \cellcolor{red!10}website & \cellcolor{red!10}download & \cellcolor{red!10}learn & \mybar{165} \\ 
    3 & \cellcolor{red!10}value & \cellcolor{red!10}market & \cellcolor{red!10}productivity & \cellcolor{red!10}customize & \cellcolor{red!10}yrs & \cellcolor{red!10}index & \mybar{126} \\ 
   49 & \cellcolor{red!10}recycling & \cellcolor{red!10}recycle & \cellcolor{red!10}garbage & \cellcolor{red!10}waste & \cellcolor{red!10}trash & \cellcolor{red!10}landfill & \mybar{408} \\ 
  111 & \cellcolor{red!10}franchisee & \cellcolor{red!10}indemnify & \cellcolor{red!10}arise & \cellcolor{red!10}harmless & \cellcolor{red!10}breach & \cellcolor{red!10}party & \mybar{307} \\ 
   17 & \cellcolor{red!10}snow & \cellcolor{red!10}plow & \cellcolor{red!10}tornado & \cellcolor{red!10}flood & \cellcolor{red!10}pothole & \cellcolor{red!10}crew & \mybar{552} \\ 
   89 & \cellcolor{red!10}vend & \cellcolor{red!10}food & \cellcolor{red!10}meat & \cellcolor{red!10}utensil & \cellcolor{red!10}calorie & \cellcolor{red!10}vending & \mybar{1174} \\ 
   45 & \cellcolor{red!10}application & \cellcolor{red!10}applicant & \cellcolor{red!10}certificate & \cellcolor{red!10}must & \cellcolor{red!10}license & \cellcolor{red!10}permit & \mybar{151} \\ 
   85 & \cellcolor{red!10}runoff & \cellcolor{red!10}sanitary & \cellcolor{red!10}infiltration & \cellcolor{red!10}storm & \cellcolor{red!10}drainage & \cellcolor{red!10}drain & \mybar{241} \\ 
  106 & \cellcolor{red!10}equipment & \cellcolor{red!10}boiler & \cellcolor{red!10}fleet & \cellcolor{red!10}crane & \cellcolor{red!10}mechanic & \cellcolor{red!10}fuel & \mybar{539} \\ 
    8 & \cellcolor{red!10}invoice & \cellcolor{red!10}payment & \cellcolor{red!10}card & \cellcolor{red!10}credit & \cellcolor{red!10}account & \cellcolor{red!10}cash & \mybar{187} \\ 
   13 & \cellcolor{red!10}class & \cellcolor{red!10}test & \cellcolor{red!10}adobe & \cellcolor{red!10}embed & \cellcolor{red!10}reader & \cellcolor{red!10}acrobat & \mybar{312} \\ 
  108 & \cellcolor{red!10}cigarette & \cellcolor{red!10}senate & \cellcolor{red!10}tobacco & \cellcolor{red!10}consumer & \cellcolor{red!10}smoking & \cellcolor{red!10}ban & \mybar{542} \\ 
   25 & \cellcolor{red!10}coal & \cellcolor{red!10}hazard & \cellcolor{red!10}hazardous & \cellcolor{red!10}toxic & \cellcolor{red!10}radiation & \cellcolor{red!10}substance & \mybar{288} \\ 
   86 & \cellcolor{red!10}groundwater & \cellcolor{red!10}sample & \cellcolor{red!10}asbestos & \cellcolor{red!10}analytical & \cellcolor{red!10}remediation & \cellcolor{red!10}remedial & \mybar{345} \\ 
    1 & \cellcolor{red!10}golf & \cellcolor{red!10}exhibit & \cellcolor{red!10}lessee & \cellcolor{red!10}course & \cellcolor{red!10}lessor & \cellcolor{red!10}lease & \mybar{401} \\ 
    9 & \cellcolor{red!10}para & \cellcolor{red!10}persona & \cellcolor{red!10}ante & \cellcolor{red!10}horas & \cellcolor{red!10}junta & \cellcolor{red!10}sin & \mybar{635} \\ 
   24 & \cellcolor{red!10}phone & \cellcolor{red!10}name & \cellcolor{red!10}page & \cellcolor{red!10}address & \cellcolor{red!10}glen & \cellcolor{red!10}cove & \mybar{158} \\ 
    7 & \cellcolor{red!10}energy & \cellcolor{red!10}renewable & \cellcolor{red!10}solar & \cellcolor{red!10}electricity & \cellcolor{red!10}climate & \cellcolor{red!10}efficiency & \mybar{399} \\ 
   66 & \cellcolor{red!10}plat & \cellcolor{red!10}thence & \cellcolor{red!10}easement & \cellcolor{red!10}pud & \cellcolor{red!10}tract & \cellcolor{red!10}subdivision & \mybar{230} \\ 
   57 & \cellcolor{red!10}dwell & \cellcolor{red!10}unit & \cellcolor{red!10}remodel & \cellcolor{red!10}condominium & \cellcolor{red!10}dwelling & \cellcolor{red!10}residential & \mybar{167} \\ 
   95 & \cellcolor{red!10}roof & \cellcolor{red!10}masonry & \cellcolor{red!10}porch & \cellcolor{red!10}exterior & \cellcolor{red!10}would & \cellcolor{red!10}brick & \mybar{611} \\ 
   26 & \cellcolor{red!10}fee & \cellcolor{red!10}charge & \cellcolor{red!10}per & \cellcolor{red!10}cost & \cellcolor{red!10}plus & \cellcolor{red!10}rate & \mybar{102} \\ 
   51 & \cellcolor{red!10}chapter & \cellcolor{red!10}code & \cellcolor{red!10}violation & \cellcolor{red!10}subsection & \cellcolor{red!10}article & \cellcolor{red!10}sec & \mybar{151} \\ 
   59 & \cellcolor{white}zoning & \cellcolor{white}conditional & \cellcolor{white}zone & \cellcolor{white}cannabis & \cellcolor{white}overlay & \cellcolor{white}district & \mybar{241} \\ 
  101 & \cellcolor{white}height & \cellcolor{white}foot & \cellcolor{white}square & \cellcolor{white}feet & \cellcolor{white}setback & \cellcolor{white}frontage & \mybar{124} \\ 
   96 & \cellcolor{red!10}house & \cellcolor{red!10}cemetery & \cellcolor{red!10}burial & \cellcolor{red!10}butler & \cellcolor{red!10}funeral & \cellcolor{red!10}barber & \mybar{472} \\ 
   65 & \cellcolor{red!10}ballot & \cellcolor{red!10}vista & \cellcolor{red!10}ranch & \cellcolor{red!10}canyon & \cellcolor{red!10}silicon & \cellcolor{red!10}voter & \mybar{518} \\ 
  120 & \cellcolor{red!10}bend & \cellcolor{red!10}fir & \cellcolor{red!10}hometown & \cellcolor{red!10}twelfth & \cellcolor{red!10}exceptional & \cellcolor{red!10}rodeo & \mybar{271} \\ 
   36 & \cellcolor{white}aviation & \cellcolor{white}airport & \cellcolor{white}airline & \cellcolor{white}runway & \cellcolor{white}aircraft & \cellcolor{white}hangar & \mybar{429} \\ 
   34 & \cellcolor{white}plan & \cellcolor{white}planning & \cellcolor{white}comprehensive & \cellcolor{white}master & \cellcolor{white}review & \cellcolor{white}amendment & \mybar{42} \\ 
   \hline
\end{tabular}
\endgroup
\caption{Top words from a structural topic model with 120 topics (first 60 topics displayed here) and FREX scoring. Colors depict partisanship based on coefficient size. White cells are non-significant topics.} 
\end{table}