### 2018-08-05 paper/tables/stmTopWords120_2.tex

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Sun Aug  5 19:55:00 2018
\begin{table}[ht]
\centering
\begingroup\scriptsize
\begin{tabular}{rllllllll}
  \hline
 \# & Top Word 1 & Top Word 2 & Top Word 3 & Top Word 4 & Top Word 5 & Top Word 6 & \multicolumn{2}{c}{Tokens assigned} \\ 
  \hline
 34 & \cellcolor{white}plan & \cellcolor{white}planning & \cellcolor{white}comprehensive & \cellcolor{white}master & \cellcolor{white}review & \cellcolor{white}amendment & \mybar{42} \\ 
   82 & \cellcolor{white}com & \cellcolor{white}mar & \cellcolor{white}spec & \cellcolor{white}jun & \cellcolor{white}est & \cellcolor{white}comm & \mybar{1388} \\ 
   22 & \cellcolor{white}server & \cellcolor{white}software & \cellcolor{white}wireless & \cellcolor{white}technology & \cellcolor{white}desktop & \cellcolor{white}broadband & \mybar{430} \\ 
   80 & \cellcolor{white}artist & \cellcolor{white}art & \cellcolor{white}artwork & \cellcolor{white}exhibition & \cellcolor{white}artistic & \cellcolor{white}sculpture & \mybar{1099} \\ 
  113 & \cellcolor{white}trench & \cellcolor{white}thickness & \cellcolor{white}compaction & \cellcolor{white}concrete & \cellcolor{white}slab & \cellcolor{white}excavation & \mybar{766} \\ 
   87 & \cellcolor{white}respondent & \cellcolor{white}survey & \cellcolor{white}census & \cellcolor{white}racial & \cellcolor{white}demographic & \cellcolor{white}score & \mybar{427} \\ 
   83 & \cellcolor{white}homeless & \cellcolor{white}homelessness & \cellcolor{white}supportive & \cellcolor{white}client & \cellcolor{white}transitional & \cellcolor{white}encampment & \mybar{229} \\ 
   20 & \cellcolor{white}noise & \cellcolor{white}fugitive & \cellcolor{white}receptor & \cellcolor{white}exhaust & \cellcolor{white}vibration & \cellcolor{white}emission & \mybar{376} \\ 
   35 & \cellcolor{white}landlord & \cellcolor{white}tenant & \cellcolor{white}owner & \cellcolor{white}property & \cellcolor{white}rent & \cellcolor{white}lien & \mybar{205} \\ 
  105 & \cellcolor{white}beach & \cellcolor{white}orange & \cellcolor{white}arena & \cellcolor{white}rainier & \cellcolor{white}ocean & \cellcolor{white}resort & \mybar{457} \\ 
    2 & \cellcolor{blue!10}yon & \cellcolor{blue!10}bay & \cellcolor{blue!10}gen & \cellcolor{blue!10}sou & \cellcolor{blue!10}coliseum & \cellcolor{blue!10}estuary & \mybar{385} \\ 
    6 & \cellcolor{white}redevelopment & \cellcolor{white}land & \cellcolor{white}developer & \cellcolor{white}parcel & \cellcolor{white}development & \cellcolor{white}area & \mybar{70} \\ 
  104 & \cellcolor{white}riparian & \cellcolor{white}wetland & \cellcolor{white}habitat & \cellcolor{white}marsh & \cellcolor{white}floodplain & \cellcolor{white}grassland & \mybar{968} \\ 
   41 & \cellcolor{white}tax & \cellcolor{white}exemption & \cellcolor{white}taxable & \cellcolor{white}deduction & \cellcolor{white}levy & \cellcolor{white}taxpayer & \mybar{172} \\ 
   68 & \cellcolor{white}economy & \cellcolor{white}workforce & \cellcolor{white}economic & \cellcolor{white}sector & \cellcolor{white}industry & \cellcolor{white}innovation & \mybar{332} \\ 
   28 & \cellcolor{white}figure & \cellcolor{white}table & \cellcolor{white}scenario & \cellcolor{white}margin & \cellcolor{white}analysis & \cellcolor{white}appendix & \mybar{207} \\ 
  110 & \cellcolor{white}bond & \cellcolor{white}maturity & \cellcolor{white}debt & \cellcolor{white}issuer & \cellcolor{white}redemption & \cellcolor{white}obligation & \mybar{232} \\ 
  102 & \cellcolor{white}sidewalk & \cellcolor{white}curb & \cellcolor{white}pole & \cellcolor{white}crosswalk & \cellcolor{white}ramp & \cellcolor{white}sign & \mybar{237} \\ 
  118 & \cellcolor{blue!10}project & \cellcolor{blue!10}phase & \cellcolor{blue!10}construction & \cellcolor{blue!10}completion & \cellcolor{blue!10}improvement & \cellcolor{blue!10}complete & \mybar{45} \\ 
   78 & \cellcolor{blue!10}parking & \cellcolor{blue!10}tow & \cellcolor{blue!10}vehicle & \cellcolor{blue!10}garage & \cellcolor{blue!10}car & \cellcolor{blue!10}motor & \mybar{210} \\ 
   71 & \cellcolor{blue!10}actuarial & \cellcolor{blue!10}retiree & \cellcolor{blue!10}retirement & \cellcolor{blue!10}pension & \cellcolor{blue!10}deductible & \cellcolor{blue!10}unfunded & \mybar{239} \\ 
   91 & \cellcolor{blue!10}prune & \cellcolor{blue!10}tree & \cellcolor{blue!10}forestry & \cellcolor{blue!10}deer & \cellcolor{blue!10}shrub & \cellcolor{blue!10}planting & \mybar{1240} \\ 
  114 & \cellcolor{blue!10}incumbent & \cellcolor{blue!10}exam & \cellcolor{blue!10}supervision & \cellcolor{blue!10}supervise & \cellcolor{blue!10}examination & \cellcolor{blue!10}ability & \mybar{432} \\ 
   16 & \cellcolor{blue!10}park & \cellcolor{blue!10}recreation & \cellcolor{blue!10}playground & \cellcolor{blue!10}zoo & \cellcolor{blue!10}trail & \cellcolor{blue!10}picnic & \mybar{290} \\ 
   53 & \cellcolor{blue!10}waterfront & \cellcolor{blue!10}boat & \cellcolor{blue!10}shoreline & \cellcolor{blue!10}maritime & \cellcolor{blue!10}dock & \cellcolor{blue!10}barge & \mybar{800} \\ 
   76 & \cellcolor{blue!10}felony & \cellcolor{blue!10}violent & \cellcolor{blue!10}offender & \cellcolor{blue!10}gang & \cellcolor{blue!10}theft & \cellcolor{blue!10}inmate & \mybar{783} \\ 
    4 & \cellcolor{blue!10}courtyard & \cellcolor{blue!10}realm & \cellcolor{blue!10}design & \cellcolor{blue!10}facade & \cellcolor{blue!10}proponent & \cellcolor{blue!10}articulation & \mybar{608} \\ 
  100 & \cellcolor{blue!10}division & \cellcolor{blue!10}manage & \cellcolor{blue!10}staffing & \cellcolor{blue!10}oversee & \cellcolor{blue!10}management & \cellcolor{blue!10}analyst & \mybar{100} \\ 
   97 & \cellcolor{blue!10}mitigation & \cellcolor{blue!10}impact & \cellcolor{blue!10}adverse & \cellcolor{blue!10}significant & \cellcolor{blue!10}alternative & \cellcolor{blue!10}propose & \mybar{132} \\ 
   11 & \cellcolor{blue!10}historic & \cellcolor{blue!10}landmark & \cellcolor{blue!10}revival & \cellcolor{blue!10}archaeological & \cellcolor{blue!10}preservation & \cellcolor{blue!10}historical & \mybar{876} \\ 
   77 & \cellcolor{blue!10}million & \cellcolor{blue!10}fiscal & \cellcolor{blue!10}forecast & \cellcolor{blue!10}revenue & \cellcolor{blue!10}quarter & \cellcolor{blue!10}billion & \mybar{138} \\ 
   74 & \cellcolor{blue!10}board & \cellcolor{blue!10}chairperson & \cellcolor{blue!10}secretary & \cellcolor{blue!10}member & \cellcolor{blue!10}appoint & \cellcolor{blue!10}executive & \mybar{118} \\ 
   47 & \cellcolor{blue!10}allegation & \cellcolor{blue!10}complainant & \cellcolor{blue!10}misconduct & \cellcolor{blue!10}bias & \cellcolor{blue!10}complaint & \cellcolor{blue!10}allege & \mybar{580} \\ 
   92 & \cellcolor{blue!10}sick & \cellcolor{blue!10}employee & \cellcolor{blue!10}wage & \cellcolor{blue!10}overtime & \cellcolor{blue!10}grievance & \cellcolor{blue!10}bargaining & \mybar{260} \\ 
   10 & \cellcolor{blue!10}ave & \cellcolor{blue!10}avenue & \cellcolor{blue!10}south & \cellcolor{blue!10}east & \cellcolor{blue!10}west & \cellcolor{blue!10}blvd & \mybar{189} \\ 
  112 & \cellcolor{blue!10}grant & \cellcolor{blue!10}loan & \cellcolor{blue!10}funding & \cellcolor{blue!10}program & \cellcolor{blue!10}recipient & \cellcolor{blue!10}federal & \mybar{85} \\ 
   56 & \cellcolor{blue!10}downtown & \cellcolor{blue!10}mall & \cellcolor{blue!10}midtown & \cellcolor{blue!10}uptown & \cellcolor{blue!10}hotel & \cellcolor{blue!10}shopping & \mybar{414} \\ 
   14 & \cellcolor{blue!10}yes & \cellcolor{blue!10}agency & \cellcolor{blue!10}successor & \cellcolor{blue!10}oversight & \cellcolor{blue!10}attachment & \cellcolor{blue!10}describe & \mybar{125} \\ 
   40 & \cellcolor{blue!10}bicycle & \cellcolor{blue!10}bike & \cellcolor{blue!10}transit & \cellcolor{blue!10}bicyclist & \cellcolor{blue!10}lane & \cellcolor{blue!10}bus & \mybar{315} \\ 
   62 & \cellcolor{blue!10}affordable & \cellcolor{blue!10}housing & \cellcolor{blue!10}affordability & \cellcolor{blue!10}income & \cellcolor{blue!10}household & \cellcolor{blue!10}moderate & \mybar{188} \\ 
   99 & \cellcolor{blue!10}memorandum & \cellcolor{blue!10}resolution & \cellcolor{blue!10}council & \cellcolor{blue!10}legislation & \cellcolor{blue!10}entitle & \cellcolor{blue!10}commission & \mybar{173} \\ 
   19 & \cellcolor{blue!10}governmental & \cellcolor{blue!10}accounting & \cellcolor{blue!10}asset & \cellcolor{blue!10}statement & \cellcolor{blue!10}financial & \cellcolor{blue!10}net & \mybar{156} \\ 
  103 & \cellcolor{blue!10}permission & \cellcolor{blue!10}ayes & \cellcolor{blue!10}correspondence & \cellcolor{blue!10}bid & \cellcolor{blue!10}smith & \cellcolor{blue!10}demolition & \mybar{203} \\ 
  107 & \cellcolor{blue!10}appropriated & \cellcolor{blue!10}dollars & \cellcolor{blue!10}thousand & \cellcolor{blue!10}ongoing & \cellcolor{blue!10}matrix & \cellcolor{blue!10}justification & \mybar{117} \\ 
   12 & \cellcolor{blue!10}approach & \cellcolor{blue!10}difficult & \cellcolor{blue!10}achieve & \cellcolor{blue!10}challenge & \cellcolor{blue!10}critical & \cellcolor{blue!10}often & \mybar{257} \\ 
   46 & \cellcolor{blue!10}variance & \cellcolor{blue!10}fence & \cellcolor{blue!10}setback & \cellcolor{blue!10}exception & \cellcolor{blue!10}yard & \cellcolor{blue!10}applicant & \mybar{136} \\ 
   90 & \cellcolor{blue!10}audit & \cellcolor{blue!10}auditor & \cellcolor{blue!10}procedure & \cellcolor{blue!10}internal & \cellcolor{blue!10}auditing & \cellcolor{blue!10}documentation & \mybar{226} \\ 
   64 & \cellcolor{blue!10}density & \cellcolor{blue!10}urban & \cellcolor{blue!10}corridor & \cellcolor{blue!10}village & \cellcolor{blue!10}orient & \cellcolor{blue!10}transit & \mybar{165} \\ 
   21 & \cellcolor{blue!10}goal & \cellcolor{blue!10}strategy & \cellcolor{blue!10}outreach & \cellcolor{blue!10}priority & \cellcolor{blue!10}strategic & \cellcolor{blue!10}implementation & \mybar{105} \\ 
   73 & \cellcolor{blue!10}parish & \cellcolor{blue!10}rouge & \cellcolor{blue!10}baton & \cellcolor{blue!10}hogan & \cellcolor{blue!10}councilman & \cellcolor{blue!10}thereto & \mybar{482} \\ 
   29 & \cellcolor{blue!10}comment & \cellcolor{blue!10}draft & \cellcolor{blue!10}discussion & \cellcolor{blue!10}feedback & \cellcolor{blue!10}discuss & \cellcolor{blue!10}presentation & \mybar{168} \\ 
   32 & \cellcolor{blue!10}budget & \cellcolor{blue!10}expenditure & \cellcolor{blue!10}appropriation & \cellcolor{blue!10}fund & \cellcolor{blue!10}endorse & \cellcolor{blue!10}balance & \mybar{129} \\ 
   54 & \cellcolor{blue!10}aye & \cellcolor{blue!10}absent & \cellcolor{blue!10}khan & \cellcolor{blue!10}nay & \cellcolor{blue!10}berry & \cellcolor{blue!10}voting & \mybar{344} \\ 
   39 & \cellcolor{blue!10}mode & \cellcolor{blue!10}accessibility & \cellcolor{blue!10}tab & \cellcolor{blue!10}focus & \cellcolor{blue!10}else & \cellcolor{blue!10}alt & \mybar{117} \\ 
  109 & \cellcolor{blue!10}auburn & \cellcolor{blue!10}buffalo & \cellcolor{blue!10}ward & \cellcolor{blue!10}brown & \cellcolor{blue!10}announce & \cellcolor{blue!10}casino & \mybar{177} \\ 
   50 & \cellcolor{blue!10}news & \cellcolor{blue!10}warren & \cellcolor{blue!10}lovely & \cellcolor{blue!10}release & \cellcolor{blue!10}leader & \cellcolor{blue!10}proud & \mybar{498} \\ 
   79 & \cellcolor{blue!20}digest & \cellcolor{blue!20}proposal & \cellcolor{blue!20}sander & \cellcolor{blue!20}reappoints & \cellcolor{blue!20}metropolitan & \cellcolor{blue!20}gray & \mybar{236} \\ 
   27 & \cellcolor{blue!20}bankruptcy & \cellcolor{blue!20}plaintiff & \cellcolor{blue!20}creditor & \cellcolor{blue!20}trial & \cellcolor{blue!20}court & \cellcolor{blue!20}supreme & \mybar{810} \\ 
   31 & \cellcolor{blue!20}agenda & \cellcolor{blue!20}speaker & \cellcolor{blue!20}item & \cellcolor{blue!20}committee & \cellcolor{blue!20}chair & \cellcolor{blue!20}divided & \mybar{146} \\ 
   48 & \cellcolor{blue!20}consolidated & \cellcolor{blue!20}contingency & \cellcolor{blue!20}reinvestment & \cellcolor{blue!20}inc & \cellcolor{blue!20}contract & \cellcolor{blue!20}authorize & \mybar{134} \\ 
   55 & \cellcolor{blue!80}suspect & \cellcolor{blue!80}shoot & \cellcolor{blue!80}fatal & \cellcolor{blue!80}homicide & \cellcolor{blue!80}stopper & \cellcolor{blue!80}pronounce & \mybar{512} \\ 
   \hline
\end{tabular}
\endgroup
\caption{Top words from a structural topic model with 120 topics (second 60 topics displayed here) and FREX scoring. Colors depict partisanship based on coefficient size. White cells are non-significant topics.} 
\end{table}