### 2018-09-14 paper/TADA2018_Poster/main.tex
updates to the poster (not done yet)

\documentclass{beamer}
%% Possible paper sizes: a0, a0b, a1, a2, a3, a4.
%% Possible orientations: portrait, landscape
%% Font sizes can be changed using the scale option.
\usepackage[size=a3,orientation=landscape,scale=3.75]{beamerposter}
\usetheme{LLT-poster}
%\usecolortheme{ComingClean}
\usecolortheme{Entrepreneur}
%\usecolortheme{ConspicuousCreep}  %% VERY garish.
\usepackage{graphicx}
\usepackage{xcolor, colortbl}
\usepackage{array}

\def\mybar#1{%%
    #1 & {\color{red}\pgfmathsetlengthmacro\x{#1*0.006mm}\rule{\x}{4pt}}}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{libertine}
\usepackage[scaled=0.92]{inconsolata}
\usepackage[libertine]{newtxmath}
\usepackage{geometry}
\geometry{paperwidth=32in,paperheight=40in,margin=2cm}
\usepackage{multicol}
\usepackage{mwe}

\author[mvn5218@psu.edu]{Markus Neumann}
\title{Government websites as data: A methodological pipeline with application to the websites of municipalities in the United States \\ \vspace{10mm} \small{Markus Neumann (mvn5218@psu.edu),  Fridolin Linder (flinder@gmail.com),  Bruce Desmarais (brucedesmarais@psu.edu) \\ Department of Political Science - Penn State University}}
\institute{The Pennsylvania State University}
% Optional foot image
%\footimage{\includegraphics[width=4cm]{IMG_1934.jpg}}

\begin{document}
\beamertemplatenavigationsymbolsempty
\begin{frame}[fragile]
\begin{columns}[T]

%%%% 1st Column
\begin{column}{.4\textwidth}

\begin{block}{Overview}
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{eriemayor.png}
    %\caption{Caption}
    %\label{fig:my_label}
\end{figure}
\begin{itemize}
\item Government websites contain important information
\item Research has largely relied on manual coding
\item Our contribution: pipeline for automated analysis
\item Application: websites of partisan municipalities
\end{itemize}
\end{block}

\begin{block}{Data Collection}
\begin{itemize}
\item City URLs are scraped from Wikipedia and the GSA
\end{itemize}
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{us_map_resized.png}
    %\caption{Caption}
    %\label{fig:my_label}
\end{figure}
%\include{tabCitiesStatesParties.tex}
\end{block}

\begin{block}{Site to Text Conversion}
\input{tabFiletypeFrequencies.tex}
\begin{itemize}
\item The file endings from city websites are sometimes wrong
\item We use file signatures to identify the correct type
\item Preprocessing: lowercase; removal of punctuation, numbers, dates, etc.; removal of non-English words; lemmatization
\end{itemize}
\end{block}

\end{column}
\begin{column}{.6\textwidth}
\begin{block}{Boilerplate Removal}
\begin{itemize}
\item Websites contain a lot repetitive and uninformative content
\end{itemize}
\begin{figure}
    \centering
    \includegraphics[width=.8\linewidth]{arcadia_water_TADA_rect.png}
    %\caption{Caption}
    %\label{fig:my_label}
\end{figure}
\begin{itemize}
\item A classifier (random forest) is trained to identify and remove boilerplate lines
\end{itemize}
\begin{minipage}{\textwidth}
  \begin{minipage}[b]{0.49\textwidth}
    \begin{figure}
        \centering
        \includegraphics[width=0.85\linewidth]{boilerplateBeforeAfterNwords.pdf}
        %\caption{Caption}
        %\label{fig:my_label}
    \end{figure}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.49\textwidth}
    \input{boilerplateClassifierRFMetrics.tex}
    \end{minipage}
  \end{minipage}
\end{block}

%\begin{column}{.63\textwidth}
%\begin{block}{Additional Preprocessing}
%\begin{itemize}
%\item Lowercase; removal of punctuation, numbers, dates, etc.
%\item Spellchecking/Removal of non-English words
%\item Lemmatization
%\end{itemize}
%\end{block}
%\end{column}

%%%% 2nd Column

%\begin{column}{.23\textwidth}

%\end{column}

%%%%  3rd column

%\begin{column}{.23\textwidth}

\begin{block}{Analysis}

\begin{center}
\begin{itemize}
\item Structural topic model with 60 topics
\item Covariates: Party, state, population, median income
\end{itemize}

\vspace{5mm}

\input{stmTopWords60.tex}

%\begin{figure}
%    \centering
%    \includegraphics[width=\linewidth]{stm_test.png}
%    %\caption{Caption}
%    \label{fig:my_label}
%\end{figure}

\end{center}

\end{block}

\begin{block}{Conclusion}
\begin{itemize}
\item Republican cities feature information on basic utilities and protection from natural disasters
\item Democratic cities feature information on policy deliberation, crime control and city budgeting
\item Politics at the municipal level is not entirely non-partisan
\item We plan to implement the pipeline in an R package
\end{itemize}
\end{block}

%\end{column}
\end{column}

\end{columns}

\end{frame}
\end{document}