### 2018-09-14 paper/TADA2018_Poster/stmTopWords60.tex

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Fri Aug 10 15:33:35 2018
\begin{table}[ht]
\centering
\begingroup\footnotesize
\begin{tabular}{rllllllll}
  \hline
 \# & Top Word 1 & Top Word 2 & Top Word 3 & Top Word 4 & Top Word 5 & Top Word 6 & \multicolumn{2}{c}{Tokens assigned} \\ 
  \hline
   42 & \cellcolor{red!20}breastfeed & \cellcolor{red!20}vaccine & \cellcolor{red!20}infection & \cellcolor{red!20}symptom & \cellcolor{red!20}asthma & \cellcolor{red!20}mosquito & \mybar{2497} \\ 
   17 & \cellcolor{red!20}alarm & \cellcolor{red!20}disaster & \cellcolor{red!20}fire & \cellcolor{red!20}rescue & \cellcolor{red!20}preparedness & \cellcolor{red!20}evacuation & \mybar{989} \\ 
   53 & \cellcolor{red!20}drinking & \cellcolor{red!20}wastewater & \cellcolor{red!20}water & \cellcolor{red!20}pipeline & \cellcolor{red!20}pump & \cellcolor{red!20}disinfection & \mybar{461} \\

   49 & \cellcolor{red!10}bin & \cellcolor{red!10}recycling & \cellcolor{red!10}garbage & \cellcolor{red!10}recyclables & \cellcolor{red!10}recyclable & \cellcolor{red!10}bag & \mybar{1791} \\ 
    7 & \cellcolor{red!10}energy & \cellcolor{red!10}garland & \cellcolor{red!10}renewable & \cellcolor{red!10}solar & \cellcolor{red!10}electricity & \cellcolor{red!10}climate & \mybar{742} \\ 
   23 & \cellcolor{red!10}bid & \cellcolor{red!10}proposer & \cellcolor{red!10}bidder & \cellcolor{red!10}contractor & \cellcolor{red!10}subcontractor & \cellcolor{red!10}contract & \mybar{447} \\ 
   57 & \cellcolor{red!10}duct & \cellcolor{red!10}conduit & \cellcolor{red!10}bolt & \cellcolor{red!10}splice & \cellcolor{red!10}valve & \cellcolor{red!10}fitting & \mybar{1373} \\ 
   13 & \cellcolor{red!10}server & \cellcolor{red!10}wireless & \cellcolor{red!10}software & \cellcolor{red!10}telecommunication & \cellcolor{red!10}subscriber & \cellcolor{red!10}desktop & \mybar{1092} \\ 
    1 & \cellcolor{red!10}storm & \cellcolor{red!10}runoff & \cellcolor{red!10}infiltration & \cellcolor{red!10}discharge & \cellcolor{red!10}drainage & \cellcolor{red!10}drain & \mybar{516} \\ 
   38 & \cellcolor{red!10}youth & \cellcolor{red!10}student & \cellcolor{red!10}parent & \cellcolor{red!10}teacher & \cellcolor{red!10}immigrant & \cellcolor{red!10}literacy & \mybar{714} \\ 
   59 & \cellcolor{red!10}sampling & \cellcolor{red!10}sample & \cellcolor{red!10}analytical & \cellcolor{red!10}concentration & \cellcolor{red!10}hydrocarbon & \cellcolor{red!10}toxicity & \mybar{1241} \\ 
   45 & \cellcolor{red!10}premise & \cellcolor{red!10}licensee & \cellcolor{red!10}violation & \cellcolor{red!10}license & \cellcolor{red!10}permit & \cellcolor{red!10}inspection & \mybar{509} \\ 
   60 & \cellcolor{red!10}exhaust & \cellcolor{red!10}fugitive & \cellcolor{red!10}aircraft & \cellcolor{red!10}airport & \cellcolor{red!10}aviation & \cellcolor{red!10}diesel & \mybar{731} \\ 
   %30 & \cellcolor{white}fort & \cellcolor{white}thence & \cellcolor{white}blvd & \cellcolor{white}worth & \cellcolor{white}ave & \cellcolor{white}west & \mybar{681} \\ 
   %58 & \cellcolor{white}councilor & \cellcolor{white}auburn & \cellcolor{white}plain & \cellcolor{white}ward & \cellcolor{white}beech & \cellcolor{white}glen & \mybar{480} \\ 
   %51 & \cellcolor{white}whereas & \cellcolor{white}councilman & \cellcolor{white}alderman & \cellcolor{white}ordain & \cellcolor{white}hereby & \cellcolor{white}resolution & \mybar{420} \\ 
   %16 & \cellcolor{white}recreation & \cellcolor{white}park & \cellcolor{white}golf & \cellcolor{white}playground & \cellcolor{white}picnic & \cellcolor{white}zoo & \mybar{682} \\ 
   %36 & \cellcolor{white}retiree & \cellcolor{white}retirement & \cellcolor{white}actuarial & \cellcolor{white}deductible & \cellcolor{white}dental & \cellcolor{white}pension & \mybar{470} \\ 
   %27 & \cellcolor{white}exam & \cellcolor{white}incumbent & \cellcolor{white}supervise & \cellcolor{white}supervision & \cellcolor{white}examination & \cellcolor{white}knowledge & \mybar{687} \\ 
   %56 & \cellcolor{white}historic & \cellcolor{white}landmark & \cellcolor{white}revival & \cellcolor{white}archaeological & \cellcolor{white}century & \cellcolor{white}historian & \mybar{2587} \\ 
   %12 & \cellcolor{white}parking & \cellcolor{white}hotel & \cellcolor{white}garage & \cellcolor{white}space & \cellcolor{white}retail & \cellcolor{white}square & \mybar{321} \\ 
   %41 & \cellcolor{white}tax & \cellcolor{white}exemption & \cellcolor{white}abatement & \cellcolor{white}real & \cellcolor{white}estate & \cellcolor{white}property & \mybar{310} \\ 
    4 & \cellcolor{white}facade & \cellcolor{white}awning & \cellcolor{white}porch & \cellcolor{white}roof & \cellcolor{white}balcony & \cellcolor{white}exterior & \mybar{1108} \\

   15 & \cellcolor{blue!10}complainant & \cellcolor{blue!10}defendant & \cellcolor{blue!10}allegation & \cellcolor{blue!10}complaint & \cellcolor{blue!10}allege & \cellcolor{blue!10}discrimination & \mybar{1384} \\

   14 & \cellcolor{blue!10}yes & \cellcolor{blue!10}agency & \cellcolor{blue!10}federal & \cellcolor{blue!10}recipient & \cellcolor{blue!10}compliance & \cellcolor{blue!10}entity & \mybar{205} \\

   25 & \cellcolor{blue!10}cannabis & \cellcolor{blue!10}marijuana & \cellcolor{blue!10}senate & \cellcolor{blue!10}dispensary & \cellcolor{blue!10}ballot & \cellcolor{blue!10}cultivation & \mybar{1188} \\

   47 & \cellcolor{blue!10}employee & \cellcolor{blue!10}overtime & \cellcolor{blue!10}sick & \cellcolor{blue!10}wage & \cellcolor{blue!10}grievance & \cellcolor{blue!10}bargaining & \mybar{511} \\

   37 & \cellcolor{blue!10}audit & \cellcolor{blue!10}auditor & \cellcolor{blue!10}internal & \cellcolor{blue!10}procedure & \cellcolor{blue!10}accountability & \cellcolor{blue!10}oversight & \mybar{420} \\ 
   21 & \cellcolor{blue!10}housing & \cellcolor{blue!10}affordable & \cellcolor{blue!10}homeless & \cellcolor{blue!10}homelessness & \cellcolor{blue!10}affordability & \cellcolor{blue!10}landlord & \mybar{318} \\ 
   34 & \cellcolor{blue!10}comment & \cellcolor{blue!10}draft & \cellcolor{blue!10}feedback & \cellcolor{blue!10}stakeholder & \cellcolor{blue!10}suggest & \cellcolor{blue!10}discussion & \mybar{289} \\ 
   19 & \cellcolor{blue!10}debt & \cellcolor{blue!10}bond & \cellcolor{blue!10}governmental & \cellcolor{blue!10}obligation & \cellcolor{blue!10}financial & \cellcolor{blue!10}accounting & \mybar{251} \\ 
   40 & \cellcolor{blue!10}bicycle & \cellcolor{blue!10}bike & \cellcolor{blue!10}lane & \cellcolor{blue!10}crosswalk & \cellcolor{blue!10}pedestrian & \cellcolor{blue!10}bicyclist & \mybar{574} \\ 
   32 & \cellcolor{blue!20}budget & \cellcolor{blue!20}revenue & \cellcolor{blue!20}expenditure & \cellcolor{blue!20}appropriation & \cellcolor{blue!20}fund & \cellcolor{blue!20}million & \mybar{242} \\ 
   48 & \cellcolor{blue!20}absent & \cellcolor{blue!20}aye & \cellcolor{blue!20}khan & \cellcolor{blue!20}nay & \cellcolor{blue!20}berry & \cellcolor{blue!20}voting & \mybar{528} \\ 
   31 & \cellcolor{blue!20}chair & \cellcolor{blue!20}agenda & \cellcolor{blue!20}commission & \cellcolor{blue!20}speaker & \cellcolor{blue!20}chairperson & \cellcolor{blue!20}committee & \mybar{314} \\ 
   55 & \cellcolor{blue!80}robbery & \cellcolor{blue!80}homicide & \cellcolor{blue!80}arrest & \cellcolor{blue!80}sergeant & \cellcolor{blue!80}suspect & \cellcolor{blue!80}burglary & \mybar{1395} \\ 
   \hline
\end{tabular}
\endgroup
%\caption{Top words from a structural topic model with 60 topics and FREX scoring. Colors depict partisanship based on coefficient size. White cells are non-significant topics.} 
\label{tabSTMtopwords60}
\end{table}