### 2018-09-28 paper/tables/tabBoilerplateIllustration.tex
addressed the point about the admittedly flawed boilerplate illustration table by making a (hopefully) better one

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Fri Sep 28 17:14:46 2018
\begin{table}[ht]
\centering
\begingroup\small
\begin{tabular}{rl}
  \hline
Boilerplate Probability & Text \\ 
  \hline
1.00 & ri ng h i l \\ 
  1.00 & any others \\ 
  1.00 & premium for \\ 
  1.00 & ra tr \\ 
  1.00 & oa od \\ 
  1.00 & te pos \\ 
  1.00 & design to \\ 
  1.00 & i pli \\ 
  1.00 & wherever possible \\ 
  1.00 & public nuisances \\ 
  1.00 & prime movers \\ 
  1.00 & of the offense \\ 
  1.00 & d nonconformities \\ 
  1.00 & yearshasexceeded costs \\ 
  1.00 & f street \\ 
   \hline
\end{tabular}
\endgroup
\caption{Lines in the corpus of Anchorage, AK, with a probability of 1 of being classified as boilerplate. This table illustrates that lines which get classified as boilerplate largely consist of irrelevant nonsense which needs to be removed.} 
\label{tabBoilerplateIllustration}
\end{table}