### 2018-12-16 paper/presentation/stmTopWordsINDem.tex
new structure

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Tue Dec 26 20:59:22 2017
\begin{table}[ht]
\centering
\scriptsize
\begin{tabular}{llllll}
  \hline
-0.027 & -0.022 & -0.016 & -0.011 & -0.011 & -0.01 \\ 
  \hline
city & school & downtown & city & \cellcolor{blue!25}trash & \cellcolor{blue!25}housing \\ 
  ordinance & community & business & department & city & property \\ 
  approve & program & project & mayor & \cellcolor{blue!25}waste & program \\ 
  resolution & student & city & police & day & \cellcolor{blue!25}fund \\ 
  property & \cellcolor{blue!25}education & \cellcolor{blue!25}development & officer & \cellcolor{blue!25}recycle & home \\ 
  purchase & \cellcolor{blue!25}university & new & public & street & city \\ 
  area & national & center & citizen & collection & project \\ 
  department & award & \cellcolor{blue!25}economic & work & resident & neighborhood \\ 
  contract & high & company & safety & \cellcolor{blue!25}recycling & \cellcolor{blue!25}grant \\ 
  service & year & community & resident & snow & unit \\ 
   \hline
\end{tabular}
\caption{Top Democratic topics and words}%\caption{Top Democratic topics and words (Indiana), according to STM. 
%The words are the top words for the most Democratic/Republican topic, determined
%by the size (and significance) of the coefficient (see table header) of the party covariate.} 
\label{tabSTMINDem}
\end{table}