### 2018-12-16 paper/presentation/stmTopWordsINRep.tex

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Tue Dec 26 20:59:22 2017
\begin{table}[ht]
\centering
\scriptsize
\begin{tabular}{lllllll}
  \hline
 0.021 & 0.019 & 0.017 & 0.017 & 0.013 & 0.012 \\ 
  \hline
 foot & team & ave & request & amp & \cellcolor{red!25}building \\ 
   \cellcolor{red!25}sign & game & inc & board &\cellcolor{red!25} traffic & historic \\ 
   use & play & cross & member & stop & \cellcolor{red!25}build \\ 
   lot & league & creek & service & vehicle & material \\ 
   \cellcolor{red!25}building & camp & \cellcolor{red!25}construction & \cellcolor{red!25}street & block & preservation \\ 
   \cellcolor{red!25}zone & class & blvd & approve & sign & \cellcolor{red!25}wall \\ 
   area & age & park & city &  \cellcolor{red!25}airport & \cellcolor{red!25}roof \\ 
   district & must & lake & purchase & ave & window \\ 
   \cellcolor{red!25}parking & child & hill & move & theft & \cellcolor{red!25}floor \\ 
   residential & participant & ridge & good & signal & new \\ 
   \hline
\end{tabular}
\caption{Top Republican topics and words}%, according to STM. 
%The words are the top words for the most Democratic/Republican topic, determined
%by the size (and significance) of the coefficient (see table header) of the party covariate.} 
\label{tabSTMINRep}
\end{table}