### 2018-12-16 paper/presentation/stmTopWordsLARep.tex

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Tue Dec 26 15:50:18 2017
\begin{table}[ht]
\centering
\begin{tabular}{llllllll}
  \hline
0.071 & 0.054 & 0.054 & 0.034 & 0.033 & 0.024 & 0.023 & 0.02 \\ 
  \hline
event & ordinance & water & street & say & city & city & mayor \\ 
  information & department & emergency & traffic & can & business & meeting & city \\ 
  show & summary & city & parking & make & new & council & parish \\ 
  park & amount & resident & lane & get & mayor & commission & town \\ 
  music & bid & storm & project & take & development & plan & office \\ 
  food & city & weather & work & people & economic & member & hall \\ 
  visit & public & waste & bike & work & million & public & contact \\ 
  weekend & police & system & downtown & need & continue & board & day \\ 
  festival & approve & power & public & city & work & committee & official \\ 
  begin & inc & service & bicycle & help & local & planning & state \\ 
   \hline
\end{tabular}
\caption{Top Republican topics and words (Louisiana), according to STM. 
The words are the top words for the most Democratic/Republican topic, determined
by the size (and significance) of the coefficient (see table header) of the party covariate.} 
\label{tabSTMLARep}
\end{table}