### 2018-12-16 paper/tables/stmTopWordsINDem.tex

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Tue Dec 26 20:59:22 2017
\begin{table}[ht]
\centering
\begin{tabular}{llllllll}
  \hline
-0.027 & -0.022 & -0.016 & -0.015 & -0.012 & -0.011 & -0.011 & -0.01 \\ 
  \hline
city & school & downtown & service & contract & city & trash & housing \\ 
  ordinance & community & business & division & bid & department & city & property \\ 
  approve & program & project & provide & contractor & mayor & waste & program \\ 
  resolution & student & city & city & city & police & day & fund \\ 
  property & education & development & management & agreement & officer & recycle & home \\ 
  purchase & university & new & public & work & public & street & city \\ 
  area & national & center & department & service & citizen & collection & project \\ 
  department & award & economic & program & department & work & resident & neighborhood \\ 
  contract & high & company & include & bidder & safety & recycling & grant \\ 
  service & year & community & office & move & resident & snow & unit \\ 
   \hline
\end{tabular}
\caption{Top Democratic topics and words (Indiana), according to STM. 
The words are the top words for the most Democratic/Republican topic, determined
by the size (and significance) of the coefficient (see table header) of the party covariate.} 
\label{tabSTMINDem}
\end{table}