### 2018-12-16 paper/tables/stmTopWordsIN_LADem.tex

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Tue Mar 20 09:25:03 2018
\begin{table}[ht]
\centering
\begingroup\tiny
\begin{tabular}{llllllll}
  \hline
-0.036 & -0.024 & -0.024 & -0.012 & -0.009 & -0.008 & -0.008 & -0.007 \\ 
  \hline
provide & applicant & community & council & department & shall & business & day \\ 
  otherwise & owner & city & councilman & division & city & development & contract \\ 
  city & application & neighborhood & rise & service & contractor & economic & total \\ 
  respect & new & public & mayor & police & agreement & downtown & payment \\ 
  authorize & building & mayor & city & city & person & job & invoice \\ 
  ordinance & exist & work & state & provide & provide & new & process \\ 
  thereto & proposal & program & grant & public & upon & city & department \\ 
  councilman & retention & resident & street & management & term & project & approval \\ 
  district & per & effort & fund & operation & provision & center & complete \\ 
  property & build & new & discussion & office & time & service & procurement \\ 
   \hline
\end{tabular}
\endgroup
\caption{Republican} 
\label{tabSTMIN_LADem}
\end{table}