### 2018-12-16 paper/tables/stmTopWordsIN_LARep.tex

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Tue Mar 20 09:25:03 2018
\begin{table}[ht]
\centering
\begingroup\tiny
\begin{tabular}{llllllll}
  \hline
0.019 & 0.018 & 0.016 & 0.015 & 0.015 & 0.013 & 0.012 & 0.01 \\ 
  \hline
library & park & motion & inc & proposal & say & lot & city \\ 
  stream & recreation & second & cross & digest & ask & district & information \\ 
  length & pool & make & construction & fund & state & use & day \\ 
  obj & facility & present & drive & grant & proposal & foot & contact \\ 
  pdf & center & approve & street & county & make & zone & can \\ 
  type & program & state & creek & authorize & get & street & resident \\ 
  branch & golf & minute & park & introduce & can & square & please \\ 
  unknown & playground & meeting & hill & district & need & area & department \\ 
  flag & trail & pass & ave & public & go & building & call \\ 
  image & play & mayor & oak & department & work & parking & amp \\ 
   \hline
\end{tabular}
\endgroup
\caption{Democratic} 
\label{tabSTMIN_LARep}
\end{table}