### 2019-05-20 paper/tables/stmTopWords120_2.tex
added new results for model with 120 topics

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Mon May 20 11:38:16 2019
\begin{table}[ht]
\centering
\begingroup\scriptsize
\begin{tabular}{rllllllll}
  \hline
 \# & Top Word 1 & Top Word 2 & Top Word 3 & Top Word 4 & Top Word 5 & Top Word 6 & \multicolumn{2}{c}{Tokens assigned} \\ 
  \hline
 29 & \cellcolor{blue!10}labor & \cellcolor{blue!10}worker & \cellcolor{blue!10}force & \cellcolor{blue!10}unemployed & \cellcolor{blue!10}earnings & \cellcolor{blue!10}civilian & \mybar{80} \\ 
  111 & \cellcolor{blue!10}discharge & \cellcolor{blue!10}pollutant & \cellcolor{blue!10}inspection & \cellcolor{blue!10}inspect & \cellcolor{blue!10}pollution & \cellcolor{blue!10}inspector & \mybar{109} \\ 
   68 & \cellcolor{blue!10}contractual & \cellcolor{blue!10}parts & \cellcolor{blue!10}duke & \cellcolor{blue!10}outside & \cellcolor{blue!10}postage & \cellcolor{blue!10}receipts & \mybar{274} \\ 
   77 & \cellcolor{blue!10}curb & \cellcolor{blue!10}pavement & \cellcolor{blue!10}sidewalk & \cellcolor{blue!10}ramp & \cellcolor{blue!10}gutter & \cellcolor{blue!10}asphalt & \mybar{390} \\ 
   65 & \cellcolor{blue!10}draft & \cellcolor{blue!10}update & \cellcolor{blue!10}process & \cellcolor{blue!10}review & \cellcolor{blue!10}staff & \cellcolor{blue!10}progress & \mybar{67} \\ 
   24 & \cellcolor{blue!10}landlord & \cellcolor{blue!10}tenant & \cellcolor{blue!10}renewal & \cellcolor{blue!10}rent & \cellcolor{blue!10}lease & \cellcolor{blue!10}expired & \mybar{255} \\ 
  106 & \cellcolor{blue!10}consultant & \cellcolor{blue!10}proposer & \cellcolor{blue!10}procurement & \cellcolor{blue!10}contract & \cellcolor{blue!10}firm & \cellcolor{blue!10}subcontractor & \mybar{179} \\ 
   43 & \cellcolor{blue!10}blanket & \cellcolor{blue!10}medicare & \cellcolor{blue!10}payroll & \cellcolor{blue!10}premium & \cellcolor{blue!10}undistributed & \cellcolor{blue!10}refund & \mybar{107} \\ 
  103 & \cellcolor{white}urban & \cellcolor{white}mixed & \cellcolor{white}density & \cellcolor{white}redevelopment & \cellcolor{white}development & \cellcolor{white}industrial & \mybar{115} \\ 
   89 & \cellcolor{blue!10}taxable & \cellcolor{blue!10}res & \cellcolor{blue!10}deed & \cellcolor{blue!10}value & \cellcolor{blue!10}homestead & \cellcolor{blue!10}star & \mybar{41} \\ 
   83 & \cellcolor{blue!10}building & \cellcolor{blue!10}demolition & \cellcolor{blue!10}story & \cellcolor{blue!10}demolish & \cellcolor{blue!10}floor & \cellcolor{blue!10}build & \mybar{82} \\ 
  119 & \cellcolor{blue!10}cost & \cellcolor{blue!10}estimate & \cellcolor{blue!10}estimated & \cellcolor{blue!10}initial & \cellcolor{blue!10}costs & \cellcolor{blue!10}change & \mybar{52} \\ 
  109 & \cellcolor{blue!10}respondent & \cellcolor{blue!10}satisfied & \cellcolor{blue!10}dissatisfied & \cellcolor{blue!10}survey & \cellcolor{blue!10}satisfaction & \cellcolor{blue!10}disagree & \mybar{403} \\ 
   64 & \cellcolor{blue!10}must & \cellcolor{blue!10}signature & \cellcolor{blue!10}copy & \cellcolor{blue!10}application & \cellcolor{blue!10}applicant & \cellcolor{blue!10}submission & \mybar{139} \\ 
   26 & \cellcolor{blue!10}tax & \cellcolor{blue!10}deduction & \cellcolor{blue!10}amt & \cellcolor{blue!10}assessed & \cellcolor{blue!10}bill & \cellcolor{blue!10}abatement & \mybar{171} \\ 
   78 & \cellcolor{blue!10}yes & \cellcolor{blue!10}worksheet & \cellcolor{blue!10}text & \cellcolor{blue!10}pic & \cellcolor{blue!10}font & \cellcolor{blue!10}button & \mybar{476} \\ 
    7 & \cellcolor{blue!10}greenhouse & \cellcolor{blue!10}emission & \cellcolor{blue!10}coal & \cellcolor{blue!10}climate & \cellcolor{blue!10}ozone & \cellcolor{blue!10}dioxide & \mybar{334} \\ 
   54 & \cellcolor{blue!10}parking & \cellcolor{blue!10}tow & \cellcolor{blue!10}taxi & \cellcolor{blue!10}vehicle & \cellcolor{blue!10}shuttle & \cellcolor{blue!10}passenger & \mybar{236} \\ 
   41 & \cellcolor{blue!10}assistant & \cellcolor{blue!10}analyst & \cellcolor{blue!10}technician & \cellcolor{blue!10}aide & \cellcolor{blue!10}specialist & \cellcolor{blue!10}asst & \mybar{119} \\ 
   22 & \cellcolor{blue!10}allocation & \cellcolor{blue!10}val & \cellcolor{blue!10}cove & \cellcolor{blue!10}acct & \cellcolor{blue!10}glen & \cellcolor{blue!10}subtotal & \mybar{79} \\ 
   63 & \cellcolor{blue!10}fee & \cellcolor{blue!10}charge & \cellcolor{blue!10}license & \cellcolor{blue!10}reservation & \cellcolor{blue!10}surcharge & \cellcolor{blue!10}refundable & \mybar{143} \\ 
  117 & \cellcolor{blue!10}delay & \cellcolor{blue!10}perm & \cellcolor{blue!10}queue & \cellcolor{blue!10}peak & \cellcolor{blue!10}flt & \cellcolor{blue!10}detector & \mybar{113} \\ 
    4 & \cellcolor{blue!10}datum & \cellcolor{blue!10}database & \cellcolor{blue!10}copyright & \cellcolor{blue!10}accuracy & \cellcolor{blue!10}data & \cellcolor{blue!10}compile & \mybar{193} \\ 
   45 & \cellcolor{blue!10}audit & \cellcolor{blue!10}auditor & \cellcolor{blue!10}auditing & \cellcolor{blue!10}internal & \cellcolor{blue!10}implemented & \cellcolor{blue!10}procedure & \mybar{222} \\ 
  100 & \cellcolor{blue!10}mitigation & \cellcolor{blue!10}impact & \cellcolor{blue!10}significant & \cellcolor{blue!10}adverse & \cellcolor{blue!10}significance & \cellcolor{blue!10}unavoidable & \mybar{136} \\ 
   88 & \cellcolor{blue!10}gender & \cellcolor{blue!10}discrimination & \cellcolor{blue!10}transgender & \cellcolor{blue!10}immigrant & \cellcolor{blue!10}immigration & \cellcolor{blue!10}religion & \mybar{859} \\ 
    9 & \cellcolor{blue!10}district & \cellcolor{blue!10}zoning & \cellcolor{blue!10}maker & \cellcolor{blue!10}vacancy & \cellcolor{blue!10}speaker & \cellcolor{blue!10}planner & \mybar{45} \\ 
   12 & \cellcolor{blue!10}artist & \cellcolor{blue!10}artwork & \cellcolor{blue!10}art & \cellcolor{blue!10}arts & \cellcolor{blue!10}mural & \cellcolor{blue!10}sculpture & \mybar{1055} \\ 
   94 & \cellcolor{blue!10}contracted & \cellcolor{blue!10}encumbrance & \cellcolor{blue!10}unencumbered & \cellcolor{blue!10}exp & \cellcolor{blue!10}expend & \cellcolor{blue!10}bud & \mybar{71} \\ 
  110 & \cellcolor{blue!10}rouge & \cellcolor{blue!10}parish & \cellcolor{blue!10}baton & \cellcolor{blue!10}thereto & \cellcolor{blue!10}sewerage & \cellcolor{blue!10}adjudicate & \mybar{464} \\ 
   46 & \cellcolor{blue!10}commissioner & \cellcolor{blue!10}chair & \cellcolor{blue!10}commission & \cellcolor{blue!10}committee & \cellcolor{blue!10}briefing & \cellcolor{blue!10}advisory & \mybar{187} \\ 
   85 & \cellcolor{blue!10}sch & \cellcolor{blue!10}min & \cellcolor{blue!10}tin & \cellcolor{blue!10}hump & \cellcolor{blue!10}carpool & \cellcolor{blue!10}qua & \mybar{390} \\ 
   15 & \cellcolor{blue!10}complainant & \cellcolor{blue!10}allegation & \cellcolor{blue!10}allege & \cellcolor{blue!10}complaint & \cellcolor{blue!10}doc & \cellcolor{blue!10}misconduct & \mybar{963} \\ 
   30 & \cellcolor{blue!10}incumbent & \cellcolor{blue!10}examination & \cellcolor{blue!10}supervision & \cellcolor{blue!10}knowledge & \cellcolor{blue!10}exam & \cellcolor{blue!10}ability & \mybar{410} \\ 
  107 & \cellcolor{blue!10}savings & \cellcolor{blue!10}ltd & \cellcolor{blue!10}village & \cellcolor{blue!10}neighborhood & \cellcolor{blue!10}excise & \cellcolor{blue!10}costs & \mybar{81} \\ 
   72 & \cellcolor{blue!10}imp & \cellcolor{blue!10}burglary & \cellcolor{blue!10}theft & \cellcolor{blue!10}testify & \cellcolor{blue!10}petitioner & \cellcolor{blue!10}mischief & \mybar{116} \\ 
   60 & \cellcolor{blue!10}bike & \cellcolor{blue!10}bicycle & \cellcolor{blue!10}bicyclist & \cellcolor{blue!10}pedestrian & \cellcolor{blue!10}route & \cellcolor{blue!10}mobility & \mybar{336} \\ 
   82 & \cellcolor{blue!10}accomplishment & \cellcolor{blue!10}narrative & \cellcolor{blue!10}grantee & \cellcolor{blue!10}outcome & \cellcolor{blue!10}objective & \cellcolor{blue!10}mod & \mybar{101} \\ 
   36 & \cellcolor{blue!10}decline & \cellcolor{blue!10}trend & \cellcolor{blue!10}recession & \cellcolor{blue!10}average & \cellcolor{blue!10}rate & \cellcolor{blue!10}percentage & \mybar{265} \\ 
   52 & \cellcolor{blue!10}homeless & \cellcolor{blue!10}homelessness & \cellcolor{blue!10}supportive & \cellcolor{blue!10}consolidated & \cellcolor{blue!10}transitional & \cellcolor{blue!10}counseling & \mybar{193} \\ 
    1 & \cellcolor{blue!10}alderman & \cellcolor{blue!10}resolved & \cellcolor{blue!10}whereas & \cellcolor{blue!10}resolution & \cellcolor{blue!10}authorizing & \cellcolor{blue!10}authorize & \mybar{245} \\ 
   92 & \cellcolor{blue!10}concept & \cellcolor{blue!10}design & \cellcolor{blue!10}realm & \cellcolor{blue!10}visual & \cellcolor{blue!10}character & \cellcolor{blue!10}conceptual & \mybar{433} \\ 
   71 & \cellcolor{blue!10}bond & \cellcolor{blue!10}obligation & \cellcolor{blue!10}proceeds & \cellcolor{blue!10}redemption & \cellcolor{blue!10}debt & \cellcolor{blue!10}series & \mybar{174} \\ 
   67 & \cellcolor{blue!10}dist & \cellcolor{blue!10}applied & \cellcolor{blue!10}col & \cellcolor{blue!10}occupancy & \cellcolor{blue!10}valuation & \cellcolor{blue!10}monoxide & \mybar{62} \\ 
   25 & \cellcolor{blue!10}scenario & \cellcolor{blue!10}figure & \cellcolor{blue!10}appendix & \cellcolor{blue!10}assume & \cellcolor{blue!10}assumption & \cellcolor{blue!10}model & \mybar{162} \\ 
   38 & \cellcolor{blue!10}horas & \cellcolor{blue!10}persona & \cellcolor{blue!10}para & \cellcolor{blue!10}yon & \cellcolor{blue!10}sou & \cellcolor{blue!10}ante & \mybar{1350} \\ 
   14 & \cellcolor{blue!10}federal & \cellcolor{blue!10}agency & \cellcolor{blue!10}entity & \cellcolor{blue!10}recipient & \cellcolor{blue!10}grant & \cellcolor{blue!10}eligible & \mybar{90} \\ 
   56 & \cellcolor{blue!10}waterfront & \cellcolor{blue!10}shoreline & \cellcolor{blue!10}marina & \cellcolor{blue!10}beach & \cellcolor{blue!10}port & \cellcolor{blue!10}boat & \mybar{844} \\ 
   61 & \cellcolor{blue!10}revenue & \cellcolor{blue!10}balance & \cellcolor{blue!10}expenditure & \cellcolor{blue!10}reserve & \cellcolor{blue!10}forecast & \cellcolor{blue!10}budget & \mybar{101} \\ 
   75 & \cellcolor{blue!10}governmental & \cellcolor{blue!10}asset & \cellcolor{blue!10}liability & \cellcolor{blue!10}assets & \cellcolor{blue!10}statement & \cellcolor{blue!10}pension & \mybar{142} \\ 
   37 & \cellcolor{blue!10}endorse & \cellcolor{blue!10}endorsed & \cellcolor{blue!10}budget & \cellcolor{blue!10}proposed & \cellcolor{blue!10}adopted & \cellcolor{blue!10}adopt & \mybar{111} \\ 
   69 & \cellcolor{blue!10}tree & \cellcolor{blue!10}planned & \cellcolor{blue!10}circumference & \cellcolor{blue!10}gross & \cellcolor{blue!10}density & \cellcolor{blue!10}infill & \mybar{211} \\ 
   90 & \cellcolor{blue!10}councilman & \cellcolor{blue!10}introduced & \cellcolor{blue!10}ordain & \cellcolor{blue!10}ordinance & \cellcolor{blue!10}digest & \cellcolor{blue!10}yea & \mybar{244} \\ 
   97 & \cellcolor{blue!10}actuarial & \cellcolor{blue!10}grievance & \cellcolor{blue!10}employee & \cellcolor{blue!10}retirement & \cellcolor{blue!10}bargaining & \cellcolor{blue!10}actuary & \mybar{250} \\ 
   39 & \cellcolor{blue!10}affordable & \cellcolor{blue!10}housing & \cellcolor{blue!10}affordability & \cellcolor{blue!10}homeowner & \cellcolor{blue!10}income & \cellcolor{blue!10}bedroom & \mybar{150} \\ 
   55 & \cellcolor{blue!10}ave & \cellcolor{blue!10}combo & \cellcolor{blue!10}blossom & \cellcolor{blue!10}pearl & \cellcolor{blue!10}cir & \cellcolor{blue!10}olive & \mybar{1091} \\ 
   13 & \cellcolor{blue!20}strategy & \cellcolor{blue!20}goal & \cellcolor{blue!20}strategic & \cellcolor{blue!20}stakeholder & \cellcolor{blue!20}focus & \cellcolor{blue!20}initiative & \mybar{162} \\ 
   57 & \cellcolor{blue!20}absent & \cellcolor{blue!20}int & \cellcolor{blue!20}preside & \cellcolor{blue!20}ordained & \cellcolor{blue!20}tag & \cellcolor{blue!20}numbers & \mybar{194} \\ 
  101 & \cellcolor{blue!20}violent & \cellcolor{blue!20}gang & \cellcolor{blue!20}firearm & \cellcolor{blue!20}offender & \cellcolor{blue!20}crime & \cellcolor{blue!20}patrol & \mybar{511} \\ 
   93 & \cellcolor{blue!80}shooting & \cellcolor{blue!80}suspect & \cellcolor{blue!80}pronounce & \cellcolor{blue!80}gunshot & \cellcolor{blue!80}flee & \cellcolor{blue!80}shoot & \mybar{730} \\ 
   \hline
\end{tabular}
\endgroup
\caption{Top words from a structural topic model with 120 topics (second 60 topics displayed here) and FREX scoring. Colors depict partisanship based on coefficient size. White cells are non-significant topics.} 
\label{tabSTMtopwords120_2}
\end{table}