### 2019-05-20 paper/tables/stmTopWords120_1.tex

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Mon May 20 11:38:16 2019
\begin{table}[ht]
\centering
\begingroup\scriptsize
\begin{tabular}{rllllllll}
  \hline
 \# & Top Word 1 & Top Word 2 & Top Word 3 & Top Word 4 & Top Word 5 & Top Word 6 & \multicolumn{2}{c}{Tokens assigned} \\ 
  \hline
 96 & \cellcolor{red!40}subcommittee & \cellcolor{red!40}agenda & \cellcolor{red!40}forum & \cellcolor{red!40}speaker & \cellcolor{red!40}item & \cellcolor{red!40}adjournment & \mybar{217} \\ 
   49 & \cellcolor{red!40}prize & \cellcolor{red!40}celebration & \cellcolor{red!40}ceremony & \cellcolor{red!40}parade & \cellcolor{red!40}follower & \cellcolor{red!40}favorite & \mybar{2043} \\ 
  102 & \cellcolor{red!30}motion & \cellcolor{red!30}second & \cellcolor{red!30}adjourn & \cellcolor{red!30}unanimous & \cellcolor{red!30}carry & \cellcolor{red!30}whiting & \mybar{207} \\ 
   73 & \cellcolor{red!20}legislator & \cellcolor{red!20}player & \cellcolor{red!20}football & \cellcolor{red!20}leg & \cellcolor{red!20}town & \cellcolor{red!20}stadium & \mybar{695} \\ 
   95 & \cellcolor{red!20}online & \cellcolor{red!20}email & \cellcolor{red!20}website & \cellcolor{red!20}browser & \cellcolor{red!20}contact & \cellcolor{red!20}server & \mybar{351} \\ 
   70 & \cellcolor{red!20}election & \cellcolor{red!20}ballot & \cellcolor{red!20}lobbyist & \cellcolor{red!20}voter & \cellcolor{red!20}candidate & \cellcolor{red!20}campaign & \mybar{407} \\ 
   74 & \cellcolor{red!20}tentative & \cellcolor{red!20}conditional & \cellcolor{red!20}approval & \cellcolor{red!20}grading & \cellcolor{red!20}attachment & \cellcolor{red!20}deviation & \mybar{177} \\ 
   79 & \cellcolor{red!20}snow & \cellcolor{red!20}remember & \cellcolor{red!20}plow & \cellcolor{red!20}lock & \cellcolor{red!20}scam & \cellcolor{red!20}sure & \mybar{888} \\ 
   28 & \cellcolor{red!20}craftsman & \cellcolor{red!20}revival & \cellcolor{red!20}historic & \cellcolor{red!20}gabled & \cellcolor{red!20}bungalow & \cellcolor{red!20}historical & \mybar{882} \\ 
  114 & \cellcolor{red!20}park & \cellcolor{red!20}playground & \cellcolor{red!20}recreation & \cellcolor{red!20}picnic & \cellcolor{red!20}mesa & \cellcolor{red!20}trail & \mybar{235} \\ 
   11 & \cellcolor{red!10}tuberculosis & \cellcolor{red!10}infection & \cellcolor{red!10}hepatitis & \cellcolor{red!10}overdose & \cellcolor{red!10}influenza & \cellcolor{red!10}vaccine & \mybar{1515} \\ 
   21 & \cellcolor{red!10}think & \cellcolor{red!10}something & \cellcolor{red!10}want & \cellcolor{red!10}thing & \cellcolor{red!10}talk & \cellcolor{red!10}everybody & \mybar{1155} \\ 
   86 & \cellcolor{red!10}sewer & \cellcolor{red!10}sanitary & \cellcolor{red!10}water & \cellcolor{red!10}pipeline & \cellcolor{red!10}drinking & \cellcolor{red!10}wastewater & \mybar{176} \\ 
   59 & \cellcolor{red!10}fort & \cellcolor{red!10}worth & \cellcolor{red!10}plot & \cellcolor{red!10}tad & \cellcolor{red!10}falls & \cellcolor{red!10}demo & \mybar{192} \\ 
   20 & \cellcolor{red!10}subsection & \cellcolor{red!10}licensee & \cellcolor{red!10}article & \cellcolor{red!10}chapter & \cellcolor{red!10}sec & \cellcolor{red!10}shall & \mybar{214} \\ 
   47 & \cellcolor{red!10}inf & \cellcolor{red!10}micron & \cellcolor{red!10}effluent & \cellcolor{red!10}eff & \cellcolor{red!10}sludge & \cellcolor{red!10}isomer & \mybar{591} \\ 
   62 & \cellcolor{red!10}bid & \cellcolor{red!10}buyer & \cellcolor{red!10}seller & \cellcolor{red!10}bidder & \cellcolor{red!10}price & \cellcolor{red!10}quote & \mybar{357} \\ 
   48 & \cellcolor{red!10}contributor & \cellcolor{red!10}instruction & \cellcolor{red!10}filer & \cellcolor{red!10}political & \cellcolor{red!10}officeholder & \cellcolor{red!10}payee & \mybar{79} \\ 
  104 & \cellcolor{red!10}provisions & \cellcolor{red!10}subcontractor & \cellcolor{red!10}surety & \cellcolor{red!10}rev & \cellcolor{red!10}bidder & \cellcolor{red!10}supplementary & \mybar{232} \\ 
   27 & \cellcolor{red!10}breach & \cellcolor{red!10}franchisee & \cellcolor{red!10}hereunder & \cellcolor{red!10}agreement & \cellcolor{red!10}remedy & \cellcolor{red!10}agree & \mybar{213} \\ 
  112 & \cellcolor{red!10}youth & \cellcolor{red!10}camp & \cellcolor{red!10}teach & \cellcolor{red!10}teen & \cellcolor{red!10}lesson & \cellcolor{red!10}yoga & \mybar{722} \\ 
   23 & \cellcolor{red!10}dog & \cellcolor{red!10}rabies & \cellcolor{red!10}euthanasia & \cellcolor{red!10}euthanized & \cellcolor{red!10}pet & \cellcolor{red!10}spay & \mybar{1710} \\ 
   35 & \cellcolor{red!10}trust & \cellcolor{red!10}revocable & \cellcolor{red!10}mfr & \cellcolor{red!10}apportionment & \cellcolor{red!10}living & \cellcolor{red!10}assn & \mybar{285} \\ 
  116 & \cellcolor{red!10}emergency & \cellcolor{red!10}preparedness & \cellcolor{red!10}null & \cellcolor{red!10}dispatch & \cellcolor{red!10}rescue & \cellcolor{red!10}fire & \mybar{340} \\ 
   80 & \cellcolor{red!10}energy & \cellcolor{red!10}efficiency & \cellcolor{red!10}customer & \cellcolor{red!10}saving & \cellcolor{red!10}rebate & \cellcolor{red!10}renewable & \mybar{382} \\ 
  113 & \cellcolor{red!10}proud & \cellcolor{red!10}leadership & \cellcolor{red!10}honor & \cellcolor{red!10}pleased & \cellcolor{red!10}grateful & \cellcolor{red!10}passion & \mybar{1168} \\ 
   18 & \cellcolor{red!10}garland & \cellcolor{red!10}invoice & \cellcolor{red!10}assoc & \cellcolor{red!10}check & \cellcolor{red!10}firefighter & \cellcolor{red!10}association & \mybar{152} \\ 
   81 & \cellcolor{red!10}page & \cellcolor{red!10}last & \cellcolor{red!10}sub & \cellcolor{red!10}update & \cellcolor{red!10}prime & \cellcolor{red!10}award & \mybar{17} \\ 
    2 & \cellcolor{red!10}mosquito & \cellcolor{red!10}insecticide & \cellcolor{red!10}spray & \cellcolor{red!10}bait & \cellcolor{red!10}repellent & \cellcolor{red!10}pesticide & \mybar{997} \\ 
  120 & \cellcolor{red!10}project & \cellcolor{red!10}improvement & \cellcolor{red!10}funding & \cellcolor{red!10}justification & \cellcolor{red!10}completion & \cellcolor{red!10}acquisition & \mybar{47} \\ 
  105 & \cellcolor{red!10}thence & \cellcolor{red!10}plat & \cellcolor{red!10}easement & \cellcolor{red!10}annexation & \cellcolor{red!10}pud & \cellcolor{red!10}westerly & \mybar{255} \\ 
  118 & \cellcolor{red!10}comment & \cellcolor{red!10}concern & \cellcolor{red!10}suggest & \cellcolor{red!10}clarify & \cellcolor{red!10}suggestion & \cellcolor{red!10}dear & \mybar{307} \\ 
   34 & \cellcolor{red!10}library & \cellcolor{red!10}campus & \cellcolor{red!10}doe & \cellcolor{red!10}branch & \cellcolor{red!10}center & \cellcolor{red!10}arena & \mybar{208} \\ 
   40 & \cellcolor{red!10}portfolio & \cellcolor{red!10}treasury & \cellcolor{red!10}investment & \cellcolor{red!10}maturity & \cellcolor{red!10}yield & \cellcolor{red!10}liquidity & \mybar{250} \\ 
  115 & \cellcolor{red!10}masonry & \cellcolor{red!10}plaster & \cellcolor{red!10}joist & \cellcolor{red!10}stud & \cellcolor{red!10}sheathing & \cellcolor{red!10}ceiling & \mybar{875} \\ 
   53 & \cellcolor{red!10}department & \cellcolor{red!10}authority & \cellcolor{red!10}dpt & \cellcolor{red!10}correction & \cellcolor{red!10}citywide & \cellcolor{red!10}transit & \mybar{109} \\ 
    3 & \cellcolor{red!10}vend & \cellcolor{red!10}utensil & \cellcolor{red!10}meat & \cellcolor{red!10}fat & \cellcolor{red!10}cheese & \cellcolor{red!10}salad & \mybar{1325} \\ 
    8 & \cellcolor{red!10}assessor & \cellcolor{red!10}taxpayer & \cellcolor{red!10}determination & \cellcolor{red!10}informal & \cellcolor{red!10}petition & \cellcolor{red!10}notification & \mybar{39} \\ 
   58 & \cellcolor{white}recycling & \cellcolor{white}bag & \cellcolor{white}garbage & \cellcolor{white}recycle & \cellcolor{white}recyclable & \cellcolor{white}recyclables & \mybar{318} \\ 
   87 & \cellcolor{red!10}sign & \cellcolor{red!10}billboard & \cellcolor{red!10}pole & \cellcolor{red!10}speeding & \cellcolor{red!10}illuminate & \cellcolor{red!10}banner & \mybar{472} \\ 
   31 & \cellcolor{red!10}student & \cellcolor{red!10}elementary & \cellcolor{red!10}school & \cellcolor{red!10}college & \cellcolor{red!10}graduate & \cellcolor{red!10}academic & \mybar{233} \\ 
   32 & \cellcolor{red!10}dwelling & \cellcolor{red!10}alteration & \cellcolor{red!10}plumbing & \cellcolor{red!10}plumb & \cellcolor{red!10}canceled & \cellcolor{red!10}mechanical & \mybar{143} \\ 
   51 & \cellcolor{red!10}combustible & \cellcolor{red!10}vent & \cellcolor{red!10}piping & \cellcolor{red!10}conductor & \cellcolor{red!10}duct & \cellcolor{red!10}flammable & \mybar{517} \\ 
   91 & \cellcolor{red!10}app & \cellcolor{red!10}credit & \cellcolor{red!10}download & \cellcolor{red!10}post & \cellcolor{red!10}issued & \cellcolor{red!10}agent & \mybar{57} \\ 
   66 & \cellcolor{white}wetland & \cellcolor{white}vernal & \cellcolor{white}riparian & \cellcolor{white}habitat & \cellcolor{white}specie & \cellcolor{white}species & \mybar{1040} \\ 
   44 & \cellcolor{white}findings & \cellcolor{white}string & \cellcolor{white}tank & \cellcolor{white}carcinogen & \cellcolor{white}qty & \cellcolor{white}lust & \mybar{128} \\ 
   42 & \cellcolor{white}contamination & \cellcolor{white}spill & \cellcolor{white}remediation & \cellcolor{white}groundwater & \cellcolor{white}asbestos & \cellcolor{white}hazardous & \mybar{343} \\ 
   99 & \cellcolor{white}prep & \cellcolor{white}batch & \cellcolor{white}qualifier & \cellcolor{white}analytical & \cellcolor{white}surrogate & \cellcolor{white}sample & \mybar{313} \\ 
   84 & \cellcolor{white}airport & \cellcolor{white}facility & \cellcolor{white}aviation & \cellcolor{white}maintenance & \cellcolor{white}operation & \cellcolor{white}aircraft & \mybar{150} \\ 
   19 & \cellcolor{white}accessory & \cellcolor{white}height & \cellcolor{white}dwell & \cellcolor{white}frontage & \cellcolor{white}setback & \cellcolor{white}subsection & \mybar{218} \\ 
    6 & \cellcolor{white}householder & \cellcolor{white}poverty & \cellcolor{white}disability & \cellcolor{white}married & \cellcolor{white}husband & \cellcolor{white}universe & \mybar{93} \\ 
   98 & \cellcolor{white}obesity & \cellcolor{white}sugary & \cellcolor{white}epidemic & \cellcolor{white}soda & \cellcolor{white}sensible & \cellcolor{white}drink & \mybar{65} \\ 
   33 & \cellcolor{white}avenue & \cellcolor{white}street & \cellcolor{white}west & \cellcolor{white}east & \cellcolor{white}boulevard & \cellcolor{white}south & \mybar{98} \\ 
   10 & \cellcolor{white}deductible & \cellcolor{white}copay & \cellcolor{white}prescription & \cellcolor{white}coinsurance & \cellcolor{white}outpatient & \cellcolor{white}inpatient & \mybar{488} \\ 
   50 & \cellcolor{white}ductile & \cellcolor{white}trench & \cellcolor{white}pipe & \cellcolor{white}manhole & \cellcolor{white}coupling & \cellcolor{white}compaction & \mybar{705} \\ 
   17 & \cellcolor{white}margin & \cellcolor{white}error & \cellcolor{white}occupied & \cellcolor{white}race & \cellcolor{white}occupy & \cellcolor{white}islander & \mybar{79} \\ 
    5 & \cellcolor{white}earthquake & \cellcolor{white}flood & \cellcolor{white}floodplain & \cellcolor{white}flooding & \cellcolor{white}landslide & \cellcolor{white}fault & \mybar{723} \\ 
   76 & \cellcolor{white}variance & \cellcolor{white}setback & \cellcolor{white}yard & \cellcolor{white}exception & \cellcolor{white}fence & \cellcolor{white}front & \mybar{94} \\ 
   16 & \cellcolor{white}business & \cellcolor{white}marijuana & \cellcolor{white}cannabis & \cellcolor{white}manufacturing & \cellcolor{white}industry & \cellcolor{white}collective & \mybar{319} \\ 
  108 & \cellcolor{blue!10}fugitive & \cellcolor{blue!10}bio & \cellcolor{blue!10}exhaust & \cellcolor{blue!10}unmitigated & \cellcolor{blue!10}noise & \cellcolor{blue!10}receptor & \mybar{262} \\ 
   \hline
\end{tabular}
\endgroup
\caption{Top words from a structural topic model with 120 topics (first 60 topics displayed here) and FREX scoring. Colors depict partisanship based on coefficient size. White cells are non-significant topics.} 
\label{tabSTMtopwords120_1}
\end{table}