### 2019-05-23 paper/tables/stmTopWords60_income.tex
added stm results on income and education as tables

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Thu May 23 15:35:04 2019
\begin{table}[ht]
\centering
\begingroup\scriptsize
\begin{tabular}{rllllllll}
  \hline
 \# & Top Word 1 & Top Word 2 & Top Word 3 & Top Word 4 & Top Word 5 & Top Word 6 & \multicolumn{2}{c}{Tokens assigned} \\ 
  \hline
 46 & \cellcolor{orange!60}chair & \cellcolor{orange!60}subcommittee & \cellcolor{orange!60}speaker & \cellcolor{orange!60}agenda & \cellcolor{orange!60}committee & \cellcolor{orange!60}commission & \mybar{446} \\ 
   19 & \cellcolor{orange!40}setback & \cellcolor{orange!40}variance & \cellcolor{orange!40}zoning & \cellcolor{orange!40}height & \cellcolor{orange!40}yard & \cellcolor{orange!40}accessory & \mybar{453} \\ 
    5 & \cellcolor{orange!30}draft & \cellcolor{orange!30}comment & \cellcolor{orange!30}review & \cellcolor{orange!30}revision & \cellcolor{orange!30}clarify & \cellcolor{orange!30}process & \mybar{356} \\ 
   58 & \cellcolor{orange!30}budget & \cellcolor{orange!30}revenue & \cellcolor{orange!30}adopted & \cellcolor{orange!30}balance & \cellcolor{orange!30}transfer & \cellcolor{orange!30}expenditure & \mybar{176} \\ 
   39 & \cellcolor{orange!30}downtown & \cellcolor{orange!30}mixed & \cellcolor{orange!30}retail & \cellcolor{orange!30}waterfront & \cellcolor{orange!30}orient & \cellcolor{orange!30}density & \mybar{419} \\ 
   50 & \cellcolor{orange!20}trench & \cellcolor{orange!20}manhole & \cellcolor{orange!20}ductile & \cellcolor{orange!20}excavation & \cellcolor{orange!20}pipe & \cellcolor{orange!20}grout & \mybar{1436} \\ 
    9 & \cellcolor{orange!20}trust & \cellcolor{orange!20}revocable & \cellcolor{orange!20}planned & \cellcolor{orange!20}mfr & \cellcolor{orange!20}apportionment & \cellcolor{orange!20}exhibit & \mybar{361} \\ 
    1 & \cellcolor{orange!20}absent & \cellcolor{orange!20}preside & \cellcolor{orange!20}authorize & \cellcolor{orange!20}ordained & \cellcolor{orange!20}int & \cellcolor{orange!20}tag & \mybar{377} \\ 
    4 & \cellcolor{orange!20}audit & \cellcolor{orange!20}auditor & \cellcolor{orange!20}procedure & \cellcolor{orange!20}timely & \cellcolor{orange!20}implemented & \cellcolor{orange!20}oversight & \mybar{472} \\ 
   25 & \cellcolor{orange!20}mitigation & \cellcolor{orange!20}impact & \cellcolor{orange!20}significant & \cellcolor{orange!20}adverse & \cellcolor{orange!20}environmental & \cellcolor{orange!20}measure & \mybar{217} \\ 
   45 & \cellcolor{orange!10}governmental & \cellcolor{orange!10}asset & \cellcolor{orange!10}actuarial & \cellcolor{orange!10}liability & \cellcolor{orange!10}financial & \cellcolor{orange!10}statement & \mybar{235} \\ 
   47 & \cellcolor{orange!10}effluent & \cellcolor{orange!10}inf & \cellcolor{orange!10}eff & \cellcolor{orange!10}infiltration & \cellcolor{orange!10}discharge & \cellcolor{orange!10}sludge & \mybar{751} \\ 
   12 & \cellcolor{orange!10}craftsman & \cellcolor{orange!10}architecture & \cellcolor{orange!10}brick & \cellcolor{orange!10}distinctive & \cellcolor{orange!10}revival & \cellcolor{orange!10}storefront & \mybar{1731} \\ 
   10 & \cellcolor{orange!10}grievance & \cellcolor{orange!10}deductible & \cellcolor{orange!10}coinsurance & \cellcolor{orange!10}dependent & \cellcolor{orange!10}employee & \cellcolor{orange!10}copay & \mybar{583} \\ 
   36 & \cellcolor{orange!10}respondent & \cellcolor{orange!10}compare & \cellcolor{orange!10}figure & \cellcolor{orange!10}trend & \cellcolor{orange!10}appendix & \cellcolor{orange!10}satisfied & \mybar{696} \\ 
   20 & \cellcolor{orange!10}customer & \cellcolor{orange!10}renewable & \cellcolor{orange!10}efficiency & \cellcolor{orange!10}energy & \cellcolor{orange!10}saving & \cellcolor{orange!10}conservation & \mybar{652} \\ 
   48 & \cellcolor{orange!10}contributor & \cellcolor{orange!10}filer & \cellcolor{orange!10}officeholder & \cellcolor{orange!10}political & \cellcolor{orange!10}rouge & \cellcolor{orange!10}payee & \mybar{293} \\ 
   56 & \cellcolor{orange!10}savings & \cellcolor{orange!10}neighborhood & \cellcolor{orange!10}village & \cellcolor{orange!10}excise & \cellcolor{orange!10}ltd & \cellcolor{orange!10}matrix & \mybar{131} \\ 
   60 & \cellcolor{orange!10}bicycle & \cellcolor{orange!10}bike & \cellcolor{orange!10}pedestrian & \cellcolor{orange!10}route & \cellcolor{orange!10}sidewalk & \cellcolor{orange!10}bicyclist & \mybar{561} \\ 
    3 & \cellcolor{orange!10}wetland & \cellcolor{orange!10}specie & \cellcolor{orange!10}species & \cellcolor{orange!10}vernal & \cellcolor{orange!10}ecological & \cellcolor{orange!10}riparian & \mybar{2293} \\ 
   28 & \cellcolor{orange!10}garland & \cellcolor{orange!10}assoc & \cellcolor{orange!10}association & \cellcolor{orange!10}firefighter & \cellcolor{orange!10}duke & \cellcolor{orange!10}xerox & \mybar{480} \\ 
   51 & \cellcolor{orange!10}vent & \cellcolor{orange!10}combustible & \cellcolor{orange!10}flammable & \cellcolor{orange!10}egress & \cellcolor{orange!10}ceiling & \cellcolor{orange!10}extinguisher & \mybar{1160} \\ 
   43 & \cellcolor{orange!10}medicare & \cellcolor{orange!10}payroll & \cellcolor{orange!10}blanket & \cellcolor{orange!10}contractual & \cellcolor{orange!10}undistributed & \cellcolor{orange!10}dept & \mybar{322} \\ 
   52 & \cellcolor{white}homeless & \cellcolor{white}homelessness & \cellcolor{white}affordable & \cellcolor{white}supportive & \cellcolor{white}housing & \cellcolor{white}affordability & \mybar{394} \\ 
   31 & \cellcolor{white}student & \cellcolor{white}teacher & \cellcolor{white}preschool & \cellcolor{white}academic & \cellcolor{white}kindergarten & \cellcolor{white}youth & \mybar{855} \\ 
   22 & \cellcolor{white}allocation & \cellcolor{white}subtotal & \cellcolor{white}admin & \cellcolor{white}cost & \cellcolor{white}yon & \cellcolor{white}allocate & \mybar{190} \\ 
   32 & \cellcolor{white}canceled & \cellcolor{white}dwelling & \cellcolor{white}suite & \cellcolor{white}ave & \cellcolor{white}tad & \cellcolor{white}alteration & \mybar{491} \\ 
   29 & \cellcolor{white}margin & \cellcolor{white}error & \cellcolor{white}disability & \cellcolor{white}speak & \cellcolor{white}employed & \cellcolor{white}language & \mybar{180} \\ 
    7 & \cellcolor{white}fugitive & \cellcolor{white}bio & \cellcolor{white}emission & \cellcolor{white}coal & \cellcolor{white}unmitigated & \cellcolor{white}exhaust & \mybar{773} \\ 
   11 & \cellcolor{white}obesity & \cellcolor{white}sugary & \cellcolor{white}epidemic & \cellcolor{white}drink & \cellcolor{white}calorie & \cellcolor{white}sensible & \mybar{96} \\ 
   34 & \cellcolor{white}playground & \cellcolor{white}recreation & \cellcolor{white}picnic & \cellcolor{white}park & \cellcolor{white}restroom & \cellcolor{white}zoo & \mybar{546} \\ 
   40 & \cellcolor{white}amt & \cellcolor{white}invoice & \cellcolor{white}acct & \cellcolor{white}exp & \cellcolor{white}unencumbered & \cellcolor{white}encumbrance & \mybar{116} \\ 
   53 & \cellcolor{white}applied & \cellcolor{white}col & \cellcolor{white}dist & \cellcolor{white}occupancy & \cellcolor{white}monoxide & \cellcolor{white}valuation & \mybar{128} \\ 
   18 & \cellcolor{teal!10}perm & \cellcolor{teal!10}queue & \cellcolor{teal!10}delay & \cellcolor{teal!10}peak & \cellcolor{teal!10}adj & \cellcolor{teal!10}flt & \mybar{187} \\ 
   55 & \cellcolor{teal!10}taxable & \cellcolor{teal!10}deed & \cellcolor{teal!10}res & \cellcolor{teal!10}homestead & \cellcolor{teal!10}value & \cellcolor{teal!10}book & \mybar{87} \\ 
    6 & \cellcolor{teal!10}race & \cellcolor{teal!10}householder & \cellcolor{teal!10}islander & \cellcolor{teal!10}census & \cellcolor{teal!10}occupied & \cellcolor{teal!10}female & \mybar{160} \\ 
   24 & \cellcolor{white}mail & \cellcolor{white}fax & \cellcolor{white}application & \cellcolor{white}click & \cellcolor{white}applicant & \cellcolor{white}copy & \mybar{367} \\ 
    8 & \cellcolor{teal!10}imp & \cellcolor{teal!10}assessor & \cellcolor{teal!10}taxpayer & \cellcolor{teal!10}petition & \cellcolor{teal!10}preliminary & \cellcolor{teal!10}determination & \mybar{91} \\ 
   17 & \cellcolor{teal!10}portfolio & \cellcolor{teal!10}micron & \cellcolor{teal!10}maturity & \cellcolor{teal!10}treasury & \cellcolor{teal!10}yield & \cellcolor{teal!10}investment & \mybar{538} \\ 
   35 & \cellcolor{teal!10}redemption & \cellcolor{teal!10}bond & \cellcolor{teal!10}increment & \cellcolor{teal!10}obligation & \cellcolor{teal!10}proceeds & \cellcolor{teal!10}lease & \mybar{339} \\ 
   38 & \cellcolor{teal!10}para & \cellcolor{teal!10}persona & \cellcolor{teal!10}horas & \cellcolor{teal!10}bud & \cellcolor{teal!10}contracted & \cellcolor{teal!10}ante & \mybar{1334} \\ 
   44 & \cellcolor{teal!10}findings & \cellcolor{teal!10}tank & \cellcolor{teal!10}string & \cellcolor{teal!10}carcinogen & \cellcolor{teal!10}lust & \cellcolor{teal!10}sic & \mybar{255} \\ 
   30 & \cellcolor{teal!10}subcontractor & \cellcolor{teal!10}bid & \cellcolor{teal!10}bidder & \cellcolor{teal!10}proposer & \cellcolor{teal!10}subcontract & \cellcolor{teal!10}bidding & \mybar{512} \\ 
   27 & \cellcolor{teal!10}article & \cellcolor{teal!10}subsection & \cellcolor{teal!10}shall & \cellcolor{teal!10}franchisee & \cellcolor{teal!10}paragraph & \cellcolor{teal!10}meaning & \mybar{658} \\ 
   15 & \cellcolor{teal!10}credit & \cellcolor{teal!10}docket & \cellcolor{teal!10}app & \cellcolor{teal!10}post & \cellcolor{teal!10}download & \cellcolor{teal!10}month & \mybar{61} \\ 
   37 & \cellcolor{teal!10}endorsed & \cellcolor{teal!10}endorse & \cellcolor{teal!10}rescue & \cellcolor{teal!10}assistant & \cellcolor{teal!10}analyst & \cellcolor{teal!10}technician & \mybar{355} \\ 
   14 & \cellcolor{teal!10}accomplishment & \cellcolor{teal!10}grantee & \cellcolor{teal!10}narrative & \cellcolor{teal!10}outcome & \cellcolor{teal!10}grant & \cellcolor{teal!10}recipient & \mybar{255} \\ 
   54 & \cellcolor{teal!10}license & \cellcolor{teal!10}licensee & \cellcolor{teal!10}citation & \cellcolor{teal!10}tow & \cellcolor{teal!10}fee & \cellcolor{teal!10}taxicab & \mybar{710} \\ 
   13 & \cellcolor{teal!10}initiative & \cellcolor{teal!10}outreach & \cellcolor{teal!10}strategy & \cellcolor{teal!10}leadership & \cellcolor{teal!10}engagement & \cellcolor{teal!10}focus & \mybar{502} \\ 
   33 & \cellcolor{teal!10}thence & \cellcolor{teal!10}east & \cellcolor{teal!10}south & \cellcolor{teal!10}corner & \cellcolor{teal!10}west & \cellcolor{teal!10}avenue & \mybar{340} \\ 
   42 & \cellcolor{teal!10}incumbent & \cellcolor{teal!10}prep & \cellcolor{teal!10}batch & \cellcolor{teal!10}qualifier & \cellcolor{teal!10}analytical & \cellcolor{teal!10}examination & \mybar{1091} \\ 
   57 & \cellcolor{teal!10}councilman & \cellcolor{teal!10}introduced & \cellcolor{teal!10}alderman & \cellcolor{teal!10}whereas & \cellcolor{teal!10}resolved & \cellcolor{teal!10}councilwoman & \mybar{615} \\ 
   23 & \cellcolor{teal!20}bag & \cellcolor{teal!20}recyclable & \cellcolor{teal!20}recyclables & \cellcolor{teal!20}reusable & \cellcolor{teal!20}vegetable & \cellcolor{teal!20}bait & \mybar{2254} \\ 
    2 & \cellcolor{teal!20}influenza & \cellcolor{teal!20}infection & \cellcolor{teal!20}vaccine & \cellcolor{teal!20}patient & \cellcolor{teal!20}tuberculosis & \cellcolor{teal!20}hepatitis & \mybar{2980} \\ 
   21 & \cellcolor{teal!20}everybody & \cellcolor{teal!20}think & \cellcolor{teal!20}something & \cellcolor{teal!20}thing & \cellcolor{teal!20}try & \cellcolor{teal!20}want & \mybar{2609} \\ 
   26 & \cellcolor{teal!20}mesa & \cellcolor{teal!20}canyon & \cellcolor{teal!20}via & \cellcolor{teal!20}odd & \cellcolor{teal!20}unidentified & \cellcolor{teal!20}paradise & \mybar{1886} \\ 
   49 & \cellcolor{teal!20}artist & \cellcolor{teal!20}fun & \cellcolor{teal!20}music & \cellcolor{teal!20}beginner & \cellcolor{teal!20}player & \cellcolor{teal!20}prize & \mybar{4565} \\ 
   16 & \cellcolor{teal!30}motion & \cellcolor{teal!30}second & \cellcolor{teal!30}adjourn & \cellcolor{teal!30}carry & \cellcolor{teal!30}unanimous & \cellcolor{teal!30}chairman & \mybar{419} \\ 
   41 & \cellcolor{teal!40}complainant & \cellcolor{teal!40}allegation & \cellcolor{teal!40}defendant & \cellcolor{teal!40}offender & \cellcolor{teal!40}commander & \cellcolor{teal!40}complaint & \mybar{1695} \\ 
   59 & \cellcolor{teal!80}burglary & \cellcolor{teal!80}robbery & \cellcolor{teal!80}theft & \cellcolor{teal!80}homicide & \cellcolor{teal!80}murder & \cellcolor{teal!80}gunshot & \mybar{945} \\ 
   \hline
\end{tabular}
\endgroup
\caption{Top words from a structural topic model with 60 topics and FREX scoring. Colors depict city median income based on coefficient size (wealthier cities are orange, poorer cities are teal). White cells are non-significant topics.} 
\label{tabSTMtopwords60_income}
\end{table}