### 2019-05-23 paper/tables/stmTopWords60_population.tex

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Thu May 23 15:37:29 2019
\begin{table}[ht]
\centering
\begingroup\scriptsize
\begin{tabular}{rllllllll}
  \hline
 \# & Top Word 1 & Top Word 2 & Top Word 3 & Top Word 4 & Top Word 5 & Top Word 6 & \multicolumn{2}{c}{Tokens assigned} \\ 
  \hline
  2 & \cellcolor{cyan!80}influenza & \cellcolor{cyan!80}infection & \cellcolor{cyan!80}vaccine & \cellcolor{cyan!80}patient & \cellcolor{cyan!80}tuberculosis & \cellcolor{cyan!80}hepatitis & \mybar{2980} \\ 
   38 & \cellcolor{cyan!40}para & \cellcolor{cyan!40}persona & \cellcolor{cyan!40}horas & \cellcolor{cyan!40}bud & \cellcolor{cyan!40}contracted & \cellcolor{cyan!40}ante & \mybar{1334} \\ 
   59 & \cellcolor{cyan!30}burglary & \cellcolor{cyan!30}robbery & \cellcolor{cyan!30}theft & \cellcolor{cyan!30}homicide & \cellcolor{cyan!30}murder & \cellcolor{cyan!30}gunshot & \mybar{945} \\ 
   52 & \cellcolor{cyan!30}homeless & \cellcolor{cyan!30}homelessness & \cellcolor{cyan!30}affordable & \cellcolor{cyan!30}supportive & \cellcolor{cyan!30}housing & \cellcolor{cyan!30}affordability & \mybar{394} \\ 
   24 & \cellcolor{cyan!30}mail & \cellcolor{cyan!30}fax & \cellcolor{cyan!30}application & \cellcolor{cyan!30}click & \cellcolor{cyan!30}applicant & \cellcolor{cyan!30}copy & \mybar{367} \\ 
   29 & \cellcolor{cyan!20}margin & \cellcolor{cyan!20}error & \cellcolor{cyan!20}disability & \cellcolor{cyan!20}speak & \cellcolor{cyan!20}employed & \cellcolor{cyan!20}language & \mybar{180} \\ 
   36 & \cellcolor{cyan!20}respondent & \cellcolor{cyan!20}compare & \cellcolor{cyan!20}figure & \cellcolor{cyan!20}trend & \cellcolor{cyan!20}appendix & \cellcolor{cyan!20}satisfied & \mybar{696} \\ 
   41 & \cellcolor{cyan!20}complainant & \cellcolor{cyan!20}allegation & \cellcolor{cyan!20}defendant & \cellcolor{cyan!20}offender & \cellcolor{cyan!20}commander & \cellcolor{cyan!20}complaint & \mybar{1695} \\ 
   13 & \cellcolor{cyan!20}initiative & \cellcolor{cyan!20}outreach & \cellcolor{cyan!20}strategy & \cellcolor{cyan!20}leadership & \cellcolor{cyan!20}engagement & \cellcolor{cyan!20}focus & \mybar{502} \\ 
    6 & \cellcolor{cyan!10}race & \cellcolor{cyan!10}householder & \cellcolor{cyan!10}islander & \cellcolor{cyan!10}census & \cellcolor{cyan!10}occupied & \cellcolor{cyan!10}female & \mybar{160} \\ 
   10 & \cellcolor{cyan!10}grievance & \cellcolor{cyan!10}deductible & \cellcolor{cyan!10}coinsurance & \cellcolor{cyan!10}dependent & \cellcolor{cyan!10}employee & \cellcolor{cyan!10}copay & \mybar{583} \\ 
   31 & \cellcolor{cyan!10}student & \cellcolor{cyan!10}teacher & \cellcolor{cyan!10}preschool & \cellcolor{cyan!10}academic & \cellcolor{cyan!10}kindergarten & \cellcolor{cyan!10}youth & \mybar{855} \\ 
   22 & \cellcolor{cyan!10}allocation & \cellcolor{cyan!10}subtotal & \cellcolor{cyan!10}admin & \cellcolor{cyan!10}cost & \cellcolor{cyan!10}yon & \cellcolor{cyan!10}allocate & \mybar{190} \\ 
   11 & \cellcolor{cyan!10}obesity & \cellcolor{cyan!10}sugary & \cellcolor{cyan!10}epidemic & \cellcolor{cyan!10}drink & \cellcolor{cyan!10}calorie & \cellcolor{cyan!10}sensible & \mybar{96} \\ 
   44 & \cellcolor{cyan!10}findings & \cellcolor{cyan!10}tank & \cellcolor{cyan!10}string & \cellcolor{cyan!10}carcinogen & \cellcolor{cyan!10}lust & \cellcolor{cyan!10}sic & \mybar{255} \\ 
   23 & \cellcolor{cyan!10}bag & \cellcolor{cyan!10}recyclable & \cellcolor{cyan!10}recyclables & \cellcolor{cyan!10}reusable & \cellcolor{cyan!10}vegetable & \cellcolor{cyan!10}bait & \mybar{2254} \\ 
   17 & \cellcolor{cyan!10}portfolio & \cellcolor{cyan!10}micron & \cellcolor{cyan!10}maturity & \cellcolor{cyan!10}treasury & \cellcolor{cyan!10}yield & \cellcolor{cyan!10}investment & \mybar{538} \\ 
    4 & \cellcolor{cyan!10}audit & \cellcolor{cyan!10}auditor & \cellcolor{cyan!10}procedure & \cellcolor{cyan!10}timely & \cellcolor{cyan!10}implemented & \cellcolor{cyan!10}oversight & \mybar{472} \\ 
   42 & \cellcolor{cyan!10}incumbent & \cellcolor{cyan!10}prep & \cellcolor{cyan!10}batch & \cellcolor{cyan!10}qualifier & \cellcolor{cyan!10}analytical & \cellcolor{cyan!10}examination & \mybar{1091} \\ 
   27 & \cellcolor{cyan!10}article & \cellcolor{cyan!10}subsection & \cellcolor{cyan!10}shall & \cellcolor{cyan!10}franchisee & \cellcolor{cyan!10}paragraph & \cellcolor{cyan!10}meaning & \mybar{658} \\ 
   15 & \cellcolor{cyan!10}credit & \cellcolor{cyan!10}docket & \cellcolor{cyan!10}app & \cellcolor{cyan!10}post & \cellcolor{cyan!10}download & \cellcolor{cyan!10}month & \mybar{61} \\ 
   26 & \cellcolor{cyan!10}mesa & \cellcolor{cyan!10}canyon & \cellcolor{cyan!10}via & \cellcolor{cyan!10}odd & \cellcolor{cyan!10}unidentified & \cellcolor{cyan!10}paradise & \mybar{1886} \\ 
   51 & \cellcolor{cyan!10}vent & \cellcolor{cyan!10}combustible & \cellcolor{cyan!10}flammable & \cellcolor{cyan!10}egress & \cellcolor{cyan!10}ceiling & \cellcolor{cyan!10}extinguisher & \mybar{1160} \\ 
    7 & \cellcolor{cyan!10}fugitive & \cellcolor{cyan!10}bio & \cellcolor{cyan!10}emission & \cellcolor{cyan!10}coal & \cellcolor{cyan!10}unmitigated & \cellcolor{cyan!10}exhaust & \mybar{773} \\ 
   18 & \cellcolor{cyan!10}perm & \cellcolor{cyan!10}queue & \cellcolor{cyan!10}delay & \cellcolor{cyan!10}peak & \cellcolor{cyan!10}adj & \cellcolor{cyan!10}flt & \mybar{187} \\ 
   54 & \cellcolor{cyan!10}license & \cellcolor{cyan!10}licensee & \cellcolor{cyan!10}citation & \cellcolor{cyan!10}tow & \cellcolor{cyan!10}fee & \cellcolor{cyan!10}taxicab & \mybar{710} \\ 
   53 & \cellcolor{white}applied & \cellcolor{white}col & \cellcolor{white}dist & \cellcolor{white}occupancy & \cellcolor{white}monoxide & \cellcolor{white}valuation & \mybar{128} \\ 
   48 & \cellcolor{white}contributor & \cellcolor{white}filer & \cellcolor{white}officeholder & \cellcolor{white}political & \cellcolor{white}rouge & \cellcolor{white}payee & \mybar{293} \\ 
   25 & \cellcolor{magenta!10}mitigation & \cellcolor{magenta!10}impact & \cellcolor{magenta!10}significant & \cellcolor{magenta!10}adverse & \cellcolor{magenta!10}environmental & \cellcolor{magenta!10}measure & \mybar{217} \\ 
    9 & \cellcolor{magenta!10}trust & \cellcolor{magenta!10}revocable & \cellcolor{magenta!10}planned & \cellcolor{magenta!10}mfr & \cellcolor{magenta!10}apportionment & \cellcolor{magenta!10}exhibit & \mybar{361} \\ 
    8 & \cellcolor{magenta!10}imp & \cellcolor{magenta!10}assessor & \cellcolor{magenta!10}taxpayer & \cellcolor{magenta!10}petition & \cellcolor{magenta!10}preliminary & \cellcolor{magenta!10}determination & \mybar{91} \\ 
   20 & \cellcolor{magenta!10}customer & \cellcolor{magenta!10}renewable & \cellcolor{magenta!10}efficiency & \cellcolor{magenta!10}energy & \cellcolor{magenta!10}saving & \cellcolor{magenta!10}conservation & \mybar{652} \\ 
   33 & \cellcolor{magenta!10}thence & \cellcolor{magenta!10}east & \cellcolor{magenta!10}south & \cellcolor{magenta!10}corner & \cellcolor{magenta!10}west & \cellcolor{magenta!10}avenue & \mybar{340} \\ 
   56 & \cellcolor{magenta!10}savings & \cellcolor{magenta!10}neighborhood & \cellcolor{magenta!10}village & \cellcolor{magenta!10}excise & \cellcolor{magenta!10}ltd & \cellcolor{magenta!10}matrix & \mybar{131} \\ 
   28 & \cellcolor{magenta!10}garland & \cellcolor{magenta!10}assoc & \cellcolor{magenta!10}association & \cellcolor{magenta!10}firefighter & \cellcolor{magenta!10}duke & \cellcolor{magenta!10}xerox & \mybar{480} \\ 
   12 & \cellcolor{magenta!10}craftsman & \cellcolor{magenta!10}architecture & \cellcolor{magenta!10}brick & \cellcolor{magenta!10}distinctive & \cellcolor{magenta!10}revival & \cellcolor{magenta!10}storefront & \mybar{1731} \\ 
   21 & \cellcolor{magenta!10}everybody & \cellcolor{magenta!10}think & \cellcolor{magenta!10}something & \cellcolor{magenta!10}thing & \cellcolor{magenta!10}try & \cellcolor{magenta!10}want & \mybar{2609} \\ 
   35 & \cellcolor{magenta!10}redemption & \cellcolor{magenta!10}bond & \cellcolor{magenta!10}increment & \cellcolor{magenta!10}obligation & \cellcolor{magenta!10}proceeds & \cellcolor{magenta!10}lease & \mybar{339} \\ 
   45 & \cellcolor{magenta!10}governmental & \cellcolor{magenta!10}asset & \cellcolor{magenta!10}actuarial & \cellcolor{magenta!10}liability & \cellcolor{magenta!10}financial & \cellcolor{magenta!10}statement & \mybar{235} \\ 
   30 & \cellcolor{magenta!10}subcontractor & \cellcolor{magenta!10}bid & \cellcolor{magenta!10}bidder & \cellcolor{magenta!10}proposer & \cellcolor{magenta!10}subcontract & \cellcolor{magenta!10}bidding & \mybar{512} \\ 
   40 & \cellcolor{magenta!10}amt & \cellcolor{magenta!10}invoice & \cellcolor{magenta!10}acct & \cellcolor{magenta!10}exp & \cellcolor{magenta!10}unencumbered & \cellcolor{magenta!10}encumbrance & \mybar{116} \\ 
   55 & \cellcolor{magenta!10}taxable & \cellcolor{magenta!10}deed & \cellcolor{magenta!10}res & \cellcolor{magenta!10}homestead & \cellcolor{magenta!10}value & \cellcolor{magenta!10}book & \mybar{87} \\ 
    3 & \cellcolor{magenta!10}wetland & \cellcolor{magenta!10}specie & \cellcolor{magenta!10}species & \cellcolor{magenta!10}vernal & \cellcolor{magenta!10}ecological & \cellcolor{magenta!10}riparian & \mybar{2293} \\ 
   37 & \cellcolor{magenta!10}endorsed & \cellcolor{magenta!10}endorse & \cellcolor{magenta!10}rescue & \cellcolor{magenta!10}assistant & \cellcolor{magenta!10}analyst & \cellcolor{magenta!10}technician & \mybar{355} \\ 
   32 & \cellcolor{magenta!10}canceled & \cellcolor{magenta!10}dwelling & \cellcolor{magenta!10}suite & \cellcolor{magenta!10}ave & \cellcolor{magenta!10}tad & \cellcolor{magenta!10}alteration & \mybar{491} \\ 
   47 & \cellcolor{magenta!10}effluent & \cellcolor{magenta!10}inf & \cellcolor{magenta!10}eff & \cellcolor{magenta!10}infiltration & \cellcolor{magenta!10}discharge & \cellcolor{magenta!10}sludge & \mybar{751} \\ 
    5 & \cellcolor{magenta!10}draft & \cellcolor{magenta!10}comment & \cellcolor{magenta!10}review & \cellcolor{magenta!10}revision & \cellcolor{magenta!10}clarify & \cellcolor{magenta!10}process & \mybar{356} \\ 
   14 & \cellcolor{magenta!10}accomplishment & \cellcolor{magenta!10}grantee & \cellcolor{magenta!10}narrative & \cellcolor{magenta!10}outcome & \cellcolor{magenta!10}grant & \cellcolor{magenta!10}recipient & \mybar{255} \\ 
   39 & \cellcolor{magenta!10}downtown & \cellcolor{magenta!10}mixed & \cellcolor{magenta!10}retail & \cellcolor{magenta!10}waterfront & \cellcolor{magenta!10}orient & \cellcolor{magenta!10}density & \mybar{419} \\ 
   43 & \cellcolor{magenta!10}medicare & \cellcolor{magenta!10}payroll & \cellcolor{magenta!10}blanket & \cellcolor{magenta!10}contractual & \cellcolor{magenta!10}undistributed & \cellcolor{magenta!10}dept & \mybar{322} \\ 
   60 & \cellcolor{magenta!10}bicycle & \cellcolor{magenta!10}bike & \cellcolor{magenta!10}pedestrian & \cellcolor{magenta!10}route & \cellcolor{magenta!10}sidewalk & \cellcolor{magenta!10}bicyclist & \mybar{561} \\ 
   58 & \cellcolor{magenta!10}budget & \cellcolor{magenta!10}revenue & \cellcolor{magenta!10}adopted & \cellcolor{magenta!10}balance & \cellcolor{magenta!10}transfer & \cellcolor{magenta!10}expenditure & \mybar{176} \\ 
   50 & \cellcolor{magenta!10}trench & \cellcolor{magenta!10}manhole & \cellcolor{magenta!10}ductile & \cellcolor{magenta!10}excavation & \cellcolor{magenta!10}pipe & \cellcolor{magenta!10}grout & \mybar{1436} \\ 
   19 & \cellcolor{magenta!20}setback & \cellcolor{magenta!20}variance & \cellcolor{magenta!20}zoning & \cellcolor{magenta!20}height & \cellcolor{magenta!20}yard & \cellcolor{magenta!20}accessory & \mybar{453} \\ 
   34 & \cellcolor{magenta!20}playground & \cellcolor{magenta!20}recreation & \cellcolor{magenta!20}picnic & \cellcolor{magenta!20}park & \cellcolor{magenta!20}restroom & \cellcolor{magenta!20}zoo & \mybar{546} \\ 
    1 & \cellcolor{magenta!20}absent & \cellcolor{magenta!20}preside & \cellcolor{magenta!20}authorize & \cellcolor{magenta!20}ordained & \cellcolor{magenta!20}int & \cellcolor{magenta!20}tag & \mybar{377} \\ 
   46 & \cellcolor{magenta!20}chair & \cellcolor{magenta!20}subcommittee & \cellcolor{magenta!20}speaker & \cellcolor{magenta!20}agenda & \cellcolor{magenta!20}committee & \cellcolor{magenta!20}commission & \mybar{446} \\ 
   57 & \cellcolor{magenta!30}councilman & \cellcolor{magenta!30}introduced & \cellcolor{magenta!30}alderman & \cellcolor{magenta!30}whereas & \cellcolor{magenta!30}resolved & \cellcolor{magenta!30}councilwoman & \mybar{615} \\ 
   16 & \cellcolor{magenta!50}motion & \cellcolor{magenta!50}second & \cellcolor{magenta!50}adjourn & \cellcolor{magenta!50}carry & \cellcolor{magenta!50}unanimous & \cellcolor{magenta!50}chairman & \mybar{419} \\ 
   49 & \cellcolor{magenta!80}artist & \cellcolor{magenta!80}fun & \cellcolor{magenta!80}music & \cellcolor{magenta!80}beginner & \cellcolor{magenta!80}player & \cellcolor{magenta!80}prize & \mybar{4565} \\ 
   \hline
\end{tabular}
\endgroup
\caption{Top words from a structural topic model with 60 topics and FREX scoring. Colors depict city population based on coefficient size (larger cities are cyan, smaller cities are magenta). White cells are non-significant topics.} 
\label{tabSTMtopwords60_population}
\end{table}