### 2019-05-26 paper/tables/stmTopWords60_KeepEverything.tex
stm results from the keepEverythingExtractor in boilerpipe

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Sun May 26 16:14:58 2019
\begin{table}[ht]
\centering
\begingroup\scriptsize
\begin{tabular}{rllllllll}
  \hline
 \# & Top Word 1 & Top Word 2 & Top Word 3 & Top Word 4 & Top Word 5 & Top Word 6 & \multicolumn{2}{c}{Tokens assigned} \\ 
  \hline
 30 & \cellcolor{red!80}riverfront & \cellcolor{red!80}department & \cellcolor{red!80}authority & \cellcolor{red!80}transit & \cellcolor{red!80}police & \cellcolor{red!80}enjoy & \mybar{212} \\ 
   55 & \cellcolor{red!60}posted & \cellcolor{red!60}dream & \cellcolor{red!60}celebration & \cellcolor{red!60}broadcast & \cellcolor{red!60}ballpark & \cellcolor{red!60}football & \mybar{544} \\ 
    5 & \cellcolor{red!50}trust & \cellcolor{red!50}revocable & \cellcolor{red!50}leisure & \cellcolor{red!50}learn & \cellcolor{red!50}mfr & \cellcolor{red!50}living & \mybar{327} \\ 
   56 & \cellcolor{red!30}click & \cellcolor{red!30}reserved & \cellcolor{red!30}trademark & \cellcolor{red!30}password & \cellcolor{red!30}search & \cellcolor{red!30}connected & \mybar{387} \\ 
   49 & \cellcolor{red!20}chair & \cellcolor{red!20}agenda & \cellcolor{red!20}subcommittee & \cellcolor{red!20}briefing & \cellcolor{red!20}presentation & \cellcolor{red!20}committee & \mybar{416} \\ 
   47 & \cellcolor{red!10}motion & \cellcolor{red!10}second & \cellcolor{red!10}adjourn & \cellcolor{red!10}carry & \cellcolor{red!10}whiting & \cellcolor{red!10}unanimous & \mybar{451} \\ 
   51 & \cellcolor{red!10}storm & \cellcolor{red!10}drain & \cellcolor{red!10}sanitary & \cellcolor{red!10}water & \cellcolor{red!10}sewer & \cellcolor{red!10}infiltration & \mybar{391} \\ 
   43 & \cellcolor{red!10}article & \cellcolor{red!10}subsection & \cellcolor{red!10}shall & \cellcolor{red!10}provision & \cellcolor{red!10}chapter & \cellcolor{red!10}unlawful & \mybar{467} \\ 
    2 & \cellcolor{red!10}virus & \cellcolor{red!10}tuberculosis & \cellcolor{red!10}infection & \cellcolor{red!10}influenza & \cellcolor{red!10}hepatitis & \cellcolor{red!10}cannabis & \mybar{2555} \\ 
    1 & \cellcolor{red!10}councilman & \cellcolor{red!10}whereas & \cellcolor{red!10}alderman & \cellcolor{red!10}resolved & \cellcolor{red!10}yea & \cellcolor{red!10}resolution & \mybar{576} \\ 
   59 & \cellcolor{red!10}subcontractor & \cellcolor{red!10}proposer & \cellcolor{red!10}bidder & \cellcolor{red!10}bid & \cellcolor{red!10}consultant & \cellcolor{red!10}subcontract & \mybar{508} \\ 
   33 & \cellcolor{red!10}eff & \cellcolor{red!10}inf & \cellcolor{red!10}effluent & \cellcolor{red!10}ether & \cellcolor{red!10}batch & \cellcolor{red!10}isomer & \mybar{1240} \\ 
   54 & \cellcolor{red!10}dwelling & \cellcolor{red!10}alteration & \cellcolor{red!10}plumbing & \cellcolor{red!10}canceled & \cellcolor{red!10}plumb & \cellcolor{red!10}mechanical & \mybar{311} \\ 
   12 & \cellcolor{red!10}think & \cellcolor{red!10}something & \cellcolor{red!10}somebody & \cellcolor{red!10}appreciate & \cellcolor{red!10}seem & \cellcolor{red!10}everything & \mybar{2911} \\ 
   21 & \cellcolor{red!10}artist & \cellcolor{red!10}ceremony & \cellcolor{red!10}jazz & \cellcolor{red!10}celebrate & \cellcolor{red!10}prize & \cellcolor{red!10}yoga & \mybar{3326} \\ 
   11 & \cellcolor{red!10}disaster & \cellcolor{red!10}evacuation & \cellcolor{red!10}marshal & \cellcolor{red!10}apparatus & \cellcolor{red!10}tornado & \cellcolor{red!10}aircraft & \mybar{1174} \\ 
    4 & \cellcolor{red!10}craftsman & \cellcolor{red!10}architecture & \cellcolor{red!10}facade & \cellcolor{red!10}distinctive & \cellcolor{red!10}architectural & \cellcolor{red!10}historic & \mybar{1634} \\ 
   34 & \cellcolor{red!10}contributor & \cellcolor{red!10}filer & \cellcolor{red!10}officeholder & \cellcolor{red!10}political & \cellcolor{red!10}payee & \cellcolor{red!10}candidate & \mybar{272} \\ 
   17 & \cellcolor{red!10}assessor & \cellcolor{red!10}informal & \cellcolor{red!10}taxpayer & \cellcolor{red!10}doc & \cellcolor{red!10}viewing & \cellcolor{red!10}determination & \mybar{442} \\ 
   18 & \cellcolor{red!10}setback & \cellcolor{red!10}variance & \cellcolor{red!10}plat & \cellcolor{red!10}height & \cellcolor{red!10}thence & \cellcolor{red!10}frontage & \mybar{472} \\ 
   19 & \cellcolor{red!10}findings & \cellcolor{red!10}tank & \cellcolor{red!10}carcinogen & \cellcolor{red!10}string & \cellcolor{red!10}qty & \cellcolor{red!10}yon & \mybar{247} \\ 
    3 & \cellcolor{white}vend & \cellcolor{white}meat & \cellcolor{white}utensil & \cellcolor{white}towel & \cellcolor{white}fat & \cellcolor{white}cheese & \mybar{2791} \\ 
   38 & \cellcolor{white}application & \cellcolor{white}applicant & \cellcolor{white}must & \cellcolor{white}copy & \cellcolor{white}tenant & \cellcolor{white}mail & \mybar{390} \\ 
   60 & \cellcolor{white}wetland & \cellcolor{white}shoreline & \cellcolor{white}vernal & \cellcolor{white}riparian & \cellcolor{white}habitat & \cellcolor{white}marsh & \mybar{1986} \\ 
   26 & \cellcolor{white}student & \cellcolor{white}teacher & \cellcolor{white}classroom & \cellcolor{white}beech & \cellcolor{white}academic & \cellcolor{white}doe & \mybar{832} \\ 
    7 & \cellcolor{blue!10}obesity & \cellcolor{blue!10}sugary & \cellcolor{blue!10}epidemic & \cellcolor{blue!10}drink & \cellcolor{blue!10}sensible & \cellcolor{blue!10}ounce & \mybar{98} \\ 
   20 & \cellcolor{white}garland & \cellcolor{white}invoice & \cellcolor{white}assoc & \cellcolor{white}rouge & \cellcolor{white}baton & \cellcolor{white}vendor & \mybar{524} \\ 
   31 & \cellcolor{blue!10}credit & \cellcolor{blue!10}docket & \cellcolor{blue!10}bbl & \cellcolor{blue!10}agent & \cellcolor{blue!10}month & \cellcolor{blue!10}app & \mybar{61} \\ 
   40 & \cellcolor{white}slideshow & \cellcolor{white}arrow & \cellcolor{white}printer & \cellcolor{white}stumble & \cellcolor{white}blogger & \cellcolor{white}google & \mybar{189} \\ 
    8 & \cellcolor{blue!10}deductible & \cellcolor{blue!10}outpatient & \cellcolor{blue!10}prescription & \cellcolor{blue!10}coinsurance & \cellcolor{blue!10}copay & \cellcolor{blue!10}inpatient & \mybar{809} \\ 
   41 & \cellcolor{blue!10}playground & \cellcolor{blue!10}park & \cellcolor{blue!10}tennis & \cellcolor{blue!10}picnic & \cellcolor{blue!10}viewpoint & \cellcolor{blue!10}ravine & \mybar{405} \\ 
   45 & \cellcolor{blue!10}taxable & \cellcolor{blue!10}res & \cellcolor{blue!10}deed & \cellcolor{blue!10}value & \cellcolor{blue!10}homestead & \cellcolor{blue!10}star & \mybar{106} \\ 
   36 & \cellcolor{blue!10}thickness & \cellcolor{blue!10}fitting & \cellcolor{blue!10}conduit & \cellcolor{blue!10}conductor & \cellcolor{blue!10}ductile & \cellcolor{blue!10}trench & \mybar{1756} \\ 
   22 & \cellcolor{blue!10}noise & \cellcolor{blue!10}mitigation & \cellcolor{blue!10}impact & \cellcolor{blue!10}significant & \cellcolor{blue!10}adverse & \cellcolor{blue!10}sensitive & \mybar{346} \\ 
   35 & \cellcolor{blue!10}householder & \cellcolor{blue!10}universe & \cellcolor{blue!10}margin & \cellcolor{blue!10}poverty & \cellcolor{blue!10}race & \cellcolor{blue!10}census & \mybar{251} \\ 
   15 & \cellcolor{blue!10}imp & \cellcolor{blue!10}amt & \cellcolor{blue!10}micron & \cellcolor{blue!10}rend & \cellcolor{blue!10}land & \cellcolor{blue!10}sustain & \mybar{117} \\ 
   16 & \cellcolor{blue!10}dist & \cellcolor{blue!10}applied & \cellcolor{blue!10}col & \cellcolor{blue!10}occupancy & \cellcolor{blue!10}monoxide & \cellcolor{blue!10}valuation & \mybar{123} \\ 
   58 & \cellcolor{blue!10}pickup & \cellcolor{blue!10}bag & \cellcolor{blue!10}recyclable & \cellcolor{blue!10}bin & \cellcolor{blue!10}landfill & \cellcolor{blue!10}curbside & \mybar{651} \\ 
   46 & \cellcolor{blue!10}contracted & \cellcolor{blue!10}medicare & \cellcolor{blue!10}allocation & \cellcolor{blue!10}subtotal & \cellcolor{blue!10}unencumbered & \cellcolor{blue!10}payroll & \mybar{236} \\ 
   14 & \cellcolor{blue!10}perm & \cellcolor{blue!10}queue & \cellcolor{blue!10}delay & \cellcolor{blue!10}peak & \cellcolor{blue!10}adj & \cellcolor{blue!10}volume & \mybar{232} \\ 
   39 & \cellcolor{blue!10}successor & \cellcolor{blue!10}franchisee & \cellcolor{blue!10}covenant & \cellcolor{blue!10}redemption & \cellcolor{blue!10}bankruptcy & \cellcolor{blue!10}obligation & \mybar{678} \\ 
   44 & \cellcolor{blue!10}prune & \cellcolor{blue!10}circumference & \cellcolor{blue!10}tree & \cellcolor{blue!10}planting & \cellcolor{blue!10}shrub & \cellcolor{blue!10}root & \mybar{1798} \\ 
   23 & \cellcolor{blue!10}tax & \cellcolor{blue!10}increment & \cellcolor{blue!10}deduction & \cellcolor{blue!10}abatement & \cellcolor{blue!10}assessed & \cellcolor{blue!10}levy & \mybar{397} \\ 
    6 & \cellcolor{blue!10}fugitive & \cellcolor{blue!10}exhaust & \cellcolor{blue!10}renewable & \cellcolor{blue!10}bio & \cellcolor{blue!10}emission & \cellcolor{blue!10}coal & \mybar{859} \\ 
   57 & \cellcolor{blue!10}savings & \cellcolor{blue!10}costs & \cellcolor{blue!10}capital & \cellcolor{blue!10}ltd & \cellcolor{blue!10}improvement & \cellcolor{blue!10}excise & \mybar{172} \\ 
   27 & \cellcolor{blue!10}density & \cellcolor{blue!10}mixed & \cellcolor{blue!10}planned & \cellcolor{blue!10}infill & \cellcolor{blue!10}orient & \cellcolor{blue!10}retail & \mybar{365} \\ 
   52 & \cellcolor{blue!10}bicycle & \cellcolor{blue!10}pedestrian & \cellcolor{blue!10}bike & \cellcolor{blue!10}route & \cellcolor{blue!10}bus & \cellcolor{blue!10}curb & \mybar{572} \\ 
   48 & \cellcolor{blue!10}actuarial & \cellcolor{blue!10}asset & \cellcolor{blue!10}governmental & \cellcolor{blue!10}assets & \cellcolor{blue!10}investment & \cellcolor{blue!10}debt & \mybar{342} \\ 
   37 & \cellcolor{blue!10}supervisor & \cellcolor{blue!10}technician & \cellcolor{blue!10}aide & \cellcolor{blue!10}incumbent & \cellcolor{blue!10}employee & \cellcolor{blue!10}trainee & \mybar{758} \\ 
   32 & \cellcolor{blue!10}introduced & \cellcolor{blue!10}absent & \cellcolor{blue!10}councilor & \cellcolor{blue!10}preside & \cellcolor{blue!10}digest & \cellcolor{blue!10}legislator & \mybar{615} \\ 
   28 & \cellcolor{blue!10}persona & \cellcolor{blue!10}para & \cellcolor{blue!10}sin & \cellcolor{blue!10}ante & \cellcolor{blue!10}junta & \cellcolor{blue!10}combo & \mybar{2376} \\ 
   24 & \cellcolor{blue!10}audit & \cellcolor{blue!10}procedure & \cellcolor{blue!10}effectiveness & \cellcolor{blue!10}ensure & \cellcolor{blue!10}software & \cellcolor{blue!10}timely & \mybar{497} \\ 
   42 & \cellcolor{blue!10}budget & \cellcolor{blue!10}endorse & \cellcolor{blue!10}revenue & \cellcolor{blue!10}endorsed & \cellcolor{blue!10}balance & \cellcolor{blue!10}expenditure & \mybar{232} \\ 
   13 & \cellcolor{blue!10}homeless & \cellcolor{blue!10}affordable & \cellcolor{blue!10}affordability & \cellcolor{blue!10}housing & \cellcolor{blue!10}supportive & \cellcolor{blue!10}homelessness & \mybar{380} \\ 
   10 & \cellcolor{blue!20}impound & \cellcolor{blue!20}taxicab & \cellcolor{blue!20}license & \cellcolor{blue!20}cat & \cellcolor{blue!20}dog & \cellcolor{blue!20}neuter & \mybar{700} \\ 
    9 & \cellcolor{blue!20}strategy & \cellcolor{blue!20}stakeholder & \cellcolor{blue!20}focus & \cellcolor{blue!20}goal & \cellcolor{blue!20}engagement & \cellcolor{blue!20}outreach & \mybar{732} \\ 
   25 & \cellcolor{blue!20}profile & \cellcolor{blue!20}executive & \cellcolor{blue!20}bend & \cellcolor{blue!20}sustainability & \cellcolor{blue!20}cleanup & \cellcolor{blue!20}rates & \mybar{110} \\ 
   53 & \cellcolor{blue!30}theft & \cellcolor{blue!30}burglary & \cellcolor{blue!30}fwy & \cellcolor{blue!30}aggravated & \cellcolor{blue!30}aggravate & \cellcolor{blue!30}robbery & \mybar{253} \\ 
   29 & \cellcolor{blue!40}sitemap & \cellcolor{blue!40}clerk & \cellcolor{blue!40}calendar & \cellcolor{blue!40}online & \cellcolor{blue!40}bureau & \cellcolor{blue!40}alert & \mybar{124} \\ 
   50 & \cellcolor{blue!60}arrest & \cellcolor{blue!60}complainant & \cellcolor{blue!60}allegation & \cellcolor{blue!60}shooting & \cellcolor{blue!60}homicide & \cellcolor{blue!60}victim & \mybar{1665} \\ 
   \hline
\end{tabular}
\endgroup
\caption{Top words from a structural topic model with 60 topics and FREX scoring. Colors depict partisanship based on coefficient size. White cells are non-significant topics.} 
\label{tabSTMtopwords60}
\end{table}