### 2021-03-23 paper/jitp_cover_letter/jitp_cover_letter.tex

\documentclass[a4paper]{letter} 
%%%% choose one of the following 4 lines to set colour and/or  date style, 
\usepackage[colour]{leedsletter} % Day, num Month year, and colour (use lp2)
%\usepackage[]{leedsletter} % Black-and-white, with US style date Month num, year
%\usepackage[myukdate]{leedsletter} 
%\usepackage[colour]{leedsletter} 
%%%% end
\department{ \vspace{-1cm} {\bf Department of Political Science}\\[38pt]
\vspace{-1.5cm}
~\\
Pennsylvania State University\\
203 Pond Lab \\
University Park, PA 16802\\ \vspace{.3cm} \\
%%%% edit the following three lines to suit you.  If you add a line, then 38pt -> 26pt above
T~~ 814.865.4598\\
F~~ 814.863.8979\\
E~~ bdesmarais@psu.edu  \\
W~~brucedesmarais.com} \vspace{-3cm}
%%%% end

\alignit{10pt} %%  tweaking parameter - probably no need to change.

%%%  need to edit  user info 
\signature{Bruce A. Desmarais Ph.D.\\[5pt] 
\rlap{DeGrandis-McCourtney Early Career Professor} \\
\rlap{Director, Graduate Programs in Social Data Analytics} \\
\rlap{Associate Director, Center for Social Data Analytics} \\ Department of Political Science\\ Institute for CyberScience}
% end personal details

% end personal details 
%%%%%%%%%%%%%%%%%%%%%% a bove this line is sender specific %%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% and probably only needs edited once %%%%%%%%%%%%
%%% what follows is letter specific

% the next line is specific to this letter, and may be used for "short" letters

\skipit{1.7}  % this amount (default = 1.7) increases 2 gaps: address-date-opening 
 \vspace{-2cm}
\begin{document} 
\begin{letter}

%% if this is a ``short'' or long address, then alter \skipit above

\opening{Dear JITP Editorial Team:}
\medskip

% overview of Ted
%% He took my course, has worked as my RA, is my advisee, has R&R at PSRM

We are very excited for the opportunity to have our manuscript entitled, ``Government Websites As Data:  A methodological pipeline with application to the websites of municipalities in the United States,'' considered for publication at {\em The Journal of Information Technology \& Politics}.  We introduce a novel methodological pipeline, along with statistical software that we have developed, for automatically scraping and processing substantive text from large sets of government websites. Our contribution lies in the fact that our methodology is capable of cutting through heterogeneous file types and site organizations to extract substantively meaningful and representative blocks of plain text. We illustrate this methodology with an application to a large sample of the websites of U.S. municipal governments, and find that the textual content of websites varies systematically with the partisanship of the mayor.

We have presented our work in the form of a Workbench Note. As JITP readers and contributors are likely to engage with research problems that involve the collection of website contents, we see JITP as an ideal outlet in which to publish our research. We appreciate your consideration of our work, and welcome feedback on our manuscript.

\closing{Sincerely,}

\end{letter} 
\end{document}