### 2021-06-04 paper/jitp_cover_letter/jitp_cover_letter_rr.tex
resubmitted files

\documentclass[a4paper]{letter} 
%%%% choose one of the following 4 lines to set colour and/or  date style, 
\usepackage[colour]{leedsletter} % Day, num Month year, and colour (use lp2)
%\usepackage[]{leedsletter} % Black-and-white, with US style date Month num, year
%\usepackage[myukdate]{leedsletter} 
%\usepackage[colour]{leedsletter} 
%%%% end
\department{ \vspace{-1cm} {\bf Department of Political Science}\\[38pt]
\vspace{-1.5cm}
~\\
Pennsylvania State University\\
203 Pond Lab \\
University Park, PA 16802\\ \vspace{.3cm} \\
%%%% edit the following three lines to suit you.  If you add a line, then 38pt -> 26pt above
T~~ 814.865.4598\\
F~~ 814.863.8979\\
E~~ bdesmarais@psu.edu  \\
W~~brucedesmarais.com} \vspace{-3cm}
%%%% end

\alignit{10pt} %%  tweaking parameter - probably no need to change.

%%%  need to edit  user info 
\signature{Bruce A. Desmarais Ph.D.\\[5pt] 
\rlap{DeGrandis-McCourtney Early Career Professor} \\
\rlap{Director, Graduate Programs in Social Data Analytics} \\
\rlap{Associate Director, Center for Social Data Analytics} \\ Department of Political Science\\ Institute for CyberScience}
% end personal details

% end personal details 
%%%%%%%%%%%%%%%%%%%%%% a bove this line is sender specific %%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% and probably only needs edited once %%%%%%%%%%%%
%%% what follows is letter specific

% the next line is specific to this letter, and may be used for "short" letters

\skipit{1.7}  % this amount (default = 1.7) increases 2 gaps: address-date-opening 
 \vspace{-2cm}
\begin{document} 
\begin{letter}

%% if this is a ``short'' or long address, then alter \skipit above

\opening{Dear {\em JITP} Editorial Team:}
\medskip

We greatly appreciate the opportunity to revise and resubmit our manuscript entitled, ``Government Websites As Data:  A methodological pipeline with application to the websites of municipalities in the United States,'' for publication at {\em The Journal of Information Technology \& Politics}.  Revising in response to the reviewer comments has sharpened the communication of our contributions, and improved the paper overall.  We are excited to continue developing our manuscript with the guidance of the {\em JITP} editorial team and reviewers.

\closing{Sincerely,}

\end{letter} 
\end{document}